### 2020-08-13 manuscript_1.0.Rmd
manuscript_1.0.Rm is the manuscript. Introduction started so far. ActCritABC.cpp implements ABC fit if the model to zegni's data. It still needs to be generalized to include sarsa. Also, a few more things to organize the code. Data/ is Zegni's data with statistical analysis

---
title: "Cognitive and ecological aspects determine cleaner performance"
author:
- Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento
  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia - andresqp@gmail.com]
- Zegni Triki^[Department of Zoology, Stockholm University, Stockholm, Sweden]
- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  word_document: default
fig_caption: yes
fontsize: 12pt
header-includes:
- \usepackage{float}
- \usepackage{setspace}\doublespacing
- \linespread{2}
- \usepackage{lineno}
- \linenumbers
keep_tex: yes
keywords: My keywords
linespread: 2
bibliography: Cleaner_learning.bib
style: C:\Users\andre\Zotero\styles\behavioral-ecology-and-sociobiology.csl
spacing: double
abstract: My abstract which has to be long enough to take multiple lines otherwise
  one does not see the effect of single-spacing.

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Cooperation among individuals of the same or different species is ubiquitous in nature. 
Among individuals of the same species it is the basis for the evolution of  higher 
forms of biological organization [@maynardsmith_Major_1998]. 
Among individuals of different species it underlays the biological mechanism for 
fundamental ecological interactions such as pollination, nitrogen fixation or 
parasite removal [@begon_Ecology_2006]. However, when cooperation is costly, 
it is not straigtforward to understand why individuals would be willing to pay that cost [@lehmann_Evolution_2006]. Furthermore, cooperative interactions might open the way for 
one partner to exploit the other, what prevents this type of exploitation? [@herre_Evolution_1999]
One reason for cooperating is if the benefits obtained from a partner´s 
cooperative action are dependent on one's own cooperative act. 
This dependency can arise, for instance, when cooperative partners
choose who they interact with depending on the level of cooperation they offer. 
In line with this, cooperative interactions can be viewed as a market-like situation 
where partner choice mediates the level of exchange between traders [@hammerstein_Biological_2016; @leimar_Cooperation_2010; @noe_Biological_1994a]. Empirical evidence shows
that several iconic examples of cooperation between individuals of different species 
are indeed mediated by partner choice [@hammerstein_Biological_2016]. Biological market theory highlights 
that changes in the supply and demand of goods and services determines the value of 
those commodities, and therefore, can influence partner's actions in the cooperative 
exchange. For instance, in the marine cleaning mutualism, where cleaner fish 
_Labroides dimidiatus_ (cleaners) removes ectoparasites from other coral reef fishes (clients) [@cote_Evolution_2000],
some client species are able to choose among different cleaners, and because of this
obtain better service [@grutter_Cleaner_2003]. Better services means for clients to have priority when 
more than one client is available [@bshary_Choosy_2002], and less cheating (when cleaners eat live tissue
instead of parasites) [@grutter_Cleaner_2003]. However, recent evidence shows that when clients experience
ecological conditions with low cleaner abundance, they are less likely to use partner
choice to leverage better service [@triki_Biological_2019]. On the other side, 
cleaners in those markets with favourable conditions, stop giving priority to choosy clients [@triki_Biological_2019; @triki_Decrease_2018].

## Cited literature

### 2020-08-13 Cartoon_actor.R

@@ -1,0 +1,20 @@
+pref<-rep(NA,1000)
+prob<-rep(NA,1000)
+
+pref[1]<-0;prob<-logist(pref[1],0)
+for(it in seq(2,1000,by=2)) {
+  pref[it]<-pref[it-1]
+  pref[it+1]<-pref[it]+runif(1,0,0.1);
+  prob[it]<-logist(pref[it],0)
+  prob[it+1]<-logist(pref[it+1],0)
+}
+
+png("c:/Users/andre/OneDrive - Universidad de los Andes/Neuchatel/images/ActorCart.png",
+    width = 640,height = 480)
+par(plt=(posPlot()-c(-0.15,-0.08,-0.06,-0.06))*0.9,las=2)
+plot(prob,type="l",xlim = c(0,200),ylim=c(0,1),xaxt="n",yaxt="n",xlab="Trial",
+     cex.lab=3, lwd=4,col="blue",ylab="",)
+axis(2,at=c(0,1),cex.axis=3,line = 0)
+mtext(text = "probability",2,line =1.3,cex=3,las=2)
+box(lwd=5)
+dev.off()

### 2020-08-13 Data/script_stats_Figures.R

@@ -1,0 +1,200 @@
+
+#####loading packages-------------
+
+require(car)
+require(visreg)
+require(ggplot2)
+require(broom)
+require(gridExtra)
+require(cowplot)
+require(EnvStats)
+require(rsq)
+require(psych)
+library("wesanderson")
+require(lme4)
+library(here)
+library(data.table)
+
+
+# set_here(path = "C:/Users/andre/OneDrive - Universidad de los Andes/Neuchatel/Data_model/data/11956158/")
+#####Uploading datasets-------------
+
+## creating work space
+# setwd("~/Google Drive/1.Files_manuscripts/Quinones, Triki and Bshary/Figshare")
+
+here()
+
+market <- fread(here("data","market_model.csv"),header=TRUE,sep=",")
+
+head(market)
+
+describe(market)
+
+
+
+### Standardisation and centering continuous predictors by scaling variables -------
+
+market$sabundance_cleaner<- c(scale(market$abundance_cleaners_100m2, scale=F))
+market$scleaner_large_ratio<- c(scale(market$cleaner_large_ratio, scale=F))
+market$scleaner_small_ratio<- c(scale(market$cleaner_small_ratio, scale=F))
+market$scleaner_client_ratio<- c(scale(market$cleaner_client_ratio, scale=F))
+market$spercentage_swim_off<- c(scale(market$percentage_swim_off, scale=F))
+
+
+## model1: cleaner fish abundance-----
+
+Anova(model1 <- glm(market_binomial~  spercentage_swim_off*
+                           poly(sabundance_cleaner,2 )
+                          ,family=binomial,data=market))
+
+##Anova(model1 <- glmer(market_binomial~  spercentage_swim_off*
+#poly(sabundance_cleaner,2 )+(1|site)+(1|year)
+#,family=binomial,data=market))## As the estimates of the random factors turned out to be zero, indicating no effect of site and year, the random factors were then removed from the models 
+
+summary(model1)
+
+rsq(model1) #explianed variance
+
+tidy(model1,conf.int=T,exponentiate=F)## confidence interval
+
+## checking model assumptions
+plot(resid(model1))
+par(mfrow=c(2,2)); plot(model1); par(mfrow=c(1,1))
+plot(model1, which=4)
+residualPlots(model1)
+influenceIndexPlot(model1)
[…]

### 2020-08-13 abundanceVSvLp.R

@@ -1,5 +1,11 @@
 ######################## Abundance analysis ######################################
 
+# Delete data reset environment ---------------------------------------------------------
+
+rm(list = ls())
+
+rm(list=ls()[grepl('data.frame', sapply(ls(), function(x) class(get(x))))])
+
 # Load libraries and external functions -----------------------------------
 
 library(here)
@@ -9,42 +15,62 @@ source(here("loadData.R"))
 source(here("..","R_files","posPlots.R"))
 source(here("..","R_files",'ternaryAEQP.R'))
 source(here("data2interp.R"))
-library('plotrix')
-library('akima')
-library("vcd")
-library("ggplot2")
-library("visreg")
+require('plotrix')
+require('akima')
+require("vcd")
+require("foreach")
+require("doParallel")
+# library("ggplot2")
+# library("visreg")
 
 
-# Delete data reset environment ---------------------------------------------------------
-
-rm(list = ls())
-
-rm(list=ls()[grepl('data.frame', sapply(ls(), function(x) class(get(x))))])
 
 
 # Load Data FIA --------------------------------------------------------------------------------------
 
-scenario<-"AbundLvp"
+scenario<-"AbundLvpAlphaTh_"
 
-simDir<-"h:/Neuch_simulations/AbundLvp_/"
-
-(listPar<-rep("Vlp",10))
-
-(listVal<-seq(0,0.9,0.1))
+simDir<-"e:/NeuchSims/AC/"
 
 
+(listPar<-rep("alphaTh",1))
+
+(listVal<-0.03)
+
+# numCores <-3
+# registerDoParallel(numCores)
+# 
+# foreach(val = c(0.01,0.02,0.03),.packages = c("data.table","here")) %dopar% {
+#   library(here)
+#   here()
+#   source(here("aesth_par.R"))
+#   source(here("loadData.R"))
+#   source(here("..","R_files","posPlots.R"))
+#   source(here("..","R_files",'ternaryAEQP.R'))
+#   source(here("data2interp.R"))
+#   require('plotrix')
+#   require('akima')
+#   require("vcd")
+  
 
 # param<-getParam(here("Simulations","AbundLvp"),listparam = listPar,values = listVal)
 
[…]

### 2020-08-13 loadData.R

@@ -15,12 +15,13 @@ getFilelist<-
            listparam=NULL, 
            # list strings providing the parameters 
            # of interest
-           values=NULL 
+           values=NULL, 
            # list of values matching the list in 
            # listparam
+           fullNam=FALSE # Boolean. Append full path name
            ){
   posAgen<-c("PAA","FAA","DP","p1")
-  listRaw<-list.files(folder,recursive = TRUE,full.names = TRUE)
+  listRaw<-list.files(folder,recursive = TRUE,full.names = fullNam)
   if(length(listRaw)<1)warning("No files in such location",immediate. = TRUE)
   listRaw<-grep(".txt",listRaw,value = TRUE)
   fullList<-vector("list",4)
@@ -130,7 +131,7 @@ file2timeInter<-function(filename,interV,maxAge=-2){
 
 # Computes RV preference for a certain final proportion of the learning trial --
 
-file2lastProp<-function(filename,prop,outPar=NULL,genfold=NULL)
+file2lastProp<-function(filename,prop,outPar=NULL,genfold=NULL,full.path=FALSE)
 {
   if(length(outPar)>0){
     extPar<-grep(outPar,strsplit(filename,"_/")[[1]],
@@ -138,8 +139,10 @@ file2lastProp<-function(filename,prop,outPar=NULL,genfold=NULL)
     parVal<-as.numeric(gsub("[[:alpha:]]",extPar,replacement = ''))
     extPar<-gsub("[[:digit:]]",extPar,replacement = '')
   }
-  if(is.null(genfold)) tmp<-fread(filename)
-  else tmp<-fread(here(genfold=filename))
+  if(is.null(genfold)) 
+    if(full.path)   tmp<-fread(filename)
+    else tmp<-fread(here("Simulations",filename))
+    else tmp<-fread(here("Simulations",genfold,filename))
   # resPtmp<-as.numeric(gsub("[[:alpha:]]", "", grep('pR',strsplit(filename,'_')[[1]],value=TRUE)))*0.1
   # visPtmp<-as.numeric(gsub("[[:alpha:]]", "", grep('pV',strsplit(filename,'_')[[1]],value=TRUE)))*0.1
   # tmp$resProb<-rep(resPtmp,dim(tmp)[1])

### 2020-08-13 loadFieldData.R

@@ -1,0 +1,45 @@
+library("data.table")
+library("here")
+
+fieldData<-fread(here("Data","market_model.csv"))
+
+str(fieldData)
+
+### Adjust data to fit IBD
+
+fieldData[,rel.abund.cleaners:=abundance_cleaners_100m2/
+         (max(abundance_cleaners_100m2)+min(abundance_cleaners_100m2))]
+fieldData[,`:=`(rel.abund.visitors=(1-rel.abund.cleaners)*abundance_large_100m2/abundance_clients_100m2,
+             rel.abund.residents=(1-rel.abund.cleaners)*abundance_small_100m2/abundance_clients_100m2)]
+
+fieldData[,prob.Vis.Leav:=percentage_swim_off/100]
+
+market.ABC<-fieldData[,.(rel.abund.cleaners,rel.abund.visitors,
+                      rel.abund.residents,prob.Vis.Leav,
+                      market_binomial)]
+
+fwrite(market.ABC,here("data","data_ABC.txt"),row.names = FALSE)
+
+
+# Random graphs
+
+par(mfrow=c(2,2))
+
+hist(fieldData[,abundance_clients_100m2])
+hist(fieldData[,abundance_large_100m2])
+hist(fieldData[,abundance_small_100m2])
+hist(fieldData[,abundance_cleaners_100m2])
+
+fieldData[,`:=`(rel_larg_clients=abundance_large_100m2/
+                  (abundance_large_100m2+abundance_small_100m2+abundance_cleaners_100m2),
+                rel_small_clients=abundance_small_100m2/
+            (abundance_large_100m2+abundance_small_100m2+abundance_cleaners_100m2),
+            rel_cleaners=abundance_cleaners_100m2/
+              (abundance_large_100m2+abundance_small_100m2+abundance_cleaners_100m2),
+            prob.Vis.Leav=percentage_swim_off*100)]
+
+par(mfro=c(2,2))
+
+hist(fieldData[,rel_larg_clients])
+hist(fieldData[,rel_small_clients])
+hist(fieldData[,rel_cleaners])

### 2020-08-13 parTOjson.R

@@ -13,44 +13,47 @@ exedir<-'/./ActCrit.exe'
 
 fileName<-"parameters.json"
 
+scenario<-"AbundLvpAlphaTh"
+
 # Baseline parameter values
-param<-list(totRounds=30000,ResReward=1,VisReward=1,
+param<-list(totRounds=20000,ResReward=1,VisReward=1,
             ResProb=c(0.2),
             VisProb=c(0.2),
             ResProbLeav=0,VisProbLeav=1,negativeRew=-0.5,scenario=0,
             inbr=0,outbr=0,trainingRep=30,forRat=0.0,
             alphaT=0.01,printGen=1,seed=1, gammaRange=c(0,0.8),
-            netaRange=c(0,1),alphaThRange=c(0.01),numlearn=1,
+            netaRange=c(0,1),alphaThRange=c(0.01,0.02,0.03),numlearn=1,
+            propfullPrint = 0.7,
             alphaThNch=0.01,
-            folderL=paste0(here(simsDir),"AbundLvpLong1/"))
+            folderL=paste0(here(simsDir),scenario,"_/"))
 
 clustfolderAnd="/hpcfs/home/a.quinones/Cleaner/AbundLvp_/"
 
-clustfolderNeu=paste0("/home/ubuntu/Cleaner/","AbundLvpLong1_/")
+clustfolderNeu=paste0("/home/ubuntu/AC/",scenario,"_/")
 
 setwd(paste("./",simsDir,sep=""))
 
 
 # Arrays with the values of external parameters
-rangLeav<-seq(0,1,by = 0.1)
+rangLeav<-seq(0,0.3,length.out = 20)
 rangAbund<-seq(0.1,0.9,length=9)
 rangScen<-c(0,1,2,3)
 rangAlphNC<-c(0,0.5,1)
 
 # General folder for analysis
-check_create.dir(here(simsDir),param = rep("AbundLvpLong1",1),
-                 values = "")
+check_create.dir(here(simsDir),param = rep(scenario,1),
+                 values = c(""))
 
-listfolders<-check_create.dir(paste(here(simsDir),"/AbundLvpLong1_/",sep=""),
-                                    param = rep("Vlp",10),
-                              values = rangLeav)
+listfolders<-check_create.dir(here("Simulations",paste0(scenario,"_")),
+                              param = rep("Vlp",length(rangLeav)),
+                              values = round(rangLeav,2))
 
 
[…]

### 2020-08-13 timeIntervals.R

@@ -4,9 +4,9 @@ projDir<-"d:/quinonesa/learning_models_c++/actCrit/"
 simsDir<-"s:/quinonesa/Simulations/actCrit/"
 
 # libraries ---------------------------------------------------------------------------------------
-source('d:/quinonesa/Dropbox/R_files/posPlots.R')
-source(paste(projDir,"aesth_par.R",sep=""))
-source(paste(projDir,"loadData.R",sep = ""))
+source(here("..","R_files","posPlots.R"))
+source(here("aesth_par.R"))
+source(here("loadData.R"))
 library('plotrix')
 
 # Load Data ---------------------------------------------------------------------------------------

### 2020-08-14 ActCritABC.cpp
Bug fixes in the cpp. Implement ABC parameter files. Change data file to tab separated.

@@ -74,7 +74,7 @@ public:
 	// function that updates the value of state-action pairs according to 
 	//current reward and estimates of future values
 	void act(client newOptions[], int &idNewOptions, double &VisProbLeav, 
-		double &ResProbLeav, double VisReward, double ResReward, double inbr, 
+		double ResProbLeav, double VisReward, double ResReward, double inbr, 
 		double outbr,  learnScenario scenario);
 		// function where the agent takes the action, gets reward, see new 
 		//state and chooses future action
@@ -126,7 +126,7 @@ public:
 	virtual void updateThet(int curState) = 0;
 	// function to update the policy parameter Theta
 	int numEst;
-	// Number of estimates characterizing bhavioural options 9 for FAA
+	// Number of estimates characterizing behavioural options 9 for FAA
 	int countExp;
 protected:
 	double values[6];																							
@@ -395,7 +395,7 @@ void agent::getExtenededMarket() {
 }
 
 void agent::act(client newOptions[], int &idNewOptions, double &VisProbLeav, 
-	double &ResProbLeav, double VisReward, double ResReward, double inbr,
+	double ResProbLeav, double VisReward, double ResReward, double inbr,
 	double outbr, learnScenario scenario){
 	// taking action, obatining reward, seeing new state, choosing future action
 	++age;																		
@@ -585,21 +585,12 @@ void draw(client trainingSet[], int rounds, double probRes, double probVis){
 	}
 }
 
-string create_filename(std::string filename, agent &individual,
-	nlohmann::json param, double pV, double pR){
+string create_filename(std::string filename,nlohmann::json param){
 	// name the file with the parameter specifications
-	filename.append("alph");
-	filename.append(douts(individual.getLearnPar(alphaPar)));
-	filename.append("_gamma");
-	filename.append(douts(individual.getLearnPar(gammaPar)));
-	filename.append("_neta");
-	filename.append(douts(individual.getLearnPar(netaPar)));
-	filename.append("_pV");
-	filename.append(douts(pV));
-	filename.append("_pR");
-	filename.append(douts(pR));
[…]

### 2020-08-14 loadFieldData.R

@@ -18,7 +18,7 @@ market.ABC<-fieldData[,.(rel.abund.cleaners,rel.abund.visitors,
                       rel.abund.residents,prob.Vis.Leav,
                       market_binomial)]
 
-fwrite(market.ABC,here("data","data_ABC.txt"),row.names = FALSE)
+fwrite(market.ABC,here("data","data_ABC.txt"),row.names = FALSE,sep = "\t")
 
 
 # Random graphs

### 2020-08-14 parTOjson.R

@@ -2,18 +2,15 @@
 
 library("jsonlite")
 library("here")
-
-projDir<-"D:/quinonesa/learning_models_c++/actCrit/"
+source(here("loadData.R"))
 
 simsDir<-"Simulations"
-
-visualDir<-"D:\\quinonesa/VisualStudio/ActCrit/"
 
 exedir<-'/./ActCrit.exe'
 
 fileName<-"parameters.json"
 
-scenario<-"AbundLvpAlphaTh"
+scenario<-"ABCtest"
 
 # Baseline parameter values
 param<-list(totRounds=20000,ResReward=1,VisReward=1,
@@ -26,6 +23,12 @@ param<-list(totRounds=20000,ResReward=1,VisReward=1,
             propfullPrint = 0.7,
             alphaThNch=0.01,
             folderL=paste0(here(simsDir),scenario,"_/"))
+
+param_ABC<-list(totRounds=20000,ResReward=1,VisReward=1,
+            ResProbLeav=0,scenario=0, inbr=0,outbr=0,forRat=0.0,
+            seed=1, propfullPrint = 0.7,sdPert=0.01,chain_length=10000,
+            init=c(0,0,0,0),
+            folderL=paste0(here(simsDir),"/",scenario,"_/"))
 
 clustfolderAnd="/hpcfs/home/a.quinones/Cleaner/AbundLvp_/"
 
@@ -90,6 +93,42 @@ for (i in 1:length(rangLeav)) {
 }
 # 
 for (k in 1:90) print(y)
+
+# Generate json parameter files for ABC fit-------------------------------------
+
+check_create.dir(here(simsDir),param = rep(scenario,1),
+                 values = c(""))
+
+param_ABC$folder<-param_ABC$folderL
+outParam<-toJSON(param_ABC,auto_unbox = TRUE,pretty = TRUE)
+fileName<-paste("parameters_ABC",".json",sep="")
+if(file.exists(paste(param_ABC$folderL,fileName,sep = ''))){
+  currFile<-fromJSON(paste(param_ABC$folderL,fileName,sep = ''))
+  if(sum(unlist(currFile)!=unlist(param_ABC))>0){
+    # warning("You are erasing old files!! n\ Check first!!!",immediate. = TRUE)
+    # print("OLD value")
+    # print(unlist(currFile)[unlist(currFile)!=unlist(param)])
+    # print("NEW value")
+    # print(unlist(param)[unlist(currFile)!=unlist(param)])
[…]

### 2021-04-30 manuscript_1.0.Rmd
ANalysis: run succesfully, but does not seem to improve fit Manuscript: finished intro. started methods

@@
 choice to leverage better service [@triki_Biological_2019]. On the other side, 
-cleaners in those markets with favourable conditions, stop giving priority to choosy clients [@triki_Biological_2019; @triki_Decrease_2018].
+cleaners in those markets with favourable conditions, stop giving priority to choosy clients [@triki_Biological_2019; @triki_Decrease_2018]. Thus, it seems that cleaners, as well as clients,
+adjust their strategy to the prevalent market conditions they find in their environment. 
+This adjustement cannot be due to local genetic adaptation, given that cleaners are open water spawners 
+and the environmental changes in different sites have occured in the last decade [@triki_Biological_2019; @triki_Decrease_2018]. Alternatively, a recent cognitive model proposes that differences in 
+cleaners preferences can arise trough the interaction of cleaner's cognitive processes 
+with the environmental conditions [@quinones_Reinforcement_2019]. Under this view,
+cleaners ontogenatically develop their preferences for different clients through learning,
+and some environmental conditions facilitate learning. Thus, the adaptive adjustment 
+seen in cleaners might be due to their learning mechanisms. Despite the general fit of the 
+model explanation to the available data, no systematic parameter fit has been performed to 
+evaluate which combination of model parameters better fit the data.
+
+## Methods
+
+In order to fit the model parameters to the data we used Aproximate Bayesian Computation (ABC).
+ABC is a statistical technique to fit parameters of complicated models to data when we can't explicitely compute the likelihood function.
+Instead, we can compare the outcome of the model, given a set of parameters, to the actual data usually using a set of summary statistics [@beaumont_Approximate_2010a].
+Thus, individual-based models such as the ones developed under the framework of reinforcement learning would be amenable to fit with this technique.
+In here, we will use the data from @triki_Biological_2019 to fit the parameters of the  model analyzed in @quinones_Reinforcement_2019.
+
+The model consist of a set individual-simulations where individuals face a series of choices; where they have to choose between to options.
+Individuals in the model represent cleaners, and the options they face simulate the natural conditions of a cleaning market.
+Thus, individual in the model choose to prioritize  different clients.
+These clients are classified in resident and visitors.
+Residents are those clients that stay in the cleaning queue when they are not given priority; while visitors leave (with a certain probability) they queue after they are not given priority.
+Individuals obtain reward from cleaning a client regardless of the type.
+Everytime individuals face and make a choice they update the probability of making that same choice.
+The update is based on the difference between the expectation of reward and the actual reward the individual obtains (prediction error); and it is carried in the direction that would lead to more reward given the new information.
+Further details of the model implementation can be found in the paper [@quinones_Reinforcement_2019].
+The model analysis points to the relevance of two parameters.
+First, $\gamma$ measures how much individuals include future rewards in their desicion updates.
+If $\gamma=0$, individuals only use the immediate reward obtained from a cleaning interaction.
+As $\gamma$ increases individuals include more in their desicion making the reward obtained from the subsequent choices.
+Second, $\eta$ measures how much individuals include in their reward the fleeing behaviour of visitor clients as a negative component.
+Both of these components allow individuals to use in ther estimates the future effects of their choices.
+The model analysis shows that without either of them, learning individuals can't develop a preference for the visitor client.
+The analysis contrast the effect of these two parameters agaisnt variation in the relative abundance of the two client types.
+The analysis shows that under very high client abundance the presence of  future rewards ($\gamma>0$) no longer favour the preference for the ephemeral clients, this effect does not occur with negative reward mediated by client behaviour ($\eta$).
+Thus, the analysis suggests that future reward as an explanation for cleaner ability to prefer visitors is supported by @triki_Biological_2019 data.
+
+## Results
 

### 2021-04-30 ActCritABC.cpp

@@ -635,7 +635,7 @@ void initializeIndFile(ofstream &chainOutput,nlohmann::json param){
 	string IndFile = create_filename(namedir, param);
 	chainOutput.open(IndFile.c_str());
 	chainOutput << "iteration	" << "alpha_actor	" << "alpha_critic	" <<
-		"gamma	" << "negReward	" << "fit" << endl;
+		"gamma	" << "negReward	" << "fit	" << "ratio" << endl;
 }
 
 
@@ -654,14 +654,28 @@ model_param perturb_parameters(model_param focal_param,json &sim_param) {
 	model_param new_param;
 	// also, you can throw in your own random number generator that you want,
 	// I just add a number N(0, sd);
-	new_param.alphaA = focal_param.alphaA + rnd::normal(-1, sim_param["sdPert"]);
-	clip_low(new_param.alphaA,0);
-	new_param.alphaC = focal_param.alphaC + rnd::normal(0, sim_param["sdPert"]);
-	clip_low(new_param.alphaC, 0);
-	new_param.gamma = focal_param.gamma + rnd::normal(0, sim_param["sdPert"]);
-	clip_range(new_param.alphaA,0,0.99999);
-	new_param.negReward = focal_param.negReward + rnd::normal(0, sim_param["sdPert"]);
-	clip_low(new_param.negReward, 0);
+	if (sim_param["pertScen"] == 0) {
+		new_param.alphaA = focal_param.alphaA + rnd::normal(0, sim_param["sdPert"]);
+		clip_low(new_param.alphaA, 0);
+		new_param.alphaC = focal_param.alphaC + rnd::normal(0, sim_param["sdPert"]);
+		clip_low(new_param.alphaC, 0);
+	}
+	if (sim_param["pertScen"] < 2) {
+		new_param.gamma = focal_param.gamma + rnd::normal(0, sim_param["sdPert"]);
+		clip_range(new_param.gamma, 0, 0.99999);
+		new_param.negReward = focal_param.negReward + rnd::normal(0, sim_param["sdPert"]);
+		clip_low(new_param.negReward, 0);
+	}
+	else if (sim_param["pertScen"] == 2)
+	{
+		new_param.gamma = focal_param.gamma + rnd::normal(0, sim_param["sdPert"]);
+		clip_range(new_param.gamma, 0, 0.99999);
+	}
+	else
+	{
+		new_param.negReward = focal_param.negReward + rnd::normal(0, sim_param["sdPert"]);
+		clip_low(new_param.negReward, 0);
+	}
[…]

### 2021-04-30 analysisABC.R

@@ -5,10 +5,24 @@
 
 require(here)
 require(data.table)
+source("../R_files/posPlots.R")
 
-listFiles<-list.files(here("Simulations","ABCtest_"),full.names = TRUE)
+(listFiles<-list.files(here("Simulations","ABC_gam_Nrew_"),full.names = TRUE))
 
-ABCraw<-fread(listFiles[1])
-names(ABCraw)<-c("iteration","alphaAct","alphaCrit","gamma","NegRew","fit")
+ABCraw<-fread(listFiles[1],)
 
-ABCraw
+head(ABCraw)
+
+par(plt=posPlot(numploty = 2,idploty = 2),mfrow=c(1,1))
+matplot(y=ABCraw[,c(4,5)],x=ABCraw[,1],type="l",lty=1,col=2:5,ylab = "Parameters",
+        xlab="",xaxt="n")
+legend("topleft",col = 2:5,legend = names(ABCraw)[c(4,5)],lty=1)
+par(plt=posPlot(numploty = 2,idploty = 1),new=TRUE)
+matplot(y=ABCraw[,6],x=ABCraw[,1],type="l",yaxt="s",xaxt="s",ylab="Fit",xlab="")
+
+
+hist(ABCraw$gamma)
+
+plot(ratio~iteration,data=ABCraw,type="l",ylab="ratio",xlab="iterations")
+
+rbinom(100,1,prob = 0.5)

### 2021-04-30 loadFieldData.R

@@ -5,6 +5,8 @@ fieldData<-fread(here("Data","market_model.csv"))
 
 str(fieldData)
 
+
+
 ### Adjust data to fit IBD
 
 fieldData[,rel.abund.cleaners:=abundance_cleaners_100m2/

### 2021-04-30 parTOjson.R

@@ -10,7 +10,7 @@ exedir<-'/./ActCrit.exe'
 
 fileName<-"parameters.json"
 
-scenario<-"ABCtest"
+scenario<-"ABC_gam_Nrew"
 
 # Baseline parameter values
 param<-list(totRounds=20000,ResReward=1,VisReward=1,
@@ -26,8 +26,9 @@ param<-list(totRounds=20000,ResReward=1,VisReward=1,
 
 param_ABC<-list(totRounds=20000,ResReward=1,VisReward=1,
             ResProbLeav=0,scenario=0, inbr=0,outbr=0,forRat=0.0,
-            seed=1, propfullPrint = 0.7,sdPert=0.01,chain_length=10000,
-            init=c(0,0,0,0),
+            seed=1, propfullPrint = 0.7,sdPert=0.001,chain_length=1000000,
+            init=c(0.01,0.01,0,0),# alphaA,AlphaC, Gamma, NegRew
+            pertScen = 1, 	# enum perturnScen {all,  bothFut, justGam, justNegRew};
             folderL=paste0(here(simsDir),"/",scenario,"_/"))
 
 clustfolderAnd="/hpcfs/home/a.quinones/Cleaner/AbundLvp_/"

### 2021-06-29 manuscript_1.0.Rmd
Log-likelihood implementation is done. Changes also to the transformation from absolute to relative cleaner abundances. New ABC simulations running for single parameters and with informed initial values. Changes reflected also in the manuscript_1.

@@
 spacing: double
-abstract: My abstract which has to be long enough to take multiple lines otherwise
-  one does not see the effect of single-spacing.
-
+abstract: My abstract.
 ---
@@
 In order to fit the model parameters to the data we used Aproximate Bayesian Computation (ABC).
-ABC is a statistical technique to fit parameters of complicated models to data when we can't explicitely compute the likelihood function.
-Instead, we can compare the outcome of the model, given a set of parameters, to the actual data usually using a set of summary statistics [@beaumont_Approximate_2010a].
+ABC is a statistical technique to fit parameters of complicated models to data when we can't explicitely compute the likelihood as a function of the model parameters.
+Instead, we can compare the outcome of the model, given a set of parameters, to the actual data usually using a set of summary statistics that summarize the fit of the model predictions to the data [@beaumont_Approximate_2010a].
 Thus, individual-based models such as the ones developed under the framework of reinforcement learning would be amenable to fit with this technique.
@@
 
-The model consist of a set individual-simulations where individuals face a series of choices; where they have to choose between to options.
+The model consist of a set individual-simulations where individuals face a series of choices; they have to choose between two options.
 Individuals in the model represent cleaners, and the options they face simulate the natural conditions of a cleaning market.
@@
 These clients are classified in resident and visitors.
-Residents are those clients that stay in the cleaning queue when they are not given priority; while visitors leave (with a certain probability) they queue after they are not given priority.
+Residents are those clients that stay in the cleaning queue when they are not given priority; while visitors leave (with a certain probability) the queue after they are not given priority.
 Individuals obtain reward from cleaning a client regardless of the type.
@@
 If $\gamma=0$, individuals only use the immediate reward obtained from a cleaning interaction.
-As $\gamma$ increases individuals include more in their desicion making the reward obtained from the subsequent choices.
+As $\gamma$ increases individuals include more, in their desicion making, the reward obtained from the subsequent choices.
+That amounts to calculating and using for deision making future expected rewards.
 Second, $\eta$ measures how much individuals include in their reward the fleeing behaviour of visitor clients as a negative component.
@@
 Thus, the analysis suggests that future reward as an explanation for cleaner ability to prefer visitors is supported by @triki_Biological_2019 data.
+However, a combination of future and negative rewards, could be a better explanation for the data.
+In the current analysis, we will use the data from @triki_Biological_2019 to fit some of the model parameters, particularly relevant we will fit $\gamma$ and $\eta$ to find out whether each or a combination of these effects is a better explanation for the pattern seen in the data.
 
-## Results
+### Data explanation
+
+*Insert here brief description of the data and data collection*
+
+In order to capture with the model the relationship between the characteristics of the market and the cleaner's performance we need to translate the absolute abundances of the data to relative abundances.
+This is because in the model, relative abundances of clients define not only the probability of residents and visitors, but also how often the cleaning station is empty (e.g.
+there are no clients to be cleaned).
+Thus, we computed a measure of cleaner abundance for each site relative to total absolute abundances of cleaners and clients.
+As for the client abundances for residents and visitors we computed a relative measure with respect with the total client abundance, and weigthted that by the relative absence of cleaners (1-relative cleaner abundance).
+Thus, all three measures of relative abundance sum up to 1, and can be used in the model as a proxy for the probability of having different options in the cleaning station.
+This translation between absolute and relative abundances is not without caveats.
+For instance, how busy a particular cleaning station is, should depend not only on the absolute abundaces of clients, but also on how often they require cleaning services.
+How often the services are needed in turn should depend on things like parasite abundances.
+These caveats are relevant because they will influence the estimates of the model parameters.
+Nevertheless, we can use the translation and the model estimates to shed some light on the cognitive mechanisms that better explain the behaviour of cleaners.
+
+### Results
 

### 2021-06-29 ActCritABC.cpp

@@ -603,8 +603,23 @@ struct data_point {
 };
 
 struct model_param {
+	//model_param(model_param const &obj);
 	double alphaC, alphaA, gamma, negReward;
+	model_param &operator= (model_param const &rhs) {
+		alphaA = rhs.alphaA;
+		alphaC = rhs.alphaC;
+		gamma = rhs.gamma;
+		negReward = rhs.negReward;
+		return *this;
+	}
 };
+
+//model_param::model_param(model_param const &obj) {
+//	alphaA = obj.alphaA;
+//	alphaC = obj.alphaC;
+//	gamma = obj.gamma;
+//	negReward = obj.negReward;
+//}
 
 vector<data_point> read_Data(ifstream &marketData) {
 	// open file
@@ -640,14 +655,25 @@ void initializeIndFile(ofstream &chainOutput,nlohmann::json param){
 
 
 
+//double calculate_fit(const std::vector< data_point>& empData,
+//					 const std::vector< data_point>& simData) {
+//	double fit = 0.0;
+//	for(int i = 0; i < empData.size(); ++i) {
+//		double d = empData[i].market_exp_success - simData[i].market_exp_success;
+//		fit += d * d;
+//	}			 
+//	return(fit);
+//}
+
 double calculate_fit(const std::vector< data_point>& empData,
-					 const std::vector< data_point>& simData) {
-	double fit = 0.0;
-	for(int i = 0; i < empData.size(); ++i) {
-		double d = empData[i].market_exp_success - simData[i].market_exp_success;
-		fit += d * d;
-	}			 
-	return(fit);
+	const std::vector< data_point>& simData) {
+	double sum_log_likelihood = 0.0;
+	for (int i = 0; i < empData.size(); ++i) {
+		sum_log_likelihood += log(simData[i].market_exp_success*empData[i].market_exp_success +
+			(1 - simData[i].market_exp_success)*(1 - empData[i].market_exp_success)+
+		0.0000000001);
+	}
+	return(sum_log_likelihood);
 }
 
 model_param perturb_parameters(model_param focal_param,json &sim_param) {
@@ -659,8 +685,12 @@ model_param perturb_parameters(model_param focal_param,json &sim_param) {
 		clip_low(new_param.alphaA, 0);
 		new_param.alphaC = focal_param.alphaC + rnd::normal(0, sim_param["sdPert"]);
[…]

### 2021-06-29 loadFieldData.R

@@ -10,11 +10,18 @@ str(fieldData)
 ### Adjust data to fit IBD
 
 fieldData[,rel.abund.cleaners:=abundance_cleaners_100m2/
-         (max(abundance_cleaners_100m2)+min(abundance_cleaners_100m2))]
+         (abundance_cleaners_100m2+abundance_clients_100m2)]
+
+
 fieldData[,`:=`(rel.abund.visitors=(1-rel.abund.cleaners)*abundance_large_100m2/abundance_clients_100m2,
              rel.abund.residents=(1-rel.abund.cleaners)*abundance_small_100m2/abundance_clients_100m2)]
 
 fieldData[,prob.Vis.Leav:=percentage_swim_off/100]
+
+fieldData[,.(min(rel.abund.cleaners),max(rel.abund.cleaners))]
+fieldData[,.(min(rel.abund.visitors),max(rel.abund.visitors))]
+fieldData[,.(min(rel.abund.residents),max(rel.abund.residents))]
+fieldData[rel.abund.residents==min(rel.abund.residents),rel.abund.cleaners*rel.abund.visitors]
 
 market.ABC<-fieldData[,.(rel.abund.cleaners,rel.abund.visitors,
                       rel.abund.residents,prob.Vis.Leav,
@@ -22,6 +29,19 @@ market.ABC<-fieldData[,.(rel.abund.cleaners,rel.abund.visitors,
 
 fwrite(market.ABC,here("data","data_ABC.txt"),row.names = FALSE,sep = "\t")
 
+
+
+str(fieldData)
+
+fieldData[,rowsum(.SD),.SDcol=grep("rel.abund",names(fieldData),value = TRUE)]
+
+plot(data=fieldData,abundance_large_100m2~rel.abund.cleaners)
+abline(lm(data=fieldData,abundance_large_100m2~rel.abund.cleaners))
+
+mod.1<-lm(data=fieldData,abundance_large_100m2~rel.abund.cleaners*rel.abund.visitors)
+summary(mod.1)
+library(car)
+avPlots(mod.1)
 
 # Random graphs
 

### 2021-09-13 ActCritABC.cpp
Implementation of the perturbation kernel with gamma and beta distributions. This also includes calculating the transition proababilities and carrecting for kernel asymmetries in the manhatan ratio. Add data from the individual fish.

@@ -40,7 +40,6 @@ Start date:
 #include <iostream>
 #include <fstream>
 #include <omp.h>
-// #include "tchar.h"
 #include "../Cpp/Routines/C++/RandomNumbers/random.h"
 //H for house pc, E for laptop, M for office
 #include "../Cpp/json.hpp"       
@@ -422,7 +421,7 @@ void agent::update(){
 	// change policy parameter according to TD error
 	currState = mapOptions(cleanOptionsT,choiceT);
 	nextState = mapOptions(cleanOptionsT1, choiceT);
-	delta = currentReward +
+	delta = currentReward -
 		negRew_curr*neta + gamma*values[nextState] - values[currState];
 	// construct the TD error
 	values[currState] += alpha*delta;
@@ -599,7 +598,15 @@ string create_filename(std::string filename,nlohmann::json param){
 
 struct data_point {
 	double rel_abund_clean, rel_abund_visitors, rel_abund_resid,prob_Vis_Leav;
-	double market_exp_success;
+	bool market_exp_success;
+};
+
+struct locat_point {
+	double rel_abund_clean, rel_abund_visitors, rel_abund_resid, prob_Vis_Leav;
+	string site_year;
+	int countMarket, totalMarket;
+	float market_exp_success;
+	double marketPred;
 };
 
 struct model_param {
@@ -621,7 +628,7 @@ struct model_param {
 //	negReward = obj.negReward;
 //}
 
-vector<data_point> read_Data(ifstream &marketData) {
+vector <locat_point> read_locData(ifstream& marketData) {
 	// open file
 	if (!marketData.is_open()) {
 		std::cerr << "error: unable to open data file\n";
@@ -630,14 +637,17 @@ vector<data_point> read_Data(ifstream &marketData) {
 	}
 	string header;
 	getline(marketData,header); // skip header
-	vector<data_point> data_set; 
-	data_point input;
+	vector<locat_point> data_set; 
+	locat_point input;
 	for (;;) { // read data
+		marketData >> input.site_year;
 		marketData >> input.rel_abund_clean;
 		marketData >> input.rel_abund_visitors;
 		marketData >> input.rel_abund_resid;
 		marketData >> input.prob_Vis_Leav;
 		marketData >> input.market_exp_success;
[…]

### 2021-09-13 analysisABC.R

@@ -7,7 +7,9 @@ require(here)
 library(plot3D) 
 require(data.table)
 library(coda)
-source("../R_files/posPlots.R")
+library(mcmcplots)
+library(lattice)
+source(here("../R_files/posPlots.R"))
 
 simdir<-"D:/Neuch_simulations/ABCfit/Simulations/"
 
@@ -15,17 +17,29 @@ scen<-"ABC_gam_Nrew_"
 
 
 (listFiles<-list.files(paste0(simdir,scen),full.names = TRUE))
+(listFiles<-list.files(here("Simulations",scen)))
 ABCruns<-grep("chain",listFiles,value = TRUE)
 
 
+# all in a single object
 ABCraw<-do.call(rbind,lapply(ABCruns, function(file){
   rundata<-fread(here("Simulations",scen,file))
-  seedNum<-strsplit(file,"_")[[1]][4]
+  seedNum<-strsplit(file,"_")[[1]][3]
   seedNum<-as.numeric(gsub("[[:alpha:]]",seedNum,replacement = ''))
   rundata[,seed:=rep(seedNum,dim(rundata)[1])]
 }))
 
-ABCraw<-fread(here("Simulations",scen,ABCruns[2]))
+# One per onject
+ABCraw<-fread(here("Simulations",scen,ABCruns[1]))
+
+# All in a list
+mcmcList<-mcmc.list(do.call(list,lapply(ABCruns, function(file){
+  rundata<-fread(here("Simulations",scen,file))
+  mcmcRun<-mcmc(rundata[,.(gamma,negReward)])
+  return(mcmcRun)
+})))
+
+
 
 
 head(ABCraw)
@@ -47,20 +61,20 @@ par(plt=posPlot(numploty = 2,idploty = 2),mfrow=c(1,1),las=1)
 matplot(y=ABCburned[,gamma],
         x=ABCburned[,iteration],
         type="l",lty=1,col=2:5,ylab = "",
-        xlab="",xaxt="n")
+        xlab="",xaxt="n",lwd=0.1)
 title(main = expression(gamma),line = -2,cex=3)
 par(plt=posPlot(numploty = 2,idploty = 1),new=TRUE)
 matplot(y=ABCburned[,negReward],
[…]

### 2021-09-13 loadFieldData.R

@@ -1,5 +1,9 @@
 library("data.table")
 library("here")
+library("readxl")
+
+
+## Load field data based on the market experiment performance criteria
 
 fieldData<-fread(here("Data","market_model.csv"))
 
@@ -8,7 +12,7 @@ str(fieldData)
 
 ### Adjust data to fit IBD
 
-fieldData[,rel.abund.cleaners:=abundance_cleaners_100m2/
+fieldData[,rel.abund.cleaners:=10*abundance_cleaners_100m2/
          (abundance_cleaners_100m2+abundance_clients_100m2)]
 
 
@@ -22,15 +26,135 @@ fieldData[,.(min(rel.abund.visitors),max(rel.abund.visitors))]
 fieldData[,.(min(rel.abund.residents),max(rel.abund.residents))]
 fieldData[rel.abund.residents==min(rel.abund.residents),rel.abund.cleaners*rel.abund.visitors]
 
-market.ABC<-fieldData[,.(rel.abund.cleaners,rel.abund.visitors,
+market.ABC<-fieldData[,.(site_year,rel.abund.cleaners,rel.abund.visitors,
                       rel.abund.residents,prob.Vis.Leav,
                       market_binomial)]
 
+hist(market.ABC[,rel.abund.cleaners])
+
 fwrite(market.ABC,here("data","data_ABC.txt"),row.names = FALSE,sep = "\t")
 
+marketABC.site<-market.ABC[,lapply(.SD,mean),by=site_year,.SDcols=c("rel.abund.cleaners",
+                                                    "rel.abund.visitors",
+                                                    "rel.abund.residents",
+                                                    "prob.Vis.Leav",
+                                                    "market_binomial")]
+marketABC.site[,countMarket:=market.ABC[,sum(market_binomial),by=site_year]$V1]
+marketABC.site[,totalMarket:=market.ABC[,length(market_binomial),by=site_year]$V1]
+marketABC.site[,site_year:=gsub(" ","_",site_year)]
 
+marketABC.site.fake<-marketABC.site
+marketABC.site.fake[,prob.Vis.Leav:=1]
+marketABC.site.fake[,rel.abund.cleaners:=seq(0.01,0.95,length.out = 12)]
+marketABC.site.fake[,rel.abund.visitors:=(1-rel.abund.cleaners)/2]
[…]

### 2021-09-13 parTOjson.R

@@ -10,7 +10,7 @@ exedir<-'/./ActCrit.exe'
 
 fileName<-"parameters.json"
 
-scenario<-"ABCtest"
+scenario<-"ABC_gam_Nrew"
 
 # Baseline parameter values
 param<-list(totRounds=20000,ResReward=1,VisReward=1,
@@ -24,12 +24,12 @@ param<-list(totRounds=20000,ResReward=1,VisReward=1,
             alphaThNch=0.01,
             folderL=paste0(here(simsDir),scenario,"_/"))
 
-param_ABC<-list(totRounds=5000,ResReward=1,VisReward=1,
+param_ABC<-list(totRounds=10000,ResReward=1,VisReward=1,
             ResProbLeav=0,scenario=0, inbr=0,outbr=0,forRat=0.0,
-            seed=1, propfullPrint = 0.7,sdPert=c(0.1,0.1,0.1,1),
-            chain_length=10000,
+            seed=1, propfullPrint = 0.7,sdPert=c(0.05,0.05,0.001,0.1),
+            chain_length=500000,
             init=c(0.05,0.05,0.8,0.8),# alphaA,AlphaC, Gamma, NegRew
-            pertScen = 3, 	# enum perturnScen {all,  bothFut, justGam, justNegRew};
+            pertScen = 1, 	# enum perturnScen {all,  bothFut, justGam, justNegRew};
             folderL=paste0(here(simsDir),"/",scenario,"_/"))
 
 clustfolderAnd="/hpcfs/home/a.quinones/Cleaner/AbundLvp_/"
@@ -104,7 +104,7 @@ check_create.dir(here(simsDir),param = rep(scenario,1),
 
 for(seed in 1:3){
   param_ABC$folder<-param_ABC$folderL
-  param_ABC$init<-c(0.05,0.05,0,runif(1,max = 10))#runif(1,max = 1),0)#
+  param_ABC$init<-c(0.05,0.05,runif(1,max = 1),runif(1,max = 2))##
   param_ABC$seed <- seed
   fileName<-paste("parameters_ABC_",seed,".json",sep="")
   outParam<-toJSON(param_ABC,auto_unbox = TRUE,pretty = TRUE)

### 2021-10-07 ActCrit.cpp
MCMC is running very well. new fiels are parameter files and outcomes of the chain

@@ -719,9 +719,9 @@ int main(int argc, char* argv[]){
 								double init = tmpGam*(1 - pow(1 - tmpRes - tmpVis, 2)) / (1 - tmpGam);
 
 							   // Initialize agents
-								learners[0] = new PAATyp1(alphaT, *itg, *itn, 
+								learners[1] = new PAATyp1(alphaT, *itg, *itn,
 									*italTh, init, alphaThNch);
-								learners[1] = new FAATyp1(alphaT, *itg, *itn, 
+								learners[0] = new FAATyp1(alphaT, *itg, *itn, 
 									*italTh, init);
 								// output of learning trials
 								ofstream printTest;

### 2021-10-07 ActCritABC.cpp

@@ -596,15 +596,21 @@ string create_filename(std::string filename,nlohmann::json param){
 	return(filename);
 }
 
-struct data_point {
-	double rel_abund_clean, rel_abund_visitors, rel_abund_resid,prob_Vis_Leav;
-	bool market_exp_success;
-};
+
 
 struct locat_point {
 	double rel_abund_clean, rel_abund_visitors, rel_abund_resid, prob_Vis_Leav;
 	string site_year;
 	int countMarket, totalMarket;
+	float market_exp_success;
+	double marketPred;
+};
+
+struct cleaner_point {
+	double rel_abund_clean, rel_abund_visitors, rel_abund_resid, prob_Vis_Leav;
+	string site_year;
+	string cleanerID;
+	int countVisitor, totalMarket;
 	float market_exp_success;
 	double marketPred;
 };
@@ -640,6 +646,7 @@ vector <locat_point> read_locData(ifstream& marketData) {
 	vector<locat_point> data_set; 
 	locat_point input;
 	for (;;) { // read data
+	    // if end of file
 		marketData >> input.site_year;
 		marketData >> input.rel_abund_clean;
 		marketData >> input.rel_abund_visitors;
@@ -648,12 +655,41 @@ vector <locat_point> read_locData(ifstream& marketData) {
 		marketData >> input.market_exp_success;
 		marketData >> input.countMarket;
 		marketData >> input.totalMarket;
-		data_set.emplace_back(input);
-		// if end of file
 		if (marketData.eof()) 	break;
+		data_set.emplace_back(input);
 	}
 	return(data_set);
 }
+
+vector <cleaner_point> read_CleanData(ifstream& marketData) {
+	// open file
+	if (!marketData.is_open()) {
+		std::cerr << "error: unable to open data file\n";
+		wait_for_return();
+		exit(EXIT_FAILURE);
+	}
+	string header;
+	getline(marketData, header); // skip header
+	vector<cleaner_point> data_set;
+	cleaner_point input;
+	for (;;) { // read data
+		marketData >> input.site_year;
+		marketData >> input.cleanerID;
+		marketData >> input.countVisitor;
+		marketData >> input.rel_abund_clean;
+		marketData >> input.rel_abund_visitors;
+		marketData >> input.rel_abund_resid;
[…]

### 2021-10-07 analysisABC.R

@@ -14,10 +14,10 @@ source(here("../R_files/posPlots.R"))
 
 simdir<-"D:/Neuch_simulations/ABCfit/Simulations/"
 
-scen<-"ABC_gam_Nrew_unif_"
+scen<-"ABCclean_gam_Nrew_exp_sca30_"
 
 
-(listFiles<-list.files(paste0(simdir,scen),full.names = TRUE))
+# (listFiles<-list.files(paste0(simdir,scen),full.names = TRUE))
 (listFiles<-list.files(here("Simulations",scen)))
 ABCruns<-grep("chain",listFiles,value = TRUE)
 
@@ -31,7 +31,7 @@ ABCraw<-do.call(rbind,lapply(ABCruns, function(file){
 }))
 
 # One per onject
-ABCraw<-fread(here("Simulations",scen,ABCruns[3]))
+ABCraw<-fread(here("Simulations",scen,ABCruns[1]))
 
 # All in a list
 mcmcList<-mcmc.list(do.call(list,lapply(ABCruns, function(file){
@@ -57,6 +57,7 @@ ABCfiltered<-ABCraw[iteration>burn.in & iteration %% thinning==0]
 ABCburned[seed==1,]
 ABC.gamma.time.short<-dcast(ABCburned,iteration~seed,value.var = c("gamma"))
 ABC.nRew.time.short<-dcast(ABCburned,iteration~seed,value.var = "negReward")
+ABC.logLike.time.short<-dcast(ABCburned,iteration~seed,value.var = "fit")
 
 par(plt=posPlot(numploty = 2,idploty = 2),mfrow=c(1,1),las=1)
 matplot(y=ABC.gamma.time.short[,c(2,3,4)],
@@ -64,13 +65,19 @@ matplot(y=ABC.gamma.time.short[,c(2,3,4)],
         type="l",lty=1,col=2:5,ylab = "",
         xlab="",xaxt="n",lwd=0.1)
 title(main = expression(gamma),line = -2,cex=3,new=T)
-par(plt=posPlot(numploty = 2,idploty = 1),mfrow=c(1,1),las=1)
+par(plt=posPlot(numploty = 2,idploty = 1),mfrow=c(1,1),las=1,new=TRUE)
 matplot(y=ABC.nRew.time.short[,c(2,3,4)],
         x=ABC.nRew.time.short[,iteration],
[…]

### 2021-10-07 loadFieldData.R

@@ -12,7 +12,9 @@ str(fieldData)
 
 ### Adjust data to fit IBD
 
-fieldData[,rel.abund.cleaners:=10*abundance_cleaners_100m2/
+clean.abund.scale<-30
+
+fieldData[,rel.abund.cleaners:=clean.abund.scale*abundance_cleaners_100m2/
          (abundance_cleaners_100m2+abundance_clients_100m2)]
 
 
@@ -32,7 +34,8 @@ market.ABC<-fieldData[,.(site_year,rel.abund.cleaners,rel.abund.visitors,
 
 hist(market.ABC[,rel.abund.cleaners])
 
-fwrite(market.ABC,here("data","data_ABC.txt"),row.names = FALSE,sep = "\t")
+fwrite(market.ABC,here("data",paste0("data_ABC_",clean.abund.scale,".txt")),
+       row.names = FALSE,sep = "\t")
 
 marketABC.site<-market.ABC[,lapply(.SD,mean),by=site_year,.SDcols=c("rel.abund.cleaners",
                                                     "rel.abund.visitors",
@@ -43,13 +46,13 @@ marketABC.site[,countMarket:=market.ABC[,sum(market_binomial),by=site_year]$V1]
 marketABC.site[,totalMarket:=market.ABC[,length(market_binomial),by=site_year]$V1]
 marketABC.site[,site_year:=gsub(" ","_",site_year)]
 
-marketABC.site.fake<-marketABC.site
-marketABC.site.fake[,prob.Vis.Leav:=1]
-marketABC.site.fake[,rel.abund.cleaners:=seq(0.01,0.95,length.out = 12)]
-marketABC.site.fake[,rel.abund.visitors:=(1-rel.abund.cleaners)/2]
-marketABC.site.fake[,rel.abund.residents:=(1-rel.abund.cleaners)/2]
+# marketABC.site.fake<-marketABC.site
+# marketABC.site.fake[,prob.Vis.Leav:=1]
+# marketABC.site.fake[,rel.abund.cleaners:=seq(0.01,0.95,length.out = 12)]
+# marketABC.site.fake[,rel.abund.visitors:=(1-rel.abund.cleaners)/2]
+# marketABC.site.fake[,rel.abund.residents:=(1-rel.abund.cleaners)/2]
 
-fwrite(marketABC.site,here("data","data_ABC_site.txt"),
[…]

### 2021-10-07 manFigures.R

@@ -4,26 +4,194 @@ library(data.table)
 library(ggplot2)
 library("RColorBrewer")
 library("patchwork") 
+library(here)
+source(here("loadData.R"))
+source(here("data2interp.R"))
+require('akima')
+source(here("aesth_par.R"))
+source(here("..","R_files",'ternaryAEQP.R'))
+source(here("../R_files/posPlots.R"))
 
-fieldData<-fread(here("Data","data_ABC_site.txt"))
-str(fieldData)
-fieldData[,predMarket:=rep(0.5,12)]
 
-ggplot(data = fieldData, 
-       aes(x=rel.abund.cleaners, y=prob.Vis.Leav,size=market_binomial))+
+scenario<-"ABCclean_gam_Nrew_unif_sca20"
+
+
+predfile<-grep("round",list.files(here("Simulations",paste0(scenario,"_"))),value = T)
+
+predictData<-fread(here("Simulations",paste0(scenario,"_"),
+                      predfile))
+
+str(predictData)
+
+fieldatabyLoc<-predictData[,.(probVisi.data=mean(visitorChoices)/20,
+                              probvisitor.pred=max(),
+                              re.abund.clean=max(rel.abund.cleaners),
+                            prob.Vis.leave=max(prob.Vis.Leav)),by=site_year]
+
+
+
+## Obv vs predicted two panels with point size ---------------------------------
+
+field.size<-ggplot(data = predictData, 
+       aes(x=rel.abund.cleaners, y=prob.Vis.Leav,size=visitorChoices/20))+
   geom_point()+theme_classic()
 
-myPalette <- colorRampPalette(rev(brewer.pal(11, "Spectral")))
-sc <- scale_colour_gradientn(colours = myPalette(100), limits=c(0,1))
+sims.size<-ggplot(data = predictData, 
+                   aes(x=rel.abund.cleaners, y=prob.Vis.Leav,size=visitorChoices_pred))+
+  geom_point()+theme_classic()
 
-obs<-ggplot(data = fieldData, 
-       aes(x=rel.abund.cleaners, y=prob.Vis.Leav,color=market_binomial))+
+field.size | sims.size
+
+##  Obv vs predicted by location from cleaner data
+
+field.size<-ggplot(data = fieldatabyLoc, 
[…]

### 2021-10-07 parTOjson.R

@@ -10,42 +10,85 @@ exedir<-'/./ActCrit.exe'
 
 fileName<-"parameters.json"
 
-scenario<-"ABC_gam_Nrew_unif2"
+scenario<-"ABCclean_gam_Nrew_unif_sca30"
 
-scenario<-"predictions"
+# scenario<-"predictions"
 
-# Baseline parameter values
-param<-list(totRounds=20000,ResReward=1,VisReward=1,
-            ResProb=c(0.2),
-            VisProb=c(0.2),
-            ResProbLeav=0,VisProbLeav=1,negativeRew=-0.5,scenario=0,
-            inbr=0,outbr=0,trainingRep=30,forRat=0.0,
-            alphaT=0.01,printGen=1,seed=1, gammaRange=c(0,0.8),
-            netaRange=c(0,1),alphaThRange=c(0.01,0.02,0.03),numlearn=1,
-            propfullPrint = 0.7,
-            alphaThNch=0.01,
-            folderL=paste0(here(simsDir),scenario,"_/"))
+# For predictions on the observed values ---------------------------------------
 
-param_ABC<-list(totRounds=10000,ResReward=1,VisReward=1,
+scenario<-"ABCclean_gam_Nrew_exp_sca30"
+
+param_pred<-list(totRounds=10000,ResReward=1,VisReward=1,
             ResProbLeav=0,scenario=0, inbr=0,outbr=0,forRat=0.0,
-            seed=1, propfullPrint = 0.7,sdPert=c(0.05,0.05,0.2,0.2),
-            chain_length=500000,
+            seed=1, propfullPrint = 0.7,sdPert=c(0.05,0.05,0.1,0.05),
+            chain_length=0,
             init=c(0.05,0.05,0,0),# alphaA,AlphaC, Gamma, NegRew
             pertScen = 1, 	# enum perturnScen {all,  bothFut, justGam, justNegRew};
-            MCMC =0,
+            MCMC =0, data="clean",nRep=30,
+            dataFile = here("Data","data_ABC_cleaner_30.txt"),
             folderL=paste0(here(simsDir),"/",scenario,"_/"))
+
+
+
+
+param_pred$folder<-param_pred$folderL
+param_pred$init[c(3,4)]<-sumMCMClist$statistics[,1]
+fileName<-paste("parameters_pred",param_pred$seed,".json",sep="")
+outParam.pred<-toJSON(param_pred,auto_unbox = TRUE,pretty = TRUE)
[…]

### 2021-10-21 manuscript_1.0.Rmd
Implementation of the perturbation kernel with gamma and beta distributions. This also includes calculating the transition proababilities and carrecting for kernel asymmetries in the manhatan ratio. Add data from the individual fish.
MCMC is running very well. new fiels are parameter files and outcomes of the chain
New set up to fit also de scaling constant. A bit of progress in the manuscript. Got things ready to run fitting procedure on simulated data. Started test to use R package to run MCMCs with any typpe of model.

@@
 knitr::opts_chunk$set(echo = TRUE)
+library(here)
 ```
@@
 
-In order to fit the model parameters to the data we used Aproximate Bayesian Computation (ABC).
-ABC is a statistical technique to fit parameters of complicated models to data when we can't explicitely compute the likelihood as a function of the model parameters.
-Instead, we can compare the outcome of the model, given a set of parameters, to the actual data usually using a set of summary statistics that summarize the fit of the model predictions to the data [@beaumont_Approximate_2010a].
-Thus, individual-based models such as the ones developed under the framework of reinforcement learning would be amenable to fit with this technique.
-In here, we will use the data from @triki_Biological_2019 to fit the parameters of the  model analyzed in @quinones_Reinforcement_2019.
+In order to fit the model parameters to the data we used Markov Chain Monte Carlo (MCMC) with 
+a Metropolis sampler to estimate the posterior distribution of the model parameters.
+We used the data from @triki_Biological_2019 to fit the parameters of the  model analyzed in @quinones_Reinforcement_2019.
+The data consist a of the number of times that set of cleaners chose to
+feed of a client plate in the context of the market task.
+This set of cleaners were collected under different locations and in each one of those locations environmental variables were measured.
+These environmental variables recorded from each location are used as input for the computational model.
+Given these inputs, the computational model produces a time series of choices in the market task.
+In the model the preference for the two option changes as the algorithm learns based on the reward that the two options provide.
+Thus, we run the model for 10000 steps and use the las 3000 steps as the preference developed in the model for visitors given the environmental variables.
+We use this preference as the predicted probability of choosing a visitor in the data, and use it to calculate the likelihood that the number of visitor choices obtained from each cleaner is generated by the model.
+We sum over all the cleaners the loglikehoods and use the total for the acceptance of new parameters in the MCMC.
 
-The model consist of a set individual-simulations where individuals face a series of choices; they have to choose between two options.
+The model consist of a set individual-simulations where individuals face a series of choices between two options.
 Individuals in the model represent cleaners, and the options they face simulate the natural conditions of a cleaning market.
-Thus, individual in the model choose to prioritize  different clients.
-These clients are classified in resident and visitors.
-Residents are those clients that stay in the cleaning queue when they are not given priority; while visitors leave (with a certain probability) the queue after they are not given priority.
+Thus, individuals in the model choose to prioritize  different clients.
+These clients are classified as resident and visitors.
+Residents are those clients that stay in the cleaning queue when they are not given priority; while visitors leave the queue (with a certain probability) after they are not given priority.
 Individuals obtain reward from cleaning a client regardless of the type.
@@
 If $\gamma=0$, individuals only use the immediate reward obtained from a cleaning interaction.
-As $\gamma$ increases individuals include more, in their desicion making, the reward obtained from the subsequent choices.
+As $\gamma$ increases individuals include more the reward obtained from the subsequent choices.
 That amounts to calculating and using for deision making future expected rewards.
@@
 Both of these components allow individuals to use in ther estimates the future effects of their choices.
-The model analysis shows that without either of them, learning individuals can't develop a preference for the visitor client.
+The model analysis shows that without either of them, learning individuals cannot develop a preference for the visitor client.
 The analysis contrast the effect of these two parameters agaisnt variation in the relative abundance of the two client types.
-The analysis shows that under very high client abundance the presence of  future rewards ($\gamma>0$) no longer favour the preference for the ephemeral clients, this effect does not occur with negative reward mediated by client behaviour ($\eta$).
+The analysis shows that under very high client abundance the presence of  future rewards ($\gamma>0$) no longer favours the preference for the ephemeral clients, this effect does not occur with negative reward mediated by client behaviour ($\eta$).
 Thus, the analysis suggests that future reward as an explanation for cleaner ability to prefer visitors is supported by @triki_Biological_2019 data.
-However, a combination of future and negative rewards, could be a better explanation for the data.
+However, a combination of future and negative rewards, could still be a better explanation for the data.
 In the current analysis, we will use the data from @triki_Biological_2019 to fit some of the model parameters, particularly relevant we will fit $\gamma$ and $\eta$ to find out whether each or a combination of these effects is a better explanation for the pattern seen in the data.
@@
 
-In order to capture with the model the relationship between the characteristics of the market and the cleaner's performance we need to translate the absolute abundances of the data to relative abundances.
+In order to capture with the model the relationship between the characteristics of the market (environment) and the cleaner's performance we need to translate the absolute abundances of the data to relative abundances.
 This is because in the model, relative abundances of clients define not only the probability of residents and visitors, but also how often the cleaning station is empty (e.g.
 there are no clients to be cleaned).
-Thus, we computed a measure of cleaner abundance for each site relative to total absolute abundances of cleaners and clients.
-As for the client abundances for residents and visitors we computed a relative measure with respect with the total client abundance, and weigthted that by the relative absence of cleaners (1-relative cleaner abundance).
+The more cleaners, the more options available for the clients and hence higher probability of empty cleaning stations.
+How often the cleaning stations are empty in turns influences the value that cleaners can expect from their choices.
+However, the frequency of empty spots in the cleaning station does not only depend on cleaner relative abundances but also on the frequency with which clients visit the station.
+The client visitation frequency will likely vary among different clients depnding on parasite abundances and loads.
+There are variables for which we do not have field estimates.
+Thus, we use
+computed a measure of relative cleaner abundance for each site relative to absolute abundances and multiplied by scalling constant that changes the range of the variable.
+We fit the value os the scalling constant as part of the statistical inference.
+As for the client abundances for residents and visitors we computed a relative measure with respect to the total client abundance, and weigthted that by the rescaled relative absence of cleaners (1-relative cleaner abundance).
 Thus, all three measures of relative abundance sum up to 1, and can be used in the model as a proxy for the probability of having different options in the cleaning station.
-This translation between absolute and relative abundances is not without caveats.
-For instance, how busy a particular cleaning station is, should depend not only on the absolute abundaces of clients, but also on how often they require cleaning services.
-How often the services are needed in turn should depend on things like parasite abundances.
-These caveats are relevant because they will influence the estimates of the model parameters.
-Nevertheless, we can use the translation and the model estimates to shed some light on the cognitive mechanisms that better explain the behaviour of cleaners.
 
@@
 
+```{r, fig1}
+knitr::include_graphics(here("Simulations","ABCclean_gam_nRew_sca_",                            "round_gamma_0.3regRew_0.11scaC_133.5alphA_0.05alphC_0.05_contour_ggplot.png"))
+```
+
 ## Cited literature

### 2021-10-21 ActCritABC.cpp

@@ -588,12 +588,12 @@ void draw(client trainingSet[], int rounds, double probRes, double probVis){
 
 string create_filename(std::string filename,nlohmann::json param){
 	// name the file with the parameter specifications
-	filename.append("ABCchain_CL");
+	filename.append("MCMCchain_CL");
 	filename.append(itos(param["chain_length"]));
 	filename.append("_seed");
 	filename.append(itos(param["seed"]));
-	filename.append(".txt");
-	return(filename);
+filename.append(".txt");
+return(filename);
 }
 
 
@@ -607,7 +607,8 @@ struct locat_point {
 };
 
 struct cleaner_point {
-	double rel_abund_clean, rel_abund_visitors, rel_abund_resid, prob_Vis_Leav;
+	double abund_clean, abund_visitors, abund_resid, prob_Vis_Leav;
+	double rel_abund_clean, rel_abund_visitors, rel_abund_resid;
 	string site_year;
 	string cleanerID;
 	int countVisitor, totalMarket;
@@ -617,12 +618,13 @@ struct cleaner_point {
 
 struct model_param {
 	//model_param(model_param const &obj);
-	double alphaC, alphaA, gamma, negReward;
+	double alphaC, alphaA, gamma, negReward, scaleConst;
 	model_param &operator= (model_param const &rhs) {
 		alphaA = rhs.alphaA;
 		alphaC = rhs.alphaC;
 		gamma = rhs.gamma;
 		negReward = rhs.negReward;
+		scaleConst = rhs.scaleConst;
 		return *this;
 	}
 };
@@ -642,11 +644,11 @@ vector <locat_point> read_locData(ifstream& marketData) {
 		exit(EXIT_FAILURE);
 	}
 	string header;
-	getline(marketData,header); // skip header
-	vector<locat_point> data_set; 
+	getline(marketData, header); // skip header
+	vector<locat_point> data_set;
 	locat_point input;
 	for (;;) { // read data
-	    // if end of file
+		// if end of file
 		marketData >> input.site_year;
 		marketData >> input.rel_abund_clean;
 		marketData >> input.rel_abund_visitors;
@@ -676,13 +678,13 @@ vector <cleaner_point> read_CleanData(ifstream& marketData) {
 		marketData >> input.site_year;
[…]

### 2021-10-21 analysisABC.R

@@ -14,12 +14,13 @@ source(here("../R_files/posPlots.R"))
 
 simdir<-"D:/Neuch_simulations/ABCfit/Simulations/"
 
-scen<-"ABCclean_gam_Nrew_exp_sca30_"
+scen<-"ABCclean_gam_Nrew_sca_"
 
+## Load files --------------------------------------------------------------
 
 # (listFiles<-list.files(paste0(simdir,scen),full.names = TRUE))
 (listFiles<-list.files(here("Simulations",scen)))
-ABCruns<-grep("chain",listFiles,value = TRUE)
+(ABCruns<-grep("MCMCchain_CL75",listFiles,value = TRUE))
 
 
 # all in a single object
@@ -31,22 +32,22 @@ ABCraw<-do.call(rbind,lapply(ABCruns, function(file){
 }))
 
 # One per onject
-ABCraw<-fread(here("Simulations",scen,ABCruns[1]))
+ABCraw<-fread(here("Simulations",scen,ABCruns[3]))
 
 # All in a list
 mcmcList<-mcmc.list(do.call(list,lapply(ABCruns, function(file){
   rundata<-fread(here("Simulations",scen,file))
-  mcmcRun<-mcmc(rundata[,.(gamma,negReward)])
+  mcmcRun<-mcmc(rundata[,.(gamma,negReward,scaleConst)],thin = 100)#
   return(mcmcRun)
 })))
 
 
-
+## Base R plots -----------------------------------------------------------------
 
 head(ABCraw)
 
 burn.in<-1000
-thinning<-10
+thinning<-100
 
 # filter
 
@@ -55,46 +56,128 @@ ABCburned<-ABCraw[iteration>burn.in]
 ABCfiltered<-ABCraw[iteration>burn.in & iteration %% thinning==0]
 
 ABCburned[seed==1,]
-ABC.gamma.time.short<-dcast(ABCburned,iteration~seed,value.var = c("gamma"))
-ABC.nRew.time.short<-dcast(ABCburned,iteration~seed,value.var = "negReward")
-ABC.logLike.time.short<-dcast(ABCburned,iteration~seed,value.var = "fit")
+ABC.gamma.time.short<-dcast(ABCfiltered,iteration~seed,value.var = c("gamma"))
[…]

### 2021-10-21 exm_bayesTools.R

@@ -1,0 +1,143 @@
+## This example shows how to run and calibrate the VSEM model 
+
+library(BayesianTools) # install.packages("BayesianTools")
+
+# Create input data for the model
+PAR <- fread(here("Data","data_ABC_cleaner_absolute.txt"))
+plot(PAR, main = "PAR (driving the model)", xlab = "Day")
+
+# PAR <- VSEMcreatePAR(1:1000)
+# plot(PAR, main = "PAR (driving the model)", xlab = "Day")
+
+# load reference parameter definition (upper, lower prior)
+refPars <- VSEMgetDefaults()
+# this adds one additional parameter for the likelihood standard deviation (see below)
+refPars[12,] <- c(2, 0.1, 4) 
+rownames(refPars)[12] <- "error-sd"
+head(refPars)
+
+# create some simulated test data 
+# generally recommended to start with simulated data before moving to real data
+referenceData <- VSEM(refPars$best[1:11], PAR) # model predictions with reference parameters  
+referenceData[,1] = 1000 * referenceData[,1] 
+# this adds the error - needs to conform to the error definition in the likelihood
+obs <- referenceData + rnorm(length(referenceData), sd = refPars$best[12])
+oldpar <- par(mfrow = c(2,2))
+for (i in 1:4) plotTimeSeries(observed = obs[,i], 
+                              predicted = referenceData[,i], main = colnames(referenceData)[i])
+
+# Best to program in a way that we can choose easily which parameters to calibrate
+parSel = c(1:6, 12)
+
+# here is the likelihood 
+likelihood <- function(par, sum = TRUE){
+  # set parameters that are not calibrated on default values 
+  x = refPars$best
+  x[parSel] = par
+  predicted <- VSEM(x[1:11], PAR) # replace here VSEM with your model 
+  predicted[,1] = 1000 * predicted[,1] # this is just rescaling
+  diff <- c(predicted[,1:4] - obs[,1:4]) # difference betweeno observed and predicted
+  # univariate normal likelihood. Note that there is a parameter involved here that is fit
+  llValues <- dnorm(diff, sd = x[12], log = TRUE)  
+  if (sum == FALSE) return(llValues)
+  else return(sum(llValues))
+}
+
+# optional, you can also directly provide lower, upper in the createBayesianSetup, see help
+prior <- createUniformPrior(lower = refPars$lower[parSel], 
[…]

### 2021-10-21 loadFieldData.R

@@ -7,21 +7,118 @@ library("readxl")
 
 fieldData<-fread(here("Data","market_model.csv"))
 
-str(fieldData)
 
+str(fieldData)
+## Summarize ecological data by site
+
+fieldData[,unique(site_year)]
+names(fieldData)
+
+fieldData.site<-fieldData[,lapply(.SD,max),by=site_year,
+                          .SDcols=c("abundance_large_100m2","abundance_small_100m2",
+                                   "abundance_cleaners_100m2","percentage_swim_off")]
+
+fieldData.cleaner<-read_xlsx(here("Data","market_raw_data.xlsx"),sheet = "round_data")
+
+fieldData.cleaner<-data.table(fieldData.cleaner)
+
+str(fieldData.cleaner)
+
+fieldData.cleaner[,use_sims:=ifelse(is.na(use_sims),0,1)]
+fieldData.cleaner[,use_sims:=as.logical(use_sims)]
+
+fieldData.filt<-fieldData.cleaner[use_sims==T,]
+fieldData.filt[,score_visitor:=as.numeric(score_visitor)]
+fieldData.filt[,stage:=as.factor(stage)]
+
+str(fieldData.filt)
+
+fieldData.sum<-fieldData.filt[,sum(score_visitor),
+                              by=.(site_year,cleaner_ID)]
+
+str(fieldData.sum)
+
+names(fieldData.sum)[3]<-"score_visitor"
+names(fieldData.sum)[1]<-"site.year"
+
+fieldData.sum[,unique(site.year)]
+fieldData.site$site_year
+
+fieldData.sum[site.year=="NHS2017",site.year:="NHS 2017"]
+
+fieldData.sum[,abund.cleaners:=
+                fieldData.site[match(site.year,site_year),abundance_cleaners_100m2]]
+
+fieldData.sum[order(abund.cleaners),max(abund.cleaners),by=site.year]$site.year==
+  
+  fieldData.site[order(abundance_cleaners_100m2),.(site_year,abundance_cleaners_100m2)]$site_year
+
+fieldData.sum[,abund.visitors:=
+                fieldData.site[match(site.year,site_year),abundance_large_100m2]]
+
+fieldData.sum[,abund.residents:=
+                fieldData.site[match(site.year,site_year),abundance_small_100m2]]
+
+fieldData.sum[,prob.Vis.Leav:=
+                fieldData.site[match(site.year,site_year),percentage_swim_off*0.01]]
+
[…]

### 2021-10-21 manFigures.R

@@ -13,7 +13,7 @@ source(here("..","R_files",'ternaryAEQP.R'))
 source(here("../R_files/posPlots.R"))
 
 
-scenario<-"ABCclean_gam_Nrew_unif_sca20"
+scenario<-"ABCclean_gam_Nrew_sca"
 
 
 predfile<-grep("round",list.files(here("Simulations",paste0(scenario,"_"))),value = T)
@@ -24,11 +24,11 @@ predictData<-fread(here("Simulations",paste0(scenario,"_"),
 str(predictData)
 
 fieldatabyLoc<-predictData[,.(probVisi.data=mean(visitorChoices)/20,
-                              probvisitor.pred=max(),
+                              probvisitor.pred=max(visitorChoices_pred),
                               re.abund.clean=max(rel.abund.cleaners),
                             prob.Vis.leave=max(prob.Vis.Leav)),by=site_year]
 
-
+fieldData
 
 ## Obv vs predicted two panels with point size ---------------------------------
 
@@ -92,7 +92,7 @@ ggplot(data = predictData, aes(x=market_binomial_pred,y=market_binomial_data))+
 
 
 ggplot(data = fieldatabyLoc, aes(x=probvisitor.pred,y=probVisi.data))+
-  geom_point()+theme_classic()+geom_abline(slope = 1)
+    geom_point()+theme_classic()+geom_abline(slope = 1)
 
 
 ## Obv vs predicted contour ----------------------------------------------------
@@ -100,13 +100,18 @@ ggplot(data = fieldatabyLoc, aes(x=probvisitor.pred,y=probVisi.data))+
 
 simsDir <- here("Simulations",paste0(scenario,"_"))
 
-simsDir <- paste0("e:/NeuchSims/AC/",paste0(scenario,"_"))
+# simsDir <- paste0("e:/NeuchSims/AC/",paste0(scenario,"_"))
 
 list.files(simsDir,recursive = TRUE,full.names = TRUE)
+
+(listPar<-c("gamma","neta"))
+
+(listVal<-c(0.2624,0))
 
 FIAlastQuarData<-do.call(rbind,lapply(
   getFilelist(simsDir,fullNam = TRUE)$p1,file2lastProp,0.70,outPar="Vlp",
   full.path=TRUE))
+
 
 range(fieldatabyLoc$prob.Vis.leave)
 range(FIAlastQuarData$Vlp)
[…]

### 2021-10-21 parTOjson.R

@@ -10,30 +10,33 @@ exedir<-'/./ActCrit.exe'
 
 fileName<-"parameters.json"
 
-scenario<-"ABCclean_gam_Nrew_unif_sca30"
+scenario<-"ABCclean_gam_Nrew_sca"
 
-# scenario<-"predictions"
+
 
 # For predictions on the observed values ---------------------------------------
 
-scenario<-"ABCclean_gam_Nrew_exp_sca30"
+scenario<-"MCMCfakedata"
 
 param_pred<-list(totRounds=10000,ResReward=1,VisReward=1,
             ResProbLeav=0,scenario=0, inbr=0,outbr=0,forRat=0.0,
-            seed=1, propfullPrint = 0.7,sdPert=c(0.05,0.05,0.1,0.05),
+            seed=1, propfullPrint = 0.7,sdPert=c(0.05,0.05,0.1,0.05,1),
             chain_length=0,
-            init=c(0.05,0.05,0,0),# alphaA,AlphaC, Gamma, NegRew
-            pertScen = 1, 	# enum perturnScen {all,  bothFut, justGam, justNegRew};
+            init=c(0.05,0.05,0,0,35),# alphaA,AlphaC, Gamma, NegRew, scalConst
+            pertScen = c(FALSE,FALSE,FALSE,FALSE,TRUE), 	
             MCMC =0, data="clean",nRep=30,
-            dataFile = here("Data","data_ABC_cleaner_30.txt"),
+            dataFile = here("Data","data_ABC_cleaner_fakePartial.txt"),
             folderL=paste0(here(simsDir),"/",scenario,"_/"))
 
 
-
+check_create.dir(here(simsDir),param = rep(scenario,1),values = c(""))
 
 param_pred$folder<-param_pred$folderL
-param_pred$init[c(3,4)]<-sumMCMClist$statistics[,1]
-fileName<-paste("parameters_pred",param_pred$seed,".json",sep="")
+param_pred$init[c(3,4,5)]<-c(runif(1,max = 0.6),runif(1,max = 1),
+                             runif(1,max = 500,min = 1))
+  #c(modeGam,modenegReward,modeScal)
+  #sumMCMC$statistics[,1]
+fileName<-paste("parameters_pred_",param_pred$seed,".json",sep="")
 outParam.pred<-toJSON(param_pred,auto_unbox = TRUE,pretty = TRUE)
[…]

### 2021-10-24 analysisABC.R
Further analysis

@@ -14,13 +14,13 @@ source(here("../R_files/posPlots.R"))
 
 simdir<-"D:/Neuch_simulations/ABCfit/Simulations/"
 
-scen<-"ABCclean_gam_Nrew_sca_"
+scen<-"MCMCfakedata_"
 
 ## Load files --------------------------------------------------------------
 
 # (listFiles<-list.files(paste0(simdir,scen),full.names = TRUE))
 (listFiles<-list.files(here("Simulations",scen)))
-(ABCruns<-grep("MCMCchain_CL75",listFiles,value = TRUE))
+(ABCruns<-grep("MCMCchain",listFiles,value = TRUE))
 
 
 # all in a single object
@@ -67,7 +67,7 @@ matplot(y=ABC.gamma.time.short[,c(2,3,4)],
         x=ABC.gamma.time.short[,iteration],
         type="l",lty=1,col=2:5,ylab = "",
         xlab="",xaxt="n",lwd=0.1)
-quantile()
+
 mtext(text =  expression(gamma),cex=3,side = 2,line = 3)
 
 par(plt=posPlot(numploty = 4,idploty = 4,numplotx = 2,idplotx = 2)-c(0.05,0.05,0,0),
@@ -132,7 +132,7 @@ legend("right",legend = c("mode","mean","median"),col = c("black","red","green")
        lty=1,lwd=2,cex=0.8)
 
 
-par(plt=posPlot(numploty = 4,idploty = 1)-c(0.05,0.05,0.05,0.05),mfrow=c(1,1),las=1,new=TRUE)
+par(plt=posPlot(numploty = 4,idploty = 1,)-c(0.05,0.05,0.05,0.05),mfrow=c(1,1),las=1,new=TRUE)
 matplot(y=ABC.logLike.time.short[,c(2,3,4)],
         x=ABC.logLike.time.short[,iteration],
         type="l",lty=1,col=2:5,ylab = "",
@@ -141,30 +141,78 @@ axis(1)
 mtext(text = "loglikelihood" ,line = 3,cex = 1,side = 2,las=0)
 mtext(text = "iteration" ,line = 2,cex = 1,side = 1,las=0)
 
+# Base R for a single chain ----------------------------------------------------
 
-par(plt=posPlot(numploty = 4,idploty = 4),mfrow=c(1,1),las=1)
-matplot(y=ABCburned[,gamma],
[…]

### 2021-10-24 manFigures.R

@@ -13,7 +13,7 @@ source(here("..","R_files",'ternaryAEQP.R'))
 source(here("../R_files/posPlots.R"))
 
 
-scenario<-"ABCclean_gam_Nrew_sca"
+scenario<-"MCMCfakedata"
 
 
 predfile<-grep("round",list.files(here("Simulations",paste0(scenario,"_"))),value = T)
@@ -174,8 +174,10 @@ png(here("Simulations",paste0(scenario,"_"),
          paste0(strsplit(predfile,"seed")[[1]][1],"contour_ggplot.png")),width = 1300,height = 700)
 
 
+fieldatabyLoc[,resids:=probVisi.data-probvisitor.pred]
+
 rsqr<-1-(fieldatabyLoc[,sum(resids^2)])/
-  fieldatabyLoc[,sum((market_binomial_data-mean(market_binomial_data))^2)]
+  fieldatabyLoc[,sum((probVisi.data-mean(probVisi.data))^2)]
 
 
 cont.obs.pred<- ggplot(data = FIAinterpData,aes(x=rel.abund.cleaners,y=prob.Vis.Leav,

### 2021-10-24 parTOjson.R

@@ -25,16 +25,16 @@ param_pred<-list(totRounds=10000,ResReward=1,VisReward=1,
             init=c(0.05,0.05,0,0,35),# alphaA,AlphaC, Gamma, NegRew, scalConst
             pertScen = c(FALSE,FALSE,FALSE,FALSE,TRUE), 	
             MCMC =0, data="clean",nRep=30,
-            dataFile = here("Data","data_ABC_cleaner_fakePartial.txt"),
+            dataFile = here("Data","data_ABC_cleaner_absolute.txt"),
             folderL=paste0(here(simsDir),"/",scenario,"_/"))
 
 
-check_create.dir(here(simsDir),param = rep(scenario,1),values = c(""))
+# check_create.dir(here(simsDir),param = rep(scenario,1),values = c(""))
 
 param_pred$folder<-param_pred$folderL
-param_pred$init[c(3,4,5)]<-c(runif(1,max = 0.6),runif(1,max = 1),
+param_pred$init[c(3,4,5)]<- #c(modeGam,modenegReward,modeScal)
+  c(runif(1,max = 1),runif(1,max = 1),
                              runif(1,max = 500,min = 1))
-  #c(modeGam,modenegReward,modeScal)
   #sumMCMC$statistics[,1]
 fileName<-paste("parameters_pred_",param_pred$seed,".json",sep="")
 outParam.pred<-toJSON(param_pred,auto_unbox = TRUE,pretty = TRUE)
@@ -149,7 +149,7 @@ for (k in 1:90) print(y)
 
 # MCMC fit - Generate json parameter files for -------------------------------------
 
-scenario<-"MCMCfakedata"
+scenario<-scen<-"ABCclean_gam_Nrew_sca_2"
 
 fake.data.pred<-fread(here("Simulations","MCMCfakedata_",
                            "round_gamma_0.2regRew_0.05scaC_359.96alphA_0.05alphC_0.05_seed1.txt "))

### 2021-10-29 manuscript_1.0.Rmd
More or  less done with the methods. Trying out to adapt the model for using BayesTools

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 library(here)
@@
 We used the data from @triki_Biological_2019 to fit the parameters of the  model analyzed in @quinones_Reinforcement_2019.
-The data consist a of the number of times that set of cleaners chose to
-feed of a client plate in the context of the market task.
-This set of cleaners were collected under different locations and in each one of those locations environmental variables were measured.
+The data consist a of the number of times that a set of cleaners chose to
+feed of a visitor client plate in the context of the market task.
+This set of cleaners were collected in different locations and in each one of those locations environmental variables were measured.
 These environmental variables recorded from each location are used as input for the computational model.
@@
 As $\gamma$ increases individuals include more the reward obtained from the subsequent choices.
-That amounts to calculating and using for deision making future expected rewards.
+That amounts to calculating and using for decision making the future expected rewards.
 Second, $\eta$ measures how much individuals include in their reward the fleeing behaviour of visitor clients as a negative component.
-Both of these components allow individuals to use in ther estimates the future effects of their choices.
+Both of these parameters allow individuals to use in ther estimates the future effects of their choices.
 The model analysis shows that without either of them, learning individuals cannot develop a preference for the visitor client.
-The analysis contrast the effect of these two parameters agaisnt variation in the relative abundance of the two client types.
+The analysis contrast the effect of these two parameters against variation in the relative abundance of the two client types.
 The analysis shows that under very high client abundance the presence of  future rewards ($\gamma>0$) no longer favours the preference for the ephemeral clients, this effect does not occur with negative reward mediated by client behaviour ($\eta$).
@@
 How often the cleaning stations are empty in turns influences the value that cleaners can expect from their choices.
+The more empty the empty cleaning stations are, the more it is to choose a client that will not stay.
 However, the frequency of empty spots in the cleaning station does not only depend on cleaner relative abundances but also on the frequency with which clients visit the station.
 The client visitation frequency will likely vary among different clients depnding on parasite abundances and loads.
-There are variables for which we do not have field estimates.
-Thus, we use
-computed a measure of relative cleaner abundance for each site relative to absolute abundances and multiplied by scalling constant that changes the range of the variable.
+These are variables for which we do not have field estimates.
+Thus, we
+computed a measure of relative cleaner abundance for each site relative to absolute abundances and multiplied by a scalling constant that changes the range of the variable.
 We fit the value os the scalling constant as part of the statistical inference.
@@
 
-```{r, fig1}
-knitr::include_graphics(here("Simulations","ABCclean_gam_nRew_sca_",                            "round_gamma_0.3regRew_0.11scaC_133.5alphA_0.05alphC_0.05_contour_ggplot.png"))
+Posterior distribution samples taken from the MCMC (Fig.\ref{fig:estimValues}) show data from the market taks experiments fit best with a learning model that includes the use of both future reward ($\gamma>0$) and negative reward triggered by visitor clients leaving the station ($\eta>0$).
+The value of negative reward is, however, small.
+
+```{r estimValues,fig.cap="Estimates posterior distributions for parameter values of $gamma$, $eta$ and the scalling constant.
+We show the kernet density estimates, the mode (black line) and the 95% credible interval for the three paramteres"}
+scenario<-"ABCclean_gam_nRew_sca_2_"
+file<-list.files(here("Simulations",scenario),full.names = TRUE,
+                      pattern = "MCMC.png")
+knitr::include_graphics(file)
+```
+
+```{r predictions, fig.cap="Observed and predicted probability of choosing a visitor.
+Left-hand side pane: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis.
+Dots show the frequency of visitor choices for the 12 locations, as well as the corresponding relative cleaner abundance (x axis) and frequency of visitors leaving the station (y axis).
+Right-hand side panel: scatter plot of the observed and predicted probabilities of choosing a visitor for 12 locations.
+Line correspond to a perfect match between observed and predicted probabilities."}
+scenario<-"ABCclean_gam_nRew_sca_2_"
+file<-list.files(here("Simulations",scenario),full.names = TRUE,
+                      pattern = ".png")
+knitr::include_graphics(file[3])
 ```

### 2021-10-29 analysisABC.R

@@ -8,13 +8,15 @@ library(plot3D)
 require(data.table)
 library(coda)
 library(mcmcplots)
+library(ggplot2)
 library(bayesplot) 
 library(lattice)
 source(here("../R_files/posPlots.R"))
+library("jsonlite")
 
 simdir<-"D:/Neuch_simulations/ABCfit/Simulations/"
 
-scen<-"MCMCfakedata_"
+scen<-"ABCclean_gam_Nrew_sca_2_"
 
 ## Load files --------------------------------------------------------------
 
@@ -49,13 +51,20 @@ head(ABCraw)
 burn.in<-1000
 thinning<-100
 
+## get original parameter values from the simulations
+
+parsOrigin<-fromJSON(here("Simulations",scen,
+              grep("parameters_pred",listFiles,value = TRUE)))
+
+
+
 # filter
 
 ABCburned<-ABCraw[iteration>burn.in]
 
 ABCfiltered<-ABCraw[iteration>burn.in & iteration %% thinning==0]
 
-ABCburned[seed==1,]
+
 ABC.gamma.time.short<-dcast(ABCfiltered,iteration~seed,value.var = c("gamma"))
 ABC.nRew.time.short<-dcast(ABCfiltered,iteration~seed,value.var = "negReward")
 ABC.sca.time.short<-dcast(ABCfiltered,iteration~seed,value.var = "scaleConst")
@@ -83,6 +92,8 @@ lines(x = rep(meanGam,2),y = range(densGamma$y),col="red")
 medianGam<-median(ABCfiltered$gamma)
 lines(x = rep(medianGam,2),y = range(densGamma$y),col="green")
 
+lines(x = rep(parsOrigin$init[3],2),y = range(densGamma$y),col="blue")
+
 par(plt=posPlot(numploty = 4,idploty = 3,numplotx = 2,idplotx = 1)-c(0.05,0.05,0,0),
     mfrow=c(1,1),las=1,new=TRUE)
 matplot(y=ABC.nRew.time.short[,c(2,3,4)],
@@ -105,6 +116,7 @@ lines(x = rep(meannegReward,2),y = range(densnegReward$y),col="red")
 mediannegReward<-median(ABCfiltered$negReward)
 lines(x = rep(mediannegReward,2),y = range(densnegReward$y),col="green")
 
+lines(x = rep(parsOrigin$init[4],2),y = range(densGamma$y),col="blue")
 
[…]

### 2021-10-29 exm_bayesTools.R

@@ -3,18 +3,25 @@
 library(BayesianTools) # install.packages("BayesianTools")
 
 # Create input data for the model
-PAR <- fread(here("Data","data_ABC_cleaner_absolute.txt"))
-plot(PAR, main = "PAR (driving the model)", xlab = "Day")
-
+fieldData <- fread(here("Data","data_ABC_cleaner_absolute.txt"))
+par(plt=posPlot())
+plot(data=fieldData,prob.Vis.Leav~abund.cleaners,type="p",cex=5*score_visitor/20,
+     xlab = "Cleaner abundance",ylab="Visitor leaving prob.",pch=20)
 # PAR <- VSEMcreatePAR(1:1000)
 # plot(PAR, main = "PAR (driving the model)", xlab = "Day")
 
 # load reference parameter definition (upper, lower prior)
 refPars <- VSEMgetDefaults()
+str(refPars)
+refPars <- data.frame(best=c(0,0,30),
+                lower=c(0,0,0),
+                upper=c(1,10,1000))
+row.names(refPars)<-c("gamma","eta","sca")
+
 # this adds one additional parameter for the likelihood standard deviation (see below)
-refPars[12,] <- c(2, 0.1, 4) 
-rownames(refPars)[12] <- "error-sd"
-head(refPars)
+# refPars[12,] <- c(2, 0.1, 4) 
+# rownames(refPars)[12] <- "error-sd"
+# head(refPars)
 
 # create some simulated test data 
 # generally recommended to start with simulated data before moving to real data

### 2021-10-29 parTOjson.R

@@ -10,13 +10,13 @@ exedir<-'/./ActCrit.exe'
 
 fileName<-"parameters.json"
 
-scenario<-"ABCclean_gam_Nrew_sca"
+scenario<-"ABCclean_gam_sca"
 
 
 
 # For predictions on the observed values ---------------------------------------
 
-scenario<-"MCMCfakedata"
+scenario<-"MCMCfakedata2"
 
 param_pred<-list(totRounds=10000,ResReward=1,VisReward=1,
             ResProbLeav=0,scenario=0, inbr=0,outbr=0,forRat=0.0,
@@ -29,7 +29,7 @@ param_pred<-list(totRounds=10000,ResReward=1,VisReward=1,
             folderL=paste0(here(simsDir),"/",scenario,"_/"))
 
 
-# check_create.dir(here(simsDir),param = rep(scenario,1),values = c(""))
+check_create.dir(here(simsDir),param = rep(scenario,1),values = c(""))
 
 param_pred$folder<-param_pred$folderL
 param_pred$init[c(3,4,5)]<- #c(modeGam,modenegReward,modeScal)
@@ -149,10 +149,11 @@ for (k in 1:90) print(y)
 
 # MCMC fit - Generate json parameter files for -------------------------------------
 
-scenario<-scen<-"ABCclean_gam_Nrew_sca_2"
+scenario<-"MCMCfakedata2_"
 
-fake.data.pred<-fread(here("Simulations","MCMCfakedata_",
-                           "round_gamma_0.2regRew_0.05scaC_359.96alphA_0.05alphC_0.05_seed1.txt "))
+filesList<-list.files(here("Simulations",scenario),full.names = TRUE)
+
+fake.data.pred<-fread(grep("round",filesList,value = TRUE))
 
 fake.data.pred<-fake.data.pred[,.(site_year,CleanerID,
                                   score_visitor=sapply(visitorChoices_pred,
@@ -165,21 +166,23 @@ fake.data.pred<-fake.data.pred[fakeData,on=.(site_year=site.year,CleanerID=clean
 fake.data.pred<-fake.data.pred[,.(site_year,CleanerID,score_visitor,abund.cleaners,
                   abund.visitors,abund.residents,prob.Vis.Leav),]
 
-fwrite(fake.data.pred,file = here("Simulations","MCMCfakedata_","data_MCMC_fake.txt"),
+fwrite(fake.data.pred,file = here("Simulations",scenario,"data_MCMC_fake.txt"),
[…]

### 2021-11-04 manuscript_1.0.Rmd
New figures including three different models. Reuslts text if finished.

@@
 author:
-- Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento
-  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia - andresqp@gmail.com]
+- Andrés E. Quiñones^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland - andresqp@gmail.com]
 - Zegni Triki^[Department of Zoology, Stockholm University, Stockholm, Sweden]
@@
 output: 
+  word_document: default
   pdf_document: default
-  word_document: default
 fig_caption: yes
@@
 - \linenumbers
+- \newcommand{\beginsupplement}{
+    \setcounter{table}{0}  
+    \renewcommand{\thetable}{S\arabic{table}} 
+    \setcounter{figure}{0} 
+    \renewcommand{\thefigure}{S\arabic{figure}}
 keep_tex: yes
@@
 spacing: double
-abstract: My abstract.
+abstract: Insert a nice abstract here.
 ---
@@
 parasite removal [@begon_Ecology_2006]. However, when cooperation is costly, 
-it is not straigtforward to understand why individuals would be willing to pay that cost [@lehmann_Evolution_2006]. Furthermore, cooperative interactions might open the way for 
+it is not straightforward to understand why individuals would be willing to pay that cost [@lehmann_Evolution_2006]. Furthermore, cooperative interactions might open the way for 
 one partner to exploit the other, what prevents this type of exploitation? [@herre_Evolution_1999]
@@
 adjust their strategy to the prevalent market conditions they find in their environment. 
-This adjustement cannot be due to local genetic adaptation, given that cleaners are open water spawners 
-and the environmental changes in different sites have occured in the last decade [@triki_Biological_2019; @triki_Decrease_2018]. Alternatively, a recent cognitive model proposes that differences in 
+This adjustment cannot be due to local genetic adaptation, given that cleaners are open water spawners 
+and the environmental changes in different sites have occurred in the last decade [@triki_Biological_2019; @triki_Decrease_2018]. Alternatively, a recent cognitive model proposes that differences in 
 cleaners preferences can arise trough the interaction of cleaner's cognitive processes 
 with the environmental conditions [@quinones_Reinforcement_2019]. Under this view,
-cleaners ontogenatically develop their preferences for different clients through learning,
+cleaners ontogenetically develop their preferences for different clients through learning,
 and some environmental conditions facilitate learning. Thus, the adaptive adjustment 
@@
 a Metropolis sampler to estimate the posterior distribution of the model parameters.
-We used the data from @triki_Biological_2019 to fit the parameters of the  model analyzed in @quinones_Reinforcement_2019.
+We used the data from @triki_Biological_2019 to fit the parameters of the  model analysed in @quinones_Reinforcement_2019.
 The data consist a of the number of times that a set of cleaners chose to
@@
 Given these inputs, the computational model produces a time series of choices in the market task.
-In the model the preference for the two option changes as the algorithm learns based on the reward that the two options provide.
-Thus, we run the model for 10000 steps and use the las 3000 steps as the preference developed in the model for visitors given the environmental variables.
+In the model the preference for the two options changes as the algorithm learns based on the reward that the two options provide.
+Thus, we run the model for 10000 steps and use the last 3000 steps as the preference developed in the model for visitors given the environmental variables.
 We use this preference as the predicted probability of choosing a visitor in the data, and use it to calculate the likelihood that the number of visitor choices obtained from each cleaner is generated by the model.
-We sum over all the cleaners the loglikehoods and use the total for the acceptance of new parameters in the MCMC.
+We sum over all the cleaners the log-likehoods and use the total for the acceptance of new parameters in the MCMC.
 
-The model consist of a set individual-simulations where individuals face a series of choices between two options.
+The model consist of a set individual-based simulations where individuals face a series of choices between two options.
 Individuals in the model represent cleaners, and the options they face simulate the natural conditions of a cleaning market.
@@
 Individuals obtain reward from cleaning a client regardless of the type.
-Everytime individuals face and make a choice they update the probability of making that same choice.
-The update is based on the difference between the expectation of reward and the actual reward the individual obtains (prediction error); and it is carried in the direction that would lead to more reward given the new information.
+Every time individuals face and make a choice they update the probability of making that same choice.
+The update is based on the difference between the expectation of reward and the actual reward the individual obtains (prediction error); and it is carried in the direction that would lead to more reward being obtained, given the new information.
 Further details of the model implementation can be found in the paper [@quinones_Reinforcement_2019].
 The model analysis points to the relevance of two parameters.
-First, $\gamma$ measures how much individuals include future rewards in their desicion updates.
+First, $\gamma$ measures how much individuals include future rewards in their decision updates.
 If $\gamma=0$, individuals only use the immediate reward obtained from a cleaning interaction.
@@
 Second, $\eta$ measures how much individuals include in their reward the fleeing behaviour of visitor clients as a negative component.
-Both of these parameters allow individuals to use in ther estimates the future effects of their choices.
+Both of these parameters allow individuals to use in their estimates the future effects of their choices.
 The model analysis shows that without either of them, learning individuals cannot develop a preference for the visitor client.
@@
 However, the frequency of empty spots in the cleaning station does not only depend on cleaner relative abundances but also on the frequency with which clients visit the station.
-The client visitation frequency will likely vary among different clients depnding on parasite abundances and loads.
+The client visitation frequency will likely vary among different clients depending on parasite abundances and loads.
 These are variables for which we do not have field estimates.
 Thus, we
-computed a measure of relative cleaner abundance for each site relative to absolute abundances and multiplied by a scalling constant that changes the range of the variable.
-We fit the value os the scalling constant as part of the statistical inference.
-As for the client abundances for residents and visitors we computed a relative measure with respect to the total client abundance, and weigthted that by the rescaled relative absence of cleaners (1-relative cleaner abundance).
+computed a measure of relative cleaner abundance for each site relative to absolute abundances and multiplied by a scaling constant that changes the range of the variable.
+We fit the value of the scaling constant as part of the statistical inference.
+As for the client abundances for residents and visitors we computed a relative measure with respect to the total client abundance, and weighted that by the rescaled relative absence of cleaners (1-relative cleaner abundance).
 Thus, all three measures of relative abundance sum up to 1, and can be used in the model as a proxy for the probability of having different options in the cleaning station.
 
-### Results
+## Results
 
-Posterior distribution samples taken from the MCMC (Fig.\ref{fig:estimValues}) show data from the market taks experiments fit best with a learning model that includes the use of both future reward ($\gamma>0$) and negative reward triggered by visitor clients leaving the station ($\eta>0$).
-The value of negative reward is, however, small.
+```{r post, fig.cap="Posterior distributions for parameter values $\\gamma$, $\\eta$ and the scalling constant.
+We show the kerne density estimates, and bellow the mean (black point) and the 65\\% and 95\\%  highest posterior density interval for the three paramteres.
+On the left (panels a, c and c) posterior distributions for a model with all three parameters; while on the right (b and d) a model without negative reward, ",out.width='85%'}
+knitr::include_graphics(here("post_both_gamma.png"))
 
-```{r estimValues,fig.cap="Estimates posterior distributions for parameter values of $gamma$, $eta$ and the scalling constant.
-We show the kernet density estimates, the mode (black line) and the 95% credible interval for the three paramteres"}
-scenario<-"ABCclean_gam_nRew_sca_2_"
-file<-list.files(here("Simulations",scenario),full.names = TRUE,
-                      pattern = "MCMC.png")
-knitr::include_graphics(file)
 ```
 
-```{r predictions, fig.cap="Observed and predicted probability of choosing a visitor.
-Left-hand side pane: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis.
+Figure \ref{fig:post} panels a,c and e show the posterior distributions of the parameter values 
+of the full model, which includes both future and negative reward. Notably, the bulk of the posterior distribution 
+for the parameter tuning negative reward ($\eta$) is close to zero.
+Given the structure of the model, we prevented $\eta$ from taking negative values.
+Nevertheless, the shape of the posterior and its credible intervals suggest the absence of negative reward to be well supported by the data.
+Thus, we run the analysis setting $\eta$ to zero.
+Figure \ref{fig:post} panels b and d, show the posterior distributions of the two parameters estimated.
+Not surprisingly, under a model without negative reward, the distribution of $\gamma$ shifts to higher values.
+
+```{r pred, fig.cap="Observed and predicted probability of choosing a visitor.
+Left-hand side panel: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis.
+Dots show the frequency of visitor choices for the 12 locations, as well as the corresponding relative cleaner abundance (x axis) and frequency of visitors leaving the station (y axis).
+Right-hand side panels: scatter plot of the observed and predicted probabilities of choosing a visitor for 12 locations.
+Line correspond to a perfect match between observed and predicted probabilities.
+Upper panels show prediction from a model including future and negative reward, lower panels froma  model with only future reward.",out.width='90%'}
+
+knitr::include_graphics(here("contour_both_gamma.png"))
+
+```
+
+Figure \ref{fig:pred} shows predictions from the two models and their fit to the data. On the left,
+the colour coded contour shows the prediction of the model for combinations of cleaner abundance (x axis) and
+visitor leaving probability (y axis). Predictions were generated by running the model using the mode of the posterior
+distributions as the parameter values, and varying systematically cleaner abundance and visitor leaving probability.
+Colour of the points on top of the contour show the average frequency with 
+which cleaner fish chose the visitor plate in their last 20 experimental trials. Each point corresponds to a 
+location with a combination of cleaner abundance and visitor leaving probability. Panels on the right show how close
+the observed values (y axis) are from the predictions of the models (x axis).
+The line represents the perfect fit between observed and predicted values.
+Predicted values were taken using the mode just as in the left panels.
+On the lower right corner of the panel we show the coefficient of determination.
+Despite having one more parameter varying, the full model gives a lower fit to the data than the model without negative reward, measured by the coefficient of determination.
+In the supplementary material, we also show that a model with only negative reward, excluding future rewards ($\gamma=0$, Fig.
+\ref{fig:nogamma}), provides a lower fit than either of the ones presented in the main text.
+
+The main reason for negative and future reward to give different predictions is the way that cleaner's relative abundance influences the preference for the visitor clients.
+Visitor leaving probability has a similar positive effect on the probability of choosing the visitor clients on all three models
+(Fig.
+\ref{fig:pred}, \ref{fig:nogamma}).
+In contrast, cleaner's relative abundance has a different effect in the model with only future reward, compared to the other two.
+In the model with only future reward
+only intermediate cleaner abundance triggers a preference for the visitor clients 
+(Fig. \ref{fig:pred} c). In models with negative reward, both intermediate and low cleaner abundance
+trigger a preference for the visitor. (Fig. \ref{fig:pred} a, \ref{fig:nogamma})
+
+## Discussion
+
+*Insert a nice discussion here*
+
+## Cited literature
+
+<div id="refs"></div>
+
+\newpage
+
+# Supplementary material
+
+\beginsupplement
+
+```{r nogamma, fig.cap="Observed and predicted probability of choosing a visitor for a model without future rewards.
+Left-hand side panel: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis.
 Dots show the frequency of visitor choices for the 12 locations, as well as the corresponding relative cleaner abundance (x axis) and frequency of visitors leaving the station (y axis).
 Right-hand side panel: scatter plot of the observed and predicted probabilities of choosing a visitor for 12 locations.
-Line correspond to a perfect match between observed and predicted probabilities."}
-scenario<-"ABCclean_gam_nRew_sca_2_"
-file<-list.files(here("Simulations",scenario),full.names = TRUE,
-                      pattern = ".png")
-knitr::include_graphics(file[3])
+Line correspond to a perfect match between observed and predicted probabilities.",out.width='90%'}
+fileNogamma<-grep("contour_gamma_",list.files(here("Simulations","ABCclean_nRew_sca_"),
+                                              full.names = TRUE),value = TRUE)
+knitr::include_graphics(fileNogamma)
+
 ```
-
-## Cited literature

### 2021-11-04 analysisABC.R

@@ -9,36 +9,58 @@ require(data.table)
 library(coda)
 library(mcmcplots)
 library(ggplot2)
+library(cowplot)
+library(ggdist)
 library(bayesplot) 
 library(lattice)
+library(ggridges)
 source(here("../R_files/posPlots.R"))
 library("jsonlite")
 
 simdir<-"D:/Neuch_simulations/ABCfit/Simulations/"
 
-scen<-"ABCclean_gam_Nrew_sca_2_"
+scen1<-"ABCclean_gam_nRew_sca_"
+scen2<-"ABCclean_gam_sca_"
 
 ## Load files --------------------------------------------------------------
 
 # (listFiles<-list.files(paste0(simdir,scen),full.names = TRUE))
-(listFiles<-list.files(here("Simulations",scen)))
-(ABCruns<-grep("MCMCchain",listFiles,value = TRUE))
+(listFiles1<-list.files(here("Simulations",scen1)))
+(ABCruns1<-grep("MCMCchain",listFiles1,value = TRUE))
+
+(listFiles2<-list.files(here("Simulations",scen2)))
+(ABCruns2<-grep("MCMCchain",listFiles2,value = TRUE))
 
 
 # all in a single object
-ABCraw<-do.call(rbind,lapply(ABCruns, function(file){
-  rundata<-fread(here("Simulations",scen,file))
+ABCraw.1<-do.call(rbind,lapply(ABCruns1, function(file){
+  rundata<-fread(here("Simulations",scen1,file))
   seedNum<-strsplit(file,"_")[[1]][3]
   seedNum<-as.numeric(gsub("[[:alpha:]]",seedNum,replacement = ''))
   rundata[,seed:=rep(seedNum,dim(rundata)[1])]
 }))
 
+ABCraw.2<-do.call(rbind,lapply(ABCruns2, function(file){
+  rundata<-fread(here("Simulations",scen2,file))
+  seedNum<-strsplit(file,"_")[[1]][3]
+  seedNum<-as.numeric(gsub("[[:alpha:]]",seedNum,replacement = ''))
+  rundata[,seed:=rep(seedNum,dim(rundata)[1])]
+}))
+
+
 # One per onject
 ABCraw<-fread(here("Simulations",scen,ABCruns[3]))
 
 # All in a list
-mcmcList<-mcmc.list(do.call(list,lapply(ABCruns, function(file){
-  rundata<-fread(here("Simulations",scen,file))
[…]

### 2021-11-04 manFigures.R

@@ -4,6 +4,7 @@ library(data.table)
 library(ggplot2)
 library("RColorBrewer")
 library("patchwork") 
+library(ggpubr)
 library(here)
 source(here("loadData.R"))
 source(here("data2interp.R"))
@@ -13,88 +14,72 @@ source(here("..","R_files",'ternaryAEQP.R'))
 source(here("../R_files/posPlots.R"))
 
 
-scenario<-"MCMCfakedata"
+scenario<-"ABCclean_gam_Nrew_sca"
+scenario2<-"ABCclean_gam_sca"
 
+predfile.both<-grep(".txt",grep("round",list.files(here("Simulations",paste0(scenario,"_"))),value = T),
+                    value = T)
+predfile.gam<-grep(".txt",grep("round",list.files(here("Simulations",paste0(scenario2,"_"))),value = T),
+                   value = T)
 
-predfile<-grep("round",list.files(here("Simulations",paste0(scenario,"_"))),value = T)
-
-predictData<-fread(here("Simulations",paste0(scenario,"_"),
-                      predfile))
+predictData.both<-fread(here("Simulations",paste0(scenario,"_"),
+                      predfile.both))
+predictData.gam<-fread(here("Simulations",paste0(scenario2,"_"),
+                             predfile.gam))
 
 str(predictData)
 
-fieldatabyLoc<-predictData[,.(probVisi.data=mean(visitorChoices)/20,
+fieldatabyLoc.both<-predictData.both[,.(probVisi.data=mean(visitorChoices)/20,
                               probvisitor.pred=max(visitorChoices_pred),
                               re.abund.clean=max(rel.abund.cleaners),
                             prob.Vis.leave=max(prob.Vis.Leav)),by=site_year]
 
-fieldData
+fieldatabyLoc.gam<-predictData.gam[,.(probVisi.data=mean(visitorChoices)/20,
+                                   probvisitor.pred=max(visitorChoices_pred),
+                                   re.abund.clean=max(rel.abund.cleaners),
+                                   prob.Vis.leave=max(prob.Vis.Leav)),by=site_year]
 
[…]

### 2021-11-04 parTOjson.R

@@ -10,13 +10,13 @@ exedir<-'/./ActCrit.exe'
 
 fileName<-"parameters.json"
 
-scenario<-"ABCclean_gam_sca"
+scenario<-"ABCclean_nRew_sca"
 
 
 
 # For predictions on the observed values ---------------------------------------
 
-scenario<-"MCMCfakedata2"
+scenario<-"ABCclean_nRew_sca"
 
 param_pred<-list(totRounds=10000,ResReward=1,VisReward=1,
             ResProbLeav=0,scenario=0, inbr=0,outbr=0,forRat=0.0,
@@ -32,9 +32,9 @@ param_pred<-list(totRounds=10000,ResReward=1,VisReward=1,
 check_create.dir(here(simsDir),param = rep(scenario,1),values = c(""))
 
 param_pred$folder<-param_pred$folderL
-param_pred$init[c(3,4,5)]<- #c(modeGam,modenegReward,modeScal)
-  c(runif(1,max = 1),runif(1,max = 1),
-                             runif(1,max = 500,min = 1))
+param_pred$init[c(3,4,5)]<- c(modeGam,modenegReward,modeScal)
+  # c(runif(1,max = 1),runif(1,max = 1),
+  #                            runif(1,max = 500,min = 1))
   #sumMCMC$statistics[,1]
 fileName<-paste("parameters_pred_",param_pred$seed,".json",sep="")
 outParam.pred<-toJSON(param_pred,auto_unbox = TRUE,pretty = TRUE)
@@ -69,7 +69,7 @@ param<-list(totRounds=10000,ResReward=1,VisReward=1,
             ResProbLeav=0,VisProbLeav=1,negativeRew=-modenegReward,#sumMCMClist$statistics[,1][2],
             scenario=0,
             inbr=0,outbr=0,trainingRep=10,forRat=0.0,
-            alphaT=0.05,printGen=1,seed=1, gammaRange=I(c(modeGam)),#c(0,sumMCMClist$statistics[,1][1])),#,
+            alphaT=0.05,printGen=1,seed=1, gammaRange=I(c(0)),#c(0,sumMCMClist$statistics[,1][1])),#,
             netaRange=I(c(1)),alphaThRange=I(c(0.05)),numlearn=1,
             propfullPrint = 0.7,
             alphaThNch=0.05,
@@ -178,7 +178,7 @@ param_mcmc<-list(totRounds=10000,ResReward=1,VisReward=1,
[…]

### 2021-11-10 manuscript_1.0.Rmd
Erased old simulations which were no longer needed. Added Redouan intro to the manuscript_. New Chains were parameters can be negative.

@@
 output: 
-  word_document: default
   pdf_document: default
@@
     \setcounter{figure}{0} 
-    \renewcommand{\thefigure}{S\arabic{figure}}
+    \renewcommand{\thefigure}{S\arabic{figure}}}
 keep_tex: yes
@@
 bibliography: Cleaner_learning.bib
-style: C:\Users\andre\Zotero\styles\behavioral-ecology-and-sociobiology.csl
+style: M:\Zotero\styles\behavioral-ecology-and-sociobiology.csl
 spacing: double
 abstract: Insert a nice abstract here.
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
@@
 
-Cooperation among individuals of the same or different species is ubiquitous in nature. 
-Among individuals of the same species it is the basis for the evolution of  higher 
-forms of biological organization [@maynardsmith_Major_1998]. 
-Among individuals of different species it underlays the biological mechanism for 
-fundamental ecological interactions such as pollination, nitrogen fixation or 
-parasite removal [@begon_Ecology_2006]. However, when cooperation is costly, 
-it is not straightforward to understand why individuals would be willing to pay that cost [@lehmann_Evolution_2006]. Furthermore, cooperative interactions might open the way for 
-one partner to exploit the other, what prevents this type of exploitation? [@herre_Evolution_1999]
-One reason for cooperating is if the benefits obtained from a partner´s 
-cooperative action are dependent on one's own cooperative act. 
-This dependency can arise, for instance, when cooperative partners
-choose who they interact with depending on the level of cooperation they offer. 
-In line with this, cooperative interactions can be viewed as a market-like situation 
-where partner choice mediates the level of exchange between traders [@hammerstein_Biological_2016; @leimar_Cooperation_2010; @noe_Biological_1994a]. Empirical evidence shows
-that several iconic examples of cooperation between individuals of different species 
-are indeed mediated by partner choice [@hammerstein_Biological_2016]. Biological market theory highlights 
-that changes in the supply and demand of goods and services determines the value of 
-those commodities, and therefore, can influence partner's actions in the cooperative 
-exchange. For instance, in the marine cleaning mutualism, where cleaner fish 
-_Labroides dimidiatus_ (cleaners) removes ectoparasites from other coral reef fishes (clients) [@cote_Evolution_2000],
-some client species are able to choose among different cleaners, and because of this
-obtain better service [@grutter_Cleaner_2003]. Better services means for clients to have priority when 
-more than one client is available [@bshary_Choosy_2002], and less cheating (when cleaners eat live tissue
-instead of parasites) [@grutter_Cleaner_2003]. However, recent evidence shows that when clients experience
-ecological conditions with low cleaner abundance, they are less likely to use partner
-choice to leverage better service [@triki_Biological_2019]. On the other side, 
-cleaners in those markets with favourable conditions, stop giving priority to choosy clients [@triki_Biological_2019; @triki_Decrease_2018]. Thus, it seems that cleaners, as well as clients,
-adjust their strategy to the prevalent market conditions they find in their environment. 
-This adjustment cannot be due to local genetic adaptation, given that cleaners are open water spawners 
-and the environmental changes in different sites have occurred in the last decade [@triki_Biological_2019; @triki_Decrease_2018]. Alternatively, a recent cognitive model proposes that differences in 
-cleaners preferences can arise trough the interaction of cleaner's cognitive processes 
-with the environmental conditions [@quinones_Reinforcement_2019]. Under this view,
-cleaners ontogenetically develop their preferences for different clients through learning,
-and some environmental conditions facilitate learning. Thus, the adaptive adjustment 
-seen in cleaners might be due to their learning mechanisms. Despite the general fit of the 
-model explanation to the available data, no systematic parameter fit has been performed to 
-evaluate which combination of model parameters better fit the data.
+Biological market theory emphasizes the importance of partner choice in
+explaining how supply-to-demand ratios translate into payoff asymmetries
+between partners [@noe_Biological_1994a; @noe_Biological_1995a]. One
+example is the marine cleaning mutualism involving the cleaner
+fish *Labroides dimidiatus* and 'client' fish. Clients offer themselves as
+food patches for cleaners to remove ectoparasites, which gives the
+cleaners calories and the clients improved health
+[@waldie_LongTerm_2011; @ros_Does_2011,@triki_Effects_2016]. As
+cleaners have about 2000 interactions per day
+[@grutter_Relationship_1995], they regularly face the situation that
+two clients seek cleaning simultaneously. In such conditions, it is
+important to distinguish between 'resident' and 'visitor' client species
+[@bshary_Cleaner_2001a]. The residents have small home ranges and hence
+access to only the local cleaner, while the visitors have large home
+ranges and hence access to two or more cleaners. Thus, visitors have the
+option to switch to another cleaner if being made to wait, while
+residents must wait for inspection. Indeed, visitors have been observed
+to use their partner choice options in that way [@bshary_Choosy_2002],
+which may explain why cleaners give visitors service priority in a field
+study in the Red Sea [@bshary_Cleaner_2001a]. Furthermore, early
+laboratory experiments showed that leaving causes service priority
+[@bshary_Asymmetric_2002]. The experiments consisted of a simultaneous
+two-choice task, where cleaners had to decide which of two Plexiglas
+plates to approach first and eat mashed prawn of it. One plate was made
+to behave like a resident, i.e. it would remain until the cleaner had
+eaten off it. The other plate was made to behave like a visitor, i.e. it
+would be removed if not inspected first. Cleaners learned to prefer the
+visitor plate and hence obtained the double amount of food
+[@bshary_Asymmetric_2002]. Adult cleaner fish outperform various
+primates as well as rats and pigeons in this original version of the
+biological market task [@salwiczek_Adult_2012; @zentall_Early_2017];
+only African grey parrots solve this task as well [@pepperberg_Can_2014].
+
+Recent research has refined our view on partner choice and priority of
+access in the *L. dimidiatus* cleaning mutualism, and as a consequence
+performance in the biological market task. First, details of the task
+presentation may matter; i.e. what relevant information is encoded in
+the design of the two plates (colour versus size; @wismer_Cuebased_2019), or
+whether the information is in the colour of the food
+[@pretot_Comparative_2021; @pretot_Comparing_2016], and whether the
+food is visible or hidden/released only when the subjects made choices
+on a touch screen [@pretot_Factors_2016]. Second, it appears that cleaner and
+client densities in nature affect how cleaners perform in the original
+market task: when densities are low, cleaners are more likely to fail to
+complete the task [@triki_Biological_2019; @triki_Decrease_2018;
+@wismer_Variation_2014]. The apparent explanation for this failure is that visitor
+clients are less likely to exert their choice options (and hence wait
+for inspection if made to wait) when cleaner fish densities are low
+[@triki_Brain_2020]. This change in behavior may be due to the fact that
+when cleaner densities are low, the cost of switching in terms of
+distance to swim and time invested increases. If visitors are not
+choosy, cleaners should not give them priority anymore. Interestingly,
+while cleaners from high-density areas have larger forebrains harbouring
+more cells than cleaners from low-density areas [@triki_Brain_2020], on a
+local scale those individuals that perform in the market task in a way
+that corresponds to the locally optimal strategy (solving/failing the
+task when living in high/low density) have relatively larger forebrains
+[@triki_Brain_2020]. Taken together, recent empirical results clearly show
+that we need a mechanistic approach to understand variation in intra-
+and interspecific performance in the market task. One approach to
+identify candidate mechanisms underlying performance in a cognitive task
+is to develop proximate learning models. By varying the cognitive tool
+kit, one can assess which minimal kit is necessary to solve the task at
+hand. This approach has been used to analyse the market task
+[@prat_Role_2021; @quinones_Reinforcement_2019] and spatial learning in bees [@dubois_Model_2021]. Learning models showed that basic reinforcement 
+learning does not suffice to solve the task, this is particularly so when model assume the more complex natural situation in which cleaners
+not only face resident-visitor pairs but also visitor-visitor and
+resident-resident pairs, as well only a resident or only a visitor. To
+be able to give visitors priority over residents, cleaners first need to
+be able to assess a client's value separately for the three possible
+forms of encounter (alone, paired with a fish with the same strategic
+option, paired with a fish with the alternative strategic option)
+[@quinones_Reinforcement_2019]. In addition, cleaners need to account for the future
+consequences of current decisions. In the model, this could be achieved
+in two non-mutually exclusive ways [@quinones_Reinforcement_2019]:
+through low temporal discounting of future effects
+and/or through perceiving a visitor client leaving as psychological
+punishment (i.e. as a negative reinforcer). Given that a negative
+reinforcer is part of the ubiquitous learning mechanism termed operant
+conditioning (Thorndike xxxx), this explanation can be considered the
+simpler one. Furthermore, accounting for the future effects of an action is 
+considered an advanced cognitive ability [@Suddenford_Evolution_2007]. 
+Here, we fitted our model to the empirical data on fish
+densities and cleaner fish performance in the market task with the aim
+to determine the likely mechanism used by cleaners to incorporate future
+consequences of current decisions.
+
+Cooperation among individuals of the same or different species is
+ubiquitous in nature. Among individuals of the same species it is the
+basis for the evolution of higher forms of biological organization
+[@maynardsmith_Major_1998]. Among individuals of different species it
+underlays the biological mechanism for fundamental ecological
+interactions such as pollination, nitrogen fixation or parasite removal
+[@begon_Ecology_2006]. However, when cooperation is costly, it is not
+straightforward to understand why individuals would be willing to pay
+that cost [@lehmann_Evolution_2006]. Furthermore, cooperative
+interactions might open the way for one partner to exploit the other,
+what prevents this type of exploitation? [@herre_Evolution_1999] One
+reason for cooperating is if the benefits obtained from a partner´s
+cooperative action are dependent on one's own cooperative act. This
+dependency can arise, for instance, when cooperative partners choose who
+they interact with depending on the level of cooperation they offer. In
+line with this, cooperative interactions can be viewed as a market-like
+situation where partner choice mediates the level of exchange between
+traders [@hammerstein_Biological_2016; @leimar_Cooperation_2010;
+@noe_Biological_1994a]. Empirical evidence shows that several iconic
+examples of cooperation between individuals of different species are
+indeed mediated by partner choice [@hammerstein_Biological_2016].
+Biological market theory highlights that changes in the supply and
+demand of goods and services determines the value of those commodities,
+and therefore, can influence partner's actions in the cooperative
+exchange. For instance, in the marine cleaning mutualism, where cleaner
+fish *Labroides dimidiatus* (cleaners) removes ectoparasites from other
+coral reef fishes (clients) [@cote_Evolution_2000], some client
+species are able to choose among different cleaners, and because of this
+obtain better service [@grutter_Cleaner_2003]. Better services means
+for clients to have priority when more than one client is available
+[@bshary_Choosy_2002], and less cheating (when cleaners eat live
+tissue instead of parasites) [@grutter_Cleaner_2003]. However, recent
+evidence shows that when clients experience ecological conditions with
+low cleaner abundance, they are less likely to use partner choice to
+leverage better service [@triki_Biological_2019]. On the other side,
+cleaners in those markets with favourable conditions, stop giving
+priority to choosy clients [@triki_Biological_2019;
+@triki_Decrease_2018]. Thus, it seems that cleaners, as well as
+clients, adjust their strategy to the prevalent market conditions they
+find in their environment. This adjustment cannot be due to local
+genetic adaptation, given that cleaners are open water spawners and the
+environmental changes in different sites have occurred in the last
+decade [@triki_Biological_2019; @triki_Decrease_2018]. Alternatively, a
+recent cognitive model proposes that differences in cleaners preferences
+can arise trough the interaction of cleaner's cognitive processes with
+the environmental conditions [@quinones_Reinforcement_2019]. Under this
+view, cleaners ontogenetically develop their preferences for different
+clients through learning, and some environmental conditions facilitate
+learning. Thus, the adaptive adjustment seen in cleaners might be due to
+their learning mechanisms. Despite the general fit of the model
+explanation to the available data, no systematic parameter fit has been
+performed to evaluate which combination of model parameters better fit
+the data.
 
@@
 
-In order to fit the model parameters to the data we used Markov Chain Monte Carlo (MCMC) with 
-a Metropolis sampler to estimate the posterior distribution of the model parameters.
-We used the data from @triki_Biological_2019 to fit the parameters of the  model analysed in @quinones_Reinforcement_2019.
-The data consist a of the number of times that a set of cleaners chose to
-feed of a visitor client plate in the context of the market task.
-This set of cleaners were collected in different locations and in each one of those locations environmental variables were measured.
-These environmental variables recorded from each location are used as input for the computational model.
-Given these inputs, the computational model produces a time series of choices in the market task.
-In the model the preference for the two options changes as the algorithm learns based on the reward that the two options provide.
-Thus, we run the model for 10000 steps and use the last 3000 steps as the preference developed in the model for visitors given the environmental variables.
-We use this preference as the predicted probability of choosing a visitor in the data, and use it to calculate the likelihood that the number of visitor choices obtained from each cleaner is generated by the model.
-We sum over all the cleaners the log-likehoods and use the total for the acceptance of new parameters in the MCMC.
+In order to fit the model parameters to the data we used Markov Chain
+Monte Carlo (MCMC) with a Metropolis sampler to estimate the posterior
+distribution of the model parameters. We used the data from
+@triki_Biological_2019 to fit the parameters of the model analysed in
+@quinones_Reinforcement_2019. The data consist a of the number of times
+that a set of cleaners chose to feed of a visitor client plate in the
+context of the market task. This set of cleaners were collected in
+different locations and in each one of those locations environmental
+variables were measured. These environmental variables recorded from
+each location are used as input for the computational model. Given these
+inputs, the computational model produces a time series of choices in the
+market task. In the model the preference for the two options changes as
+the algorithm learns based on the reward that the two options provide.
+Thus, we run the model for 10000 steps and use the last 3000 steps as
+the preference developed in the model for visitors given the
+environmental variables. We use this preference as the predicted
+probability of choosing a visitor in the data, and use it to calculate
+the likelihood that the number of visitor choices obtained from each
+cleaner is generated by the model. We sum over all the cleaners the
+log-likehoods and use the total for the acceptance of new parameters in
+the MCMC.
 
-The model consist of a set individual-based simulations where individuals face a series of choices between two options.
-Individuals in the model represent cleaners, and the options they face simulate the natural conditions of a cleaning market.
-Thus, individuals in the model choose to prioritize  different clients.
-These clients are classified as resident and visitors.
-Residents are those clients that stay in the cleaning queue when they are not given priority; while visitors leave the queue (with a certain probability) after they are not given priority.
-Individuals obtain reward from cleaning a client regardless of the type.
-Every time individuals face and make a choice they update the probability of making that same choice.
-The update is based on the difference between the expectation of reward and the actual reward the individual obtains (prediction error); and it is carried in the direction that would lead to more reward being obtained, given the new information.
-Further details of the model implementation can be found in the paper [@quinones_Reinforcement_2019].
-The model analysis points to the relevance of two parameters.
-First, $\gamma$ measures how much individuals include future rewards in their decision updates.
-If $\gamma=0$, individuals only use the immediate reward obtained from a cleaning interaction.
-As $\gamma$ increases individuals include more the reward obtained from the subsequent choices.
-That amounts to calculating and using for decision making the future expected rewards.
-Second, $\eta$ measures how much individuals include in their reward the fleeing behaviour of visitor clients as a negative component.
-Both of these parameters allow individuals to use in their estimates the future effects of their choices.
-The model analysis shows that without either of them, learning individuals cannot develop a preference for the visitor client.
-The analysis contrast the effect of these two parameters against variation in the relative abundance of the two client types.
-The analysis shows that under very high client abundance the presence of  future rewards ($\gamma>0$) no longer favours the preference for the ephemeral clients, this effect does not occur with negative reward mediated by client behaviour ($\eta$).
-Thus, the analysis suggests that future reward as an explanation for cleaner ability to prefer visitors is supported by @triki_Biological_2019 data.
-However, a combination of future and negative rewards, could still be a better explanation for the data.
-In the current analysis, we will use the data from @triki_Biological_2019 to fit some of the model parameters, particularly relevant we will fit $\gamma$ and $\eta$ to find out whether each or a combination of these effects is a better explanation for the pattern seen in the data.
+The model consist of a set individual-based simulations where
+individuals face a series of choices between two options. Individuals in
+the model represent cleaners, and the options they face simulate the
+natural conditions of a cleaning market. Thus, individuals in the model
+choose to prioritize different clients. These clients are classified as
+resident and visitors. Residents are those clients that stay in the
+cleaning queue when they are not given priority; while visitors leave
+the queue (with a certain probability) after they are not given
+priority. Individuals obtain reward from cleaning a client regardless of
+the type. Every time individuals face and make a choice they update the
+probability of making that same choice. The update is based on the
+difference between the expectation of reward and the actual reward the
+individual obtains (prediction error); and it is carried in the
+direction that would lead to more reward being obtained, given the new
+information. Further details of the model implementation can be found in
+the paper [@quinones_Reinforcement_2019]. The model analysis points to
+the relevance of two parameters. First, $\gamma$ measures how much
+individuals include future rewards in their decision updates. If
+$\gamma=0$, individuals only use the immediate reward obtained from a
+cleaning interaction. As $\gamma$ increases individuals include more the
+reward obtained from the subsequent choices. That amounts to calculating
+and using for decision making the future expected rewards. Second,
+$\eta$ measures how much individuals include in their reward the fleeing
+behaviour of visitor clients as a negative component. Both of these
+parameters allow individuals to use in their estimates the future
+effects of their choices. The model analysis shows that without either
+of them, learning individuals cannot develop a preference for the
+visitor client. The analysis contrast the effect of these two parameters
+against variation in the relative abundance of the two client types. The
+analysis shows that under very high client abundance the presence of
+future rewards ($\gamma>0$) no longer favours the preference for the
+ephemeral clients, this effect does not occur with negative reward
+mediated by client behaviour ($\eta$). Thus, the analysis suggests that
+future reward as an explanation for cleaner ability to prefer visitors
+is supported by @triki_Biological_2019 data. However, a combination of
+future and negative rewards, could still be a better explanation for the
+data. In the current analysis, we will use the data from
+@triki_Biological_2019 to fit some of the model parameters, particularly
+relevant we will fit $\gamma$ and $\eta$ to find out whether each or a
+combination of these effects is a better explanation for the pattern
+seen in the data.
 
@@
 
-In order to capture with the model the relationship between the characteristics of the market (environment) and the cleaner's performance we need to translate the absolute abundances of the data to relative abundances.
-This is because in the model, relative abundances of clients define not only the probability of residents and visitors, but also how often the cleaning station is empty (e.g.
-there are no clients to be cleaned).
-The more cleaners, the more options available for the clients and hence higher probability of empty cleaning stations.
-How often the cleaning stations are empty in turns influences the value that cleaners can expect from their choices.
-The more empty the empty cleaning stations are, the more it is to choose a client that will not stay.
-However, the frequency of empty spots in the cleaning station does not only depend on cleaner relative abundances but also on the frequency with which clients visit the station.
-The client visitation frequency will likely vary among different clients depending on parasite abundances and loads.
-These are variables for which we do not have field estimates.
-Thus, we
-computed a measure of relative cleaner abundance for each site relative to absolute abundances and multiplied by a scaling constant that changes the range of the variable.
-We fit the value of the scaling constant as part of the statistical inference.
-As for the client abundances for residents and visitors we computed a relative measure with respect to the total client abundance, and weighted that by the rescaled relative absence of cleaners (1-relative cleaner abundance).
-Thus, all three measures of relative abundance sum up to 1, and can be used in the model as a proxy for the probability of having different options in the cleaning station.
+In order to capture with the model the relationship between the
+characteristics of the market (environment) and the cleaner's
+performance we need to translate the absolute abundances of the data to
+relative abundances. This is because in the model, relative abundances
+of clients define not only the probability of residents and visitors,
+but also how often the cleaning station is empty (e.g. there are no
+clients to be cleaned). The more cleaners, the more options available
+for the clients and hence higher probability of empty cleaning stations.
+How often the cleaning stations are empty in turns influences the value
+that cleaners can expect from their choices. The more empty the empty
+cleaning stations are, the more it is to choose a client that will not
+stay. However, the frequency of empty spots in the cleaning station does
+not only depend on cleaner relative abundances but also on the frequency
+with which clients visit the station. The client visitation frequency
+will likely vary among different clients depending on parasite
+abundances and loads. These are variables for which we do not have field
+estimates. Thus, we computed a measure of relative cleaner abundance for
+each site relative to absolute abundances and multiplied by a scaling
+constant that changes the range of the variable. We fit the value of the
+scaling constant as part of the statistical inference. As for the client
+abundances for residents and visitors we computed a relative measure
+with respect to the total client abundance, and weighted that by the
+rescaled relative absence of cleaners (1-relative cleaner abundance).
+Thus, all three measures of relative abundance sum up to 1, and can be
+used in the model as a proxy for the probability of having different
+options in the cleaning station.
 
@@
 ```{r post, fig.cap="Posterior distributions for parameter values $\\gamma$, $\\eta$ and the scalling constant.
-We show the kerne density estimates, and bellow the mean (black point) and the 65\\% and 95\\%  highest posterior density interval for the three paramteres.
+We show the kerne density estimates, and bellow the mode (black point) and the 65\\% (light blue) and 95\\% (grey)  highest posterior density interval for the three paramteres.
 On the left (panels a, c and c) posterior distributions for a model with all three parameters; while on the right (b and d) a model without negative reward, ",out.width='85%'}
@@
 
-Figure \ref{fig:post} panels a,c and e show the posterior distributions of the parameter values 
-of the full model, which includes both future and negative reward. Notably, the bulk of the posterior distribution 
-for the parameter tuning negative reward ($\eta$) is close to zero.
-Given the structure of the model, we prevented $\eta$ from taking negative values.
-Nevertheless, the shape of the posterior and its credible intervals suggest the absence of negative reward to be well supported by the data.
-Thus, we run the analysis setting $\eta$ to zero.
-Figure \ref{fig:post} panels b and d, show the posterior distributions of the two parameters estimated.
-Not surprisingly, under a model without negative reward, the distribution of $\gamma$ shifts to higher values.
+Figure \ref{fig:post} panels a,c and e show the posterior distributions
+of the parameter values of the full model, which includes both future
+and negative reward. Notably, the bulk of the posterior distribution for
+the parameter tuning negative reward ($\eta$) is close to zero. Given
+the structure of the model, we prevented $\eta$ from taking negative
+values. Nevertheless, the shape of the posterior and its credible
+intervals suggest the absence of negative reward to be well supported by
+the data. Thus, we run the analysis setting $\eta$ to zero. Figure
+\ref{fig:post} panels b and d, show the posterior distributions of the
+two parameters estimated. Not surprisingly, under a model without
+negative reward, the distribution of $\gamma$ shifts to higher values.
 
@@
 
-Figure \ref{fig:pred} shows predictions from the two models and their fit to the data. On the left,
-the colour coded contour shows the prediction of the model for combinations of cleaner abundance (x axis) and
-visitor leaving probability (y axis). Predictions were generated by running the model using the mode of the posterior
-distributions as the parameter values, and varying systematically cleaner abundance and visitor leaving probability.
-Colour of the points on top of the contour show the average frequency with 
-which cleaner fish chose the visitor plate in their last 20 experimental trials. Each point corresponds to a 
-location with a combination of cleaner abundance and visitor leaving probability. Panels on the right show how close
-the observed values (y axis) are from the predictions of the models (x axis).
-The line represents the perfect fit between observed and predicted values.
+Figure \ref{fig:pred} shows predictions from the two models and their
+fit to the data. On the left, the colour coded contour shows the
+prediction of the model for combinations of cleaner abundance (x axis)
+and visitor leaving probability (y axis). Predictions were generated by
+running the model using the mode of the posterior distributions as the
+parameter values, and varying systematically cleaner abundance and
+visitor leaving probability. Colour of the points on top of the contour
+show the average frequency with which cleaner fish chose the visitor
+plate in their last 20 experimental trials. Each point corresponds to a
+location with a combination of cleaner abundance and visitor leaving
+probability. Panels on the right show how close the observed values (y
+axis) are from the predictions of the models (x axis). The line
+represents the perfect fit between observed and predicted values.
 Predicted values were taken using the mode just as in the left panels.
-On the lower right corner of the panel we show the coefficient of determination.
-Despite having one more parameter varying, the full model gives a lower fit to the data than the model without negative reward, measured by the coefficient of determination.
-In the supplementary material, we also show that a model with only negative reward, excluding future rewards ($\gamma=0$, Fig.
-\ref{fig:nogamma}), provides a lower fit than either of the ones presented in the main text.
+On the lower right corner of the panel we show the coefficient of
+determination. Despite having one more parameter varying, the full model
+gives a lower fit to the data than the model without negative reward,
+measured by the coefficient of determination. In the supplementary
+material, we also show that a model with only negative reward, excluding
+future rewards ($\gamma=0$, Fig. \ref{fig:nogamma}), provides a lower
+fit than either of the ones presented in the main text.
 
-The main reason for negative and future reward to give different predictions is the way that cleaner's relative abundance influences the preference for the visitor clients.
-Visitor leaving probability has a similar positive effect on the probability of choosing the visitor clients on all three models
-(Fig.
-\ref{fig:pred}, \ref{fig:nogamma}).
-In contrast, cleaner's relative abundance has a different effect in the model with only future reward, compared to the other two.
-In the model with only future reward
-only intermediate cleaner abundance triggers a preference for the visitor clients 
-(Fig. \ref{fig:pred} c). In models with negative reward, both intermediate and low cleaner abundance
-trigger a preference for the visitor. (Fig. \ref{fig:pred} a, \ref{fig:nogamma})
+The main reason for negative and future reward to give different
+predictions is the way that cleaner's relative abundance influences the
+preference for the visitor clients. Visitor leaving probability has a
+similar positive effect on the probability of choosing the visitor
+clients on all three models (Fig. \ref{fig:pred}, \ref{fig:nogamma}). In
+contrast, cleaner's relative abundance has a different effect in the
+model with only future reward, compared to the other two. In the model
+with only future reward only intermediate cleaner abundance triggers a
+preference for the visitor clients (Fig. \ref{fig:pred} c). In models
+with negative reward, both intermediate and low cleaner abundance
+trigger a preference for the visitor. (Fig. \ref{fig:pred} a,
+\ref{fig:nogamma}).
 
@@
 
-<div id="refs"></div>
+::: {#refs}
+:::
 
@@
 
+```{=tex}
 \beginsupplement
-
+```
 ```{r nogamma, fig.cap="Observed and predicted probability of choosing a visitor for a model without future rewards.

### 2021-11-10 ActCritABC.cpp

@@ -962,10 +962,10 @@ model_param perturb_parameters_uniform(model_param focal_param, json &sim_param)
 		float(sim_param["sdPert"][1]), 0, INFINITY)+
 		(!bool(sim_param["pertScen"][1]))*focal_param.alphaC;
 	new_param.gamma = bool(sim_param["pertScen"][2]) * boundedParUnifPert(focal_param.gamma,
-		sim_param["sdPert"][2], 0, 1) +
+		sim_param["sdPert"][2], -1, 1) +
 		(!bool(sim_param["pertScen"][2]))*focal_param.gamma;
 	new_param.negReward = bool(sim_param["pertScen"][3])* boundedParUnifPert(focal_param.negReward,
-		sim_param["sdPert"][3], 0, INFINITY)+
+		sim_param["sdPert"][3], -INFINITY, INFINITY)+
 		(!bool(sim_param["pertScen"][3]))*focal_param.negReward;
 	new_param.scaleConst = bool(sim_param["pertScen"][4])* boundedParUnifPert(focal_param.scaleConst,
 		sim_param["sdPert"][4], 0, INFINITY) +
@@ -1180,12 +1180,12 @@ int main(int argc, char* argv[]){
 		for (int r = 0; r < sim_param["chain_length"]; ++r) {  // 
 			//cout << "Iteration	" << r << endl;
 			model_param new_param = perturb_parameters_uniform(focal_param, sim_param);
-			if ((new_param.gamma < 1 && new_param.gamma >= 0) &&
-				((new_param.negReward <= INFINITY && new_param.negReward >= 0) &&
-				(new_param.scaleConst <= INFINITY && new_param.scaleConst >= 0))) {
+			if ((new_param.gamma < 1 && new_param.gamma > -1) &&
+				((new_param.negReward <= INFINITY && new_param.negReward >= -INFINITY) &&
+				(new_param.scaleConst <= INFINITY && new_param.scaleConst > 0))) {
 				if (sim_param["data"] == "loc") {
 					do_simulation(//focal_model, 
-						emp_data_loc, focal_param, sim_param);	
+						emp_data_loc, focal_param, sim_param);
 					curr_loglike = calculate_fit(emp_data_loc);
 					do_simulation(//focal_model, 
 						emp_data_loc, new_param, sim_param);

### 2021-11-10 analysisABC.R

@@ -13,13 +13,12 @@ library(cowplot)
 library(ggdist)
 library(bayesplot) 
 library(lattice)
-library(ggridges)
 source(here("../R_files/posPlots.R"))
 library("jsonlite")
 
-simdir<-"D:/Neuch_simulations/ABCfit/Simulations/"
+# simdir<-"D:/Neuch_simulations/ABCfit/Simulations/"
 
-scen1<-"ABCclean_gam_nRew_sca_"
+scen1<-"MCMCclean_gam_Nrew_sca_"
 scen2<-"ABCclean_gam_sca_"
 
 ## Load files --------------------------------------------------------------
@@ -50,6 +49,7 @@ ABCraw.2<-do.call(rbind,lapply(ABCruns2, function(file){
 
 # One per onject
 ABCraw<-fread(here("Simulations",scen,ABCruns[3]))
+
 
 # All in a list
 mcmcList.1<-mcmc.list(do.call(list,lapply(ABCruns1, function(file){
@@ -182,72 +182,104 @@ mtext(text = "iteration" ,line = 2,cex = 1,side = 1,las=0)
 
 modeGam;modenegReward;modeScal
 
+ABCraw.1[negReward<0]
+
 plot(data=ABCfiltered.1,gamma~negReward,type="p",cex=0.2)
 plot(data=ABCfiltered.1,negReward~scaleConst,type="p",cex=0.2)
 plot(data=ABCfiltered.1,gamma~scaleConst,type="p",cex=0.2)
 
 ## Density plots with gg -------------------------------------------------------
 
-ABCfiltered[round(runif(10,min = 1,max=dim(ABCfiltered)[1]),digits = 0),]
-
-ABCfilt.long.1<-melt(ABCfiltered.1,measure.vars = c("gamma","negReward","scaleConst"),
-                   variable.name = "variable",id.vars = "seed")
-
-ABCfilt.long.2<-melt(ABCfiltered.2,measure.vars = c("gamma","negReward","scaleConst"),
-                     variable.name = "variable",id.vars = "seed")
-
-
-typeof(mcmcList)
-
-areas<-
-  mcmc_areas(x = mcmcList)+
-  facet_wrap(~parameter,scales = "free_x")
-
-
-areas.data<-mcmc_areas_data(x=mcmcList)
-
-densplots<-ggplot(data = ABCfilt.long,aes(x=value,y=variable))+
-  stat_halfeye()
-
[…]

### 2021-11-10 loadData.R

@@ -210,3 +210,20 @@ check_create.dir<-function(folder,param,values){
     return(listfolders)
   }
 }
+
+loadMCMCrun<-function(scen,burn.in=1000,thinning=100){
+  listFiles<-list.files(here("Simulations",scen))
+  runs<-grep("MCMCchain",listFiles,value = TRUE)
+  raw<-do.call(rbind,lapply(runs, function(file){
+    rundata<-fread(here("Simulations",scen,file))
+    seedNum<-strsplit(file,"_")[[1]][3]
+    seedNum<-as.numeric(gsub("[[:alpha:]]",seedNum,replacement = ''))
+    rundata[,seed:=rep(seedNum,dim(rundata)[1])]
+  }))
+  filtered<-raw[iteration>burn.in & iteration %% thinning==0]
+  return(filtered)
+}
+  
+  
+  
+  

### 2021-11-10 manFigures.R

@@ -6,6 +6,7 @@ library("RColorBrewer")
 library("patchwork") 
 library(ggpubr)
 library(here)
+library(cowplot)
 source(here("loadData.R"))
 source(here("data2interp.R"))
 require('akima')
@@ -14,30 +15,64 @@ source(here("..","R_files",'ternaryAEQP.R'))
 source(here("../R_files/posPlots.R"))
 
 
-scenario<-"ABCclean_gam_Nrew_sca"
+scenario<-"ABCclean_gam_nRew_sca_2"
 scenario2<-"ABCclean_gam_sca"
 
-predfile.both<-grep(".txt",grep("round",list.files(here("Simulations",paste0(scenario,"_"))),value = T),
+predfileMode.both<-grep(".txt",grep("round",list.files(here("Simulations",paste0(scenario,"_"))),value = T),
                     value = T)
-predfile.gam<-grep(".txt",grep("round",list.files(here("Simulations",paste0(scenario2,"_"))),value = T),
+predfileMode.gam<-grep(".txt",grep("round",list.files(here("Simulations",paste0(scenario2,"_"))),value = T),
                    value = T)
 
-predictData.both<-fread(here("Simulations",paste0(scenario,"_"),
-                      predfile.both))
-predictData.gam<-fread(here("Simulations",paste0(scenario2,"_"),
-                             predfile.gam))
+predfileSamples.both<-grep(".txt",grep("round",list.files(here("Simulations",
+                                                               paste0(scenario,"_"),"samplesPost_"),
+                                                          full.names = TRUE)
+                                       ,value = T),
+                         value = T)
 
-str(predictData)
+predfileSamples.gam<-grep(".txt",grep("round",list.files(here("Simulations",
+                                                               paste0(scenario2,"_"),"samplesPost_"),
+                                                          full.names = TRUE)
+                                       ,value = T),
+                           value = T)
 
[…]

### 2021-11-10 parTOjson.R

@@ -18,6 +18,8 @@ scenario<-"ABCclean_nRew_sca"
 
 scenario<-"ABCclean_nRew_sca"
 
+MCMCdata<-loadMCMCrun(paste0(scenario,"_"))
+
 param_pred<-list(totRounds=10000,ResReward=1,VisReward=1,
             ResProbLeav=0,scenario=0, inbr=0,outbr=0,forRat=0.0,
             seed=1, propfullPrint = 0.7,sdPert=c(0.05,0.05,0.1,0.05,1),
@@ -26,40 +28,50 @@ param_pred<-list(totRounds=10000,ResReward=1,VisReward=1,
             pertScen = c(FALSE,FALSE,FALSE,FALSE,TRUE), 	
             MCMC =0, data="clean",nRep=30,
             dataFile = here("Data","data_ABC_cleaner_absolute.txt"),
-            folderL=paste0(here(simsDir),"/",scenario,"_/"))
+            folderL=paste0(here(simsDir),"/",scenario,"_/","samplesPost_/"))
 
 
-check_create.dir(here(simsDir),param = rep(scenario,1),values = c(""))
+
+check_create.dir(here(simsDir,paste0(scenario,"_")),param = rep("samplesPost",1),
+                 values = c(""))
 
 param_pred$folder<-param_pred$folderL
-param_pred$init[c(3,4,5)]<- c(modeGam,modenegReward,modeScal)
+# param_pred$init[c(3,4,5)]<- c(modeGam,modenegReward,modeScal)
   # c(runif(1,max = 1),runif(1,max = 1),
   #                            runif(1,max = 500,min = 1))
   #sumMCMC$statistics[,1]
-fileName<-paste("parameters_pred_",param_pred$seed,".json",sep="")
-outParam.pred<-toJSON(param_pred,auto_unbox = TRUE,pretty = TRUE)
-if(file.exists(paste(param_pred$folderL,fileName,sep = ''))){
-  currFile<-fromJSON(paste(param_pred$folderL,fileName,sep = ''))
-  if(sum(unlist(currFile)!=unlist(param_pred))>0){
-    # warning("You are erasing old files!! n\ Check first!!!",immediate. = TRUE)
-    # print("OLD value")
-    # print(unlist(currFile)[unlist(currFile)!=unlist(param)])
-    # print("NEW value")
-    # print(unlist(param)[unlist(currFile)!=unlist(param)])
-    # ans<-readline("Want to continue?")
-    # if(substr(ans, 1, 1) == "y"){
[…]

### 2021-11-16 manuscript_1.0.Rmd
started  with the dicussion in the manuscript. Used functions to make the data loading more straightforwd

@@
 author:
-- Andrés E. Quiñones^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland - andresqp@gmail.com]
-- Zegni Triki^[Department of Zoology, Stockholm University, Stockholm, Sweden]
-- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
+- Andrés E. Quiñones^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
+  - andresqp@gmail.com]
 date: "`r format(Sys.time(), '%B %d, %Y')`"
 output:
+  word_document: default
   pdf_document: default
@@
 - \linenumbers
-- \newcommand{\beginsupplement}{
-    \setcounter{table}{0}  
-    \renewcommand{\thetable}{S\arabic{table}} 
-    \setcounter{figure}{0} 
-    \renewcommand{\thefigure}{S\arabic{figure}}}
+- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
+  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
 keep_tex: yes
@@
 
-Biological market theory emphasizes the importance of partner choice in
-explaining how supply-to-demand ratios translate into payoff asymmetries
-between partners [@noe_Biological_1994a; @noe_Biological_1995a]. One
-example is the marine cleaning mutualism involving the cleaner
+Cooperation among individuals of the same or different species is
+ubiquitous in nature. Among individuals of different species it
+underlays the biological mechanism for fundamental ecological
+interactions such as pollination, nitrogen fixation or parasite removal
+[@begon_Ecology_2006]. When partners in these mutualistic interactions 
+are able to choose their interaction partner, their investment and the return 
+they obtain can be understood as a 
+biological market [@hammerstein_Biological_2016; @leimar_Cooperation_2010;
+@noe_Biological_1994a,@noe_Biological_1995a]. Thus, we expect supply 
+and demand of the good 
+and services being exchanged to determine the action of individuals, at least in
+an evolutionary time-scale.
+However, supply and demand conditions of a biological market, which typically
+depend on the abundance of the species involved, certainly vary in time and space. 
+How do partners adjust their strategy in the face of this variation? 
+In animals an obvious general candidate mechanism for the strategic adjustment is the 
+cognitive machinery. However, it is not clear which specific cognitive mechanisms
+allow individual to adjust their behaviour to the varying market conditions.
+
+One example of strategic adjustment in a biological market is the marine cleaning 
+mutualism involving the cleaner
 fish *Labroides dimidiatus* and 'client' fish. Clients offer themselves as
@@
 for inspection if made to wait) when cleaner fish densities are low
-[@triki_Brain_2020]. This change in behavior may be due to the fact that
+[@triki_Brain_2020]. This change in behaviour may be due to the fact that
 when cleaner densities are low, the cost of switching in terms of
@@
 [@triki_Brain_2020]. Taken together, recent empirical results clearly show
-that we need a mechanistic approach to understand variation in intra-
-and interspecific performance in the market task. One approach to
+that cognitive processes are involved in the pattern of intra-
+and inter-specific variation in  performance in the market task. One approach to
 identify candidate mechanisms underlying performance in a cognitive task
-is to develop proximate learning models. By varying the cognitive tool
+is to develop proximate learning models and to evaluated in light of the behavioural data 
+(e.g. @dubois_Model_2021). By varying the cognitive tool
 kit, one can assess which minimal kit is necessary to solve the task at
 hand. This approach has been used to analyse the market task
-[@prat_Role_2021; @quinones_Reinforcement_2019] and spatial learning in bees [@dubois_Model_2021]. Learning models showed that basic reinforcement 
-learning does not suffice to solve the task, this is particularly so when model assume the more complex natural situation in which cleaners
+[@prat_Role_2021; @quinones_Reinforcement_2019]. Learning models showed that basic reinforcement learning does not suffice to solve the task, this is 
+particularly so when model assume the more complex natural situation 
+in which cleaners
 not only face resident-visitor pairs but also visitor-visitor and
@@
 simpler one. Furthermore, accounting for the future effects of an action is 
-considered an advanced cognitive ability [@Suddenford_Evolution_2007]. 
+considered an advanced cognitive ability [@suddendorf_Evolution_2007]. 
 Here, we fitted our model to the empirical data on fish
@@
 
-Cooperation among individuals of the same or different species is
-ubiquitous in nature. Among individuals of the same species it is the
-basis for the evolution of higher forms of biological organization
-[@maynardsmith_Major_1998]. Among individuals of different species it
-underlays the biological mechanism for fundamental ecological
-interactions such as pollination, nitrogen fixation or parasite removal
-[@begon_Ecology_2006]. However, when cooperation is costly, it is not
-straightforward to understand why individuals would be willing to pay
-that cost [@lehmann_Evolution_2006]. Furthermore, cooperative
-interactions might open the way for one partner to exploit the other,
-what prevents this type of exploitation? [@herre_Evolution_1999] One
-reason for cooperating is if the benefits obtained from a partner´s
-cooperative action are dependent on one's own cooperative act. This
-dependency can arise, for instance, when cooperative partners choose who
-they interact with depending on the level of cooperation they offer. In
-line with this, cooperative interactions can be viewed as a market-like
-situation where partner choice mediates the level of exchange between
-traders [@hammerstein_Biological_2016; @leimar_Cooperation_2010;
-@noe_Biological_1994a]. Empirical evidence shows that several iconic
-examples of cooperation between individuals of different species are
-indeed mediated by partner choice [@hammerstein_Biological_2016].
-Biological market theory highlights that changes in the supply and
-demand of goods and services determines the value of those commodities,
-and therefore, can influence partner's actions in the cooperative
-exchange. For instance, in the marine cleaning mutualism, where cleaner
-fish *Labroides dimidiatus* (cleaners) removes ectoparasites from other
-coral reef fishes (clients) [@cote_Evolution_2000], some client
-species are able to choose among different cleaners, and because of this
-obtain better service [@grutter_Cleaner_2003]. Better services means
-for clients to have priority when more than one client is available
-[@bshary_Choosy_2002], and less cheating (when cleaners eat live
-tissue instead of parasites) [@grutter_Cleaner_2003]. However, recent
-evidence shows that when clients experience ecological conditions with
-low cleaner abundance, they are less likely to use partner choice to
-leverage better service [@triki_Biological_2019]. On the other side,
-cleaners in those markets with favourable conditions, stop giving
-priority to choosy clients [@triki_Biological_2019;
-@triki_Decrease_2018]. Thus, it seems that cleaners, as well as
-clients, adjust their strategy to the prevalent market conditions they
-find in their environment. This adjustment cannot be due to local
-genetic adaptation, given that cleaners are open water spawners and the
-environmental changes in different sites have occurred in the last
-decade [@triki_Biological_2019; @triki_Decrease_2018]. Alternatively, a
-recent cognitive model proposes that differences in cleaners preferences
-can arise trough the interaction of cleaner's cognitive processes with
-the environmental conditions [@quinones_Reinforcement_2019]. Under this
-view, cleaners ontogenetically develop their preferences for different
-clients through learning, and some environmental conditions facilitate
-learning. Thus, the adaptive adjustment seen in cleaners might be due to
-their learning mechanisms. Despite the general fit of the model
-explanation to the available data, no systematic parameter fit has been
-performed to evaluate which combination of model parameters better fit
-the data.
-
 ## Methods
 
-In order to fit the model parameters to the data we used Markov Chain
-Monte Carlo (MCMC) with a Metropolis sampler to estimate the posterior
-distribution of the model parameters. We used the data from
-@triki_Biological_2019 to fit the parameters of the model analysed in
-@quinones_Reinforcement_2019. The data consist a of the number of times
+### Overview
+
+We used the data from @triki_Biological_2019 to fit the parameters of the model analysed in
+@quinones_Reinforcement_2019. For the bayesian fitting procedure we used Markov Chain Monte Carlo (MCMC) with a Metropolis sampler to estimate the posterior
+distribution of the model parameters.  The data consist a of the number of times
 that a set of cleaners chose to feed of a visitor client plate in the
@@
 environmental variables. We use this preference as the predicted
-probability of choosing a visitor in the data, and use it to calculate
+probability in a binomial distribution to calculate
 the likelihood that the number of visitor choices obtained from each
@@
 the MCMC.
+
+### The model
 
@@
 reward obtained from the subsequent choices. That amounts to calculating
-and using for decision making the future expected rewards. Second,
+and using for decision making the future expected rewards in action. Second,
 $\eta$ measures how much individuals include in their reward the fleeing
@@
 against variation in the relative abundance of the two client types. The
-analysis shows that under very high client abundance the presence of
+analysis shows that under very high client (relative) abundance the presence of
 future rewards ($\gamma>0$) no longer favours the preference for the
@@
 
-```{r post, fig.cap="Posterior distributions for parameter values $\\gamma$, $\\eta$ and the scalling constant.
-We show the kerne density estimates, and bellow the mode (black point) and the 65\\% (light blue) and 95\\% (grey)  highest posterior density interval for the three paramteres.
-On the left (panels a, c and c) posterior distributions for a model with all three parameters; while on the right (b and d) a model without negative reward, ",out.width='85%'}
+```{r post, fig.cap="Posterior distributions for parameter values $\\gamma$, $\\eta$ and the scaling constant.
+We show the kernel density estimates, below the mode (black point) and the 65\\% (light blue) and 95\\% (grey)  highest posterior density interval for the three paramteres.
+On the left (panels a, c and c) posterior distributions for a model with   all three parameters; while on the right (b and d) a model without negative reward.
+Panel f shows a measure of fit for both models, namely the distribution of pseudo-$R^2$ obtained from sampling the posterior distribution of parameter values.",out.width='85%'}
 knitr::include_graphics(here("post_both_gamma.png"))
@@
 and negative reward. Notably, the bulk of the posterior distribution for
-the parameter tuning negative reward ($\eta$) is close to zero. Given
-the structure of the model, we prevented $\eta$ from taking negative
-values. Nevertheless, the shape of the posterior and its credible
-intervals suggest the absence of negative reward to be well supported by
-the data. Thus, we run the analysis setting $\eta$ to zero. Figure
+the parameter tuning negative reward ($\eta$ panerl e) is around zero.
+Thus, the shape of the posterior and its credible intervals suggest the absence of negative reward to be well supported by the data.
+As for $\gamma$, the confidence intervals also includes zero, but the mode of the posterior is around 0.5.
+Thus, we run the analysis setting $\eta$ to zero.
+Figure
 \ref{fig:post} panels b and d, show the posterior distributions of the
 two parameters estimated. Not surprisingly, under a model without
-negative reward, the distribution of $\gamma$ shifts to higher values.
+negative reward, the distribution of $\gamma$ shifts to higher values. Under the new model, the estimates of $\gamma$ do not include zero, the distribution is centered arounf $0.5$.
+To asses the fit of the two models we calculated the  $pseudo-R^2$ [@mcfadden_Conditional_1974], which is a standard measure of fit for logistic regression. 
+In that context, $pseudo-R^2$ uses the log-likelihood of the data given the model, relative to the log-likelihood of the data given a model without covariates, as a measure of fit. 
+Given that our model is not a logistic regression, we measure the $pseudo-R^2$ 
+in relation to the log-likelihood of a model with parameters $\gamma$ and $\eta$ set to zero. This in practice amounts to a model that has neutral preference between 
+the two options. In panel f of figure \ref{fig:post} we show the distribution of $pseudo-R^2$ calculated using samples from the posterior distributions shown before, as well as from a 
+model with negative reward but without future reward. 
+Note, $pseudo-R^2$ can have negative values, that is when the log-likelihood of 
+the model is lower than that of a model that triggers neutral preferences. 
+Even though, the peak of the three $pseudo-R^2$ distributions is not very different, 
+the model with only future reward produces a distribution of $pseudo-R^2$ where more values 
+are positive (to the right of black line in Fig. \ref{fig:post} f). This 
+shows that, accounting for variation in the parameter estimates, the model with only future
+reward gives a better fit to the data, despite having one parameter less.
 
@@
 Line correspond to a perfect match between observed and predicted probabilities.
+Pseudo $R^2$ shown in the lower right corner corresponds to the fit of a model parameterized with the mode of the posterior distributions.
 Upper panels show prediction from a model including future and negative reward, lower panels froma  model with only future reward.",out.width='90%'}
@@
 
-Figure \ref{fig:pred} shows predictions from the two models and their
-fit to the data. On the left, the colour coded contour shows the
+Figure \ref{fig:pred} shows predictions from the two models. 
+On the left, the colour coded contour shows the
 prediction of the model for combinations of cleaner abundance (x axis)
@@
 location with a combination of cleaner abundance and visitor leaving
-probability. Panels on the right show how close the observed values (y
+probability. Panels on the right show how close the observed values for 
+the probability of choosing the visitor (y
 axis) are from the predictions of the models (x axis). The line
@@
 Predicted values were taken using the mode just as in the left panels.
-On the lower right corner of the panel we show the coefficient of
-determination. Despite having one more parameter varying, the full model
-gives a lower fit to the data than the model without negative reward,
-measured by the coefficient of determination. In the supplementary
+On the lower right corner of the panel we show the $pseudo-R^2$ for predictions
+using the mode of the posterior distributions. 
+Despite having one more parameter varying, the full model
+gives a lower fit to the data than the model without negative reward. 
+The difference in $pseudo-R^2$ for the point estimates is small, but using the whole posterior distribution
+shows the improvement in fit achieved by eliminating negative reward 
+from the model (Fig. \ref{fig:post}f). In the supplementary
 material, we also show that a model with only negative reward, excluding
@@
 
-*Insert a nice discussion here*
+We have shown that using future rewards in the learning process is the cognitive 
+strategy that best explains the intra-specific variation in cleaner performance 
+in the market task. We compared future reward with the use of client behaviour as 
+a primary source of negative reward, and a combination of both. All three models 
+capture well the positive relation between visitor leaving behaviour and cleaner
+performance in the market task [@triki_Biological_2019]. For both mechanisms it is 
+indeed necessary to have fleeing behaviour on the side of the visitors for cleaners
+to develop a preference for them. The key difference between future and negative 
+reward emerges in the predicted relation between cleaner relative abundance and 
+cleaner performance. When cleaner relative abundance is low a model with only 
+future reward predicts low performance in the market task regardless of the visitor
+leaving probability. In contrast, A model with negative reward predicts the highest 
+performance with low relative cleaner abundance, particularly together with high 
+probability of visitor leaving (Fig. \ref{fig:pred}). Low cleaner abundances 
+mean the market has an excess of demand for cleaning services. In the model,
+this translates to a cleaning station that is frequently full. Thus, when visitors 
+leave, it's likely that the cleaner will have access to another client, and there won't be much difference in future reward between choosing a visitor and a resident. Thus,
+estimation of future reward will not develop a preference for the visitor in this 
+conditions, because it matches the state of the market.
+On the other hand, the effect of negative reward is unaffected by this market situation.
+In a full station, cleaners will get more often the heterotypic (resident-visitor) option and will develop a preference for the visitor faster.
+Estimation of future reward is then the cognitive mechanism
+that allows cleaner to adaptively respond to the ecological conditions of the biological market.
+
+<!-- ### Correlation of visitor leaving and cleaner density  -->
+
+<!-- ### Cognition as an explanation -->
+
+<!-- bash on social competence -->
+
+<!-- Cognitive processes gain support as an explanation for inter and intra-specific  -->
+<!-- variation in the market task. Even though the model considered here does not  -->
+<!-- explicitly model perception, associative processes must include the perception  -->
+<!-- of different stimuli which allow cleaners to discriminate -->
+<!-- residents and visitors. Thus, the fact that recent experimental work shows -->
+<!-- an effect of the perceptual conditions in the experiments on the performance of  -->
+<!-- the subjects, supports the idea that learning is involved in the decision  -->
+<!-- making process of the market task. Furthermore, and notably, certain features  -->
+<!-- of brain anatomy correlate with performance in the market task and with the  -->
+<!-- abundance of cleaners in the location where the fish were collected [@triki_Brain_2020].  -->
+<!-- All this, supports the idea that learning processes are behind the behavioural  -->
+<!-- adjustment that cleaners show to environmental conditions.  -->
 

### 2021-11-16 analysisABC.R

@@ -19,48 +19,42 @@ library("jsonlite")
 # simdir<-"D:/Neuch_simulations/ABCfit/Simulations/"
 
 scen1<-"MCMCclean_gam_Nrew_sca_"
-scen2<-"ABCclean_gam_sca_"
+scen2<-"MCMCclean_gam_sca_"
+scen3<-"MCMCclean_nRew_sca_"
 
 ## Load files --------------------------------------------------------------
 
-# (listFiles<-list.files(paste0(simdir,scen),full.names = TRUE))
-(listFiles1<-list.files(here("Simulations",scen1)))
-(ABCruns1<-grep("MCMCchain",listFiles1,value = TRUE))
+# all chains in a single object
 
-(listFiles2<-list.files(here("Simulations",scen2)))
-(ABCruns2<-grep("MCMCchain",listFiles2,value = TRUE))
-
-
-# all in a single object
-ABCraw.1<-do.call(rbind,lapply(ABCruns1, function(file){
-  rundata<-fread(here("Simulations",scen1,file))
-  seedNum<-strsplit(file,"_")[[1]][3]
-  seedNum<-as.numeric(gsub("[[:alpha:]]",seedNum,replacement = ''))
-  rundata[,seed:=rep(seedNum,dim(rundata)[1])]
-}))
-
-ABCraw.2<-do.call(rbind,lapply(ABCruns2, function(file){
-  rundata<-fread(here("Simulations",scen2,file))
-  seedNum<-strsplit(file,"_")[[1]][3]
-  seedNum<-as.numeric(gsub("[[:alpha:]]",seedNum,replacement = ''))
-  rundata[,seed:=rep(seedNum,dim(rundata)[1])]
-}))
-
+MCMCdata1<-loadMCMCrun(scen1)
+MCMCdata2<-loadMCMCrun(scen2)
+MCMCdata3<-loadMCMCrun(scen3)
 
 # One per onject
 ABCraw<-fread(here("Simulations",scen,ABCruns[3]))
 
+# Defaults
+burn.in<-1000
+thinning<-100
+
 
 # All in a list
-mcmcList.1<-mcmc.list(do.call(list,lapply(ABCruns1, function(file){
-  rundata<-fread(here("Simulations",scen1,file))
+mcmcList.1<-mcmc.list(do.call(list,lapply(unique(MCMCdata1$seed), function(repli){
+  rundata<-MCMCdata1[seed==repli]
[…]

### 2021-11-16 loadData.R

@@ -220,7 +220,7 @@ loadMCMCrun<-function(scen,burn.in=1000,thinning=100){
     seedNum<-as.numeric(gsub("[[:alpha:]]",seedNum,replacement = ''))
     rundata[,seed:=rep(seedNum,dim(rundata)[1])]
   }))
-  filtered<-raw[iteration>burn.in & iteration %% thinning==0]
+  filtered<-raw[iteration>burn.in & iteration %% thinning==0,]
   return(filtered)
 }
   

### 2021-11-16 loadFieldData.R

@@ -18,6 +18,14 @@ fieldData.site<-fieldData[,lapply(.SD,max),by=site_year,
                           .SDcols=c("abundance_large_100m2","abundance_small_100m2",
                                    "abundance_cleaners_100m2","percentage_swim_off")]
 
+fieldData.site %>%
+  ggplot(aes(y=percentage_swim_off,x=abundance_cleaners_100m2))+
+  geom_point()+geom_smooth(method=lm)+
+  theme_classic()
+model.ml<-lm(data = fieldData.site,percentage_swim_off~abundance_cleaners_100m2)
+anova(model.ml)
+summary(model.ml)
+
 fieldData.cleaner<-read_xlsx(here("Data","market_raw_data.xlsx"),sheet = "round_data")
 
 fieldData.cleaner<-data.table(fieldData.cleaner)

### 2021-11-16 manFigures.R

@@ -15,8 +15,8 @@ source(here("..","R_files",'ternaryAEQP.R'))
 source(here("../R_files/posPlots.R"))
 
 
-scenario<-"ABCclean_gam_nRew_sca_2"
-scenario2<-"ABCclean_gam_sca"
+scenario<-"MCMCclean_gam_Nrew_sca"
+scenario2<-"MCMCclean_gam_sca"
 
 predfileMode.both<-grep(".txt",grep("round",list.files(here("Simulations",paste0(scenario,"_"))),value = T),
                     value = T)
@@ -82,6 +82,7 @@ fieldatabyLoc.gam
 myPalette <- colorRampPalette(rev(brewer.pal(11, "Spectral")))
 sc <- scale_colour_gradientn(colours = myPalette(100), limits=c(0,0.8))
 
+typeof(sc)
 
 obs.both<-ggplot(data = fieldatabyLoc.both, 
             aes(x=re.abund.clean, y=prob.Vis.leave,color=probVisi.data))+
@@ -246,15 +247,6 @@ names(FIAinterpData.gam)<-c("rel.abund.cleaners","prob.Vis.Leav","market_binomia
 names(fieldatabyLoc.gam)[c(4,5,2)]<-c("rel.abund.cleaners","prob.Vis.Leav","market_binomial_data")
 
 
-
-fieldatabyLoc.both[,resids:=market_binomial_data-probvisitor.pred]
-fieldatabyLoc.gam[,resids:=market_binomial_data-probvisitor.pred]
-
-fieldatabyLocSamps.both[,resids:=probVisi.data-probvisitor.pred]
-fieldatabyLocSamps.gam[,resids:=probVisi.data-probvisitor.pred]
-
-
-
 predictDataMode.both[,
                                          log.like:=dbinom(visitorChoices,
                                 size = 20,prob = visitorChoices_pred,log = TRUE)]
@@ -286,15 +278,15 @@ axis.title.x = element_text(size=12),axis.title.y = element_text(size=12))
 scatter.obs.pred.both<-ggplot(data = fieldatabyLoc.both,aes(y=market_binomial_data,x=probvisitor.pred))+
   geom_point(data = fieldatabyLocSamps.both,aes(x=probvisitor.pred,y=probVisi.data),
              color='grey')+
-  geom_point(size=4)+ylim(0.45,0.8)+xlim(0.45,0.8)+
[…]

### 2021-11-16 parTOjson.R

@@ -16,9 +16,16 @@ scenario<-"ABCclean_nRew_sca"
 
 # For predictions on the observed values ---------------------------------------
 
-scenario<-"ABCclean_nRew_sca"
+scenario<-"MCMCclean_gam_sca"
 
 MCMCdata<-loadMCMCrun(paste0(scenario,"_"))
+densGamma<-density(MCMCdata$gamma)
+modeGamma<-densGamma$x[densGamma$y==max(densGamma$y)]
+densNR<-density(MCMCdata$negReward)
+modeNegrew<-densNR$x[densNR$y==max(densNR$y)]
+densScal<-density(MCMCdata$scaleConst)
+modescal<-densScal$x[densScal$y==max(densScal$y)]
+
 
 param_pred<-list(totRounds=10000,ResReward=1,VisReward=1,
             ResProbLeav=0,scenario=0, inbr=0,outbr=0,forRat=0.0,
@@ -28,15 +35,15 @@ param_pred<-list(totRounds=10000,ResReward=1,VisReward=1,
             pertScen = c(FALSE,FALSE,FALSE,FALSE,TRUE), 	
             MCMC =0, data="clean",nRep=30,
             dataFile = here("Data","data_ABC_cleaner_absolute.txt"),
-            folderL=paste0(here(simsDir),"/",scenario,"_/","samplesPost_/"))
-
+            # folderL=paste0(here(simsDir),"/",scenario,"_/","samplesPost_/"))
+            folderL=paste0(here(simsDir),"/",scenario,"_/"))
 
 
 check_create.dir(here(simsDir,paste0(scenario,"_")),param = rep("samplesPost",1),
                  values = c(""))
 
 param_pred$folder<-param_pred$folderL
-# param_pred$init[c(3,4,5)]<- c(modeGam,modenegReward,modeScal)
+# param_pred$init[c(3,4,5)]<- c(modeGamma,0,modeScal)
   # c(runif(1,max = 1),runif(1,max = 1),
   #                            runif(1,max = 500,min = 1))
   #sumMCMC$statistics[,1]
@@ -46,7 +53,7 @@ postSamp<-MCMCdata[round(runif(nsamples,min = 1,max=dim(MCMCdata)[1]),
                               digits = 0),]
 
 for(i in 1:nsamples){
-  param_pred$init[c(3,4,5)]<- postSamp[i,.(gamma,negReward,scaleConst)]
[…]

### 2021-11-18 manuscript_1.0.Rmd
Finished writting the first version of the discussion. Abstract still to be written

@@
 author:
-- Andrés E. Quiñones^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
-  - andresqp@gmail.com]
+- Andrés E. Quiñones^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland - andresqp@gmail.com]
+- Zegni Triki^[Department of Zoology, Stockholm University, Stockholm, Sweden]
+- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 of the parameter values of the full model, which includes both future
-and negative reward. Notably, the bulk of the posterior distribution for
-the parameter tuning negative reward ($\eta$ panerl e) is around zero.
+and negative reward. Posterior distributions show the probability density (y axis)
+of the estimated parameter values (x axis). Notably, the bulk of the posterior distribution for
+the parameter tuning negative reward ($\eta$ panel e) is around zero.
 Thus, the shape of the posterior and its credible intervals suggest the absence of negative reward to be well supported by the data.
@@
 two parameters estimated. Not surprisingly, under a model without
-negative reward, the distribution of $\gamma$ shifts to higher values. Under the new model, the estimates of $\gamma$ do not include zero, the distribution is centered arounf $0.5$.
+negative reward, the distribution of $\gamma$ shifts to higher values. Under the new model, the estimates of $\gamma$ do not include zero, the distribution is centered around $0.5$.
 To asses the fit of the two models we calculated the  $pseudo-R^2$ [@mcfadden_Conditional_1974], which is a standard measure of fit for logistic regression. 
@@
 strategy that best explains the intra-specific variation in cleaner performance 
-in the market task. We compared future reward with the use of client behaviour as 
+in the market task. We compared future reward with the use of visitors behaviour as 
 a primary source of negative reward, and a combination of both. All three models 
@@
 future reward predicts low performance in the market task regardless of the visitor
-leaving probability. In contrast, A model with negative reward predicts the highest 
+leaving probability. In contrast, a model with negative reward predicts the highest 
 performance with low relative cleaner abundance, particularly together with high 
@@
 this translates to a cleaning station that is frequently full. Thus, when visitors 
-leave, it's likely that the cleaner will have access to another client, and there won't be much difference in future reward between choosing a visitor and a resident. Thus,
+leave, it's likely that the cleaner will have access to another client, and there will not be much difference in future reward between choosing a visitor and a resident. Thus,
 estimation of future reward will not develop a preference for the visitor in this 
@@
 In a full station, cleaners will get more often the heterotypic (resident-visitor) option and will develop a preference for the visitor faster.
-Estimation of future reward is then the cognitive mechanism
-that allows cleaner to adaptively respond to the ecological conditions of the biological market.
+Note that when cleaner abundance is high situations with a resident and a visitor will be so rare that neither mechanism will be very efficient in generating a preference for visitors.
+When such choice arises it is still be best to choose the visitor; however, the learning machinery will not be able to develop this preference efficiently.
+Overall, estimation of future reward is then the cognitive mechanism that allows cleaners to adaptively respond conditions to the ecological conditions of the biological market.
 
-<!-- ### Correlation of visitor leaving and cleaner density  -->
+The cognitive mechanism that our analysis highlights does not require a direct flow of information between the cleaner and its social environment.
+The adaptive responses to local conditions in cleaners is not only evidenced by their performance in  experiments, but also by changes in their brain anatomy.
+Indeed, individual cleaners that performed in the laboratory test according to what is best at the location in which they were caught had larger forebrains harbouring more cells than individuals whose performance mismatched locally optimal decision-rules [@triki_Brain_2020].
+This suggests that despite the overall behavioural adjustment, evidenced here and elsewhere
+[@triki_Biological_2019; @triki_Decrease_2018], there are some cleaners better than others at adjusting their behaviour.
+@triki_Brain_2020 refer to these as socially competent cleaners.
+Social competence refers to the ability to optimise social behaviour depending on the available social information [@taborsky_Social_2012a;@bshary_Cooperation_2015].
+However, we have shown here that cleaners are able to adjust their preference 
+according to the social environment without using any direct information about the social
+environment.
+Instead cleaners seem to adjust their preference using information from the visitation patterns of their clients.
+These patters are of course determined by the feedback with the cleaner's social environment (relative abundance).
+Our models, however, do not include the possibility of cleaners responding to their relative abundance directly.
+The mechanism we present is a general way to learn from the environment
+in both asocial and social situations. It would be interesting to contrast this learning mechanism with one tailored to a social situation, and see whether such mechanism 
+fits the data better. More generally, an important challenge ahead is to understand 
+which cognitive mechanisms evolved exclusively for social life.
 
-<!-- ### Cognition as an explanation -->
+One potential difference between cleaners that adjusted adaptively, like the ones
+reported in @triki_Brain_2020, is configurational learning or chunking. The original analysis of the learning model showed that agents needed two mechanism to solve the 
+market task; to account for the future, and to assign value to the combination of clients
+present in the station [@quinones_Reinforcement_2019]. The latter is referred to as configurational learning or chunking [@sutherland_Configural_1989;@miller_Magical_1956]. 
+Theoretical models suggest that the lack of configurational learning is related to a low 
+frequency with which stimuli are encountered together [@kolodny_Evolution_2014;@prat_Role_2021].
+Thus, cleaner relative abundance could be related to this mechanisms as well.
+Given that the lack of either of these two mechanisms yields
+neutral preferences [@quinones_Reinforcement_2019], we can't, based on the current analysis, distinguish between the two possibilities.
+One potential way to distinguish between the two would be to use an experimental set up designed to test configuration learning on cleaners
+with high and low performance in the market task [@woodbury_Learning_1943;@sutherland_Configural_1989].
 
-<!-- bash on social competence -->
-
-<!-- Cognitive processes gain support as an explanation for inter and intra-specific  -->
-<!-- variation in the market task. Even though the model considered here does not  -->
-<!-- explicitly model perception, associative processes must include the perception  -->
-<!-- of different stimuli which allow cleaners to discriminate -->
-<!-- residents and visitors. Thus, the fact that recent experimental work shows -->
-<!-- an effect of the perceptual conditions in the experiments on the performance of  -->
-<!-- the subjects, supports the idea that learning is involved in the decision  -->
-<!-- making process of the market task. Furthermore, and notably, certain features  -->
-<!-- of brain anatomy correlate with performance in the market task and with the  -->
-<!-- abundance of cleaners in the location where the fish were collected [@triki_Brain_2020].  -->
-<!-- All this, supports the idea that learning processes are behind the behavioural  -->
-<!-- adjustment that cleaners show to environmental conditions.  -->
+Combining mechanistic modelling with empirical data provides a powerful tool to identify the cognitive processes underlying learning in animals.
+The combination overcomes the limitations of the general philosophy in animal cognition to apply the logic of Lloyd Morgen’s canon (Occam’s razor).
+Excluding basic reinforcement learning explanations (operant and/or classical conditioning) as a potential explanation in experiments often warrants animals solving the task on the first possible occasion.
+For example, any theory of mind task needs to be solved in the first trial in order to exclude fast conditioning [@heyes_Theory_1998].
+Similarly, subjects need to solve a social learning task on the first trial to accept imitation as a mechanism over stimulus/local enhancement.
+Such strict conditions are virtually never met.
+For example, potato washing by Japanese macaques, an iconic example of social learning, took several years to spread within the group [@kawamura_Process_1959], meaning that any learner had been repeatedly exposed to demonstrations before acquisition.
+Importantly, @galef_Question_1992 refuted imitation as mechanism not simply because of the repeated exposure but because a (rather qualitative) analysis of the spread of potato washing across individuals did not follow the prediction based on imitation learning (see also @hirata_SweetPotato_2001).
+In our case, the number of trials it took cleaners to learn the solution to the market task would never allow to exclude an important role of negative reinforcement based on the data alone.
+However, fitting model predictions to our data set revealed that a more complex mechanism (estimation of future reward) fits the data better.
 

### 2021-11-18 loadFieldData.R

@@ -1,7 +1,8 @@
 library("data.table")
 library("here")
 library("readxl")
-
+library('dplyr')
+library('ggplot2')
 
 ## Load field data based on the market experiment performance criteria
 
@@ -75,6 +76,16 @@ fieldData.sum[,site.year:=gsub(" ",replacement = "_",x = site.year)]
 fieldData.sum[,cleaner_ID:=gsub(" ",replacement = "",x = cleaner_ID)]
 
 str(fieldData.sum)
+
+fieldData.sum %>%
+  ggplot(aes(x=abund.cleaners,y=score_visitor))+
+  geom_point()+theme_classic()
+
+
+fieldData.site %>%
+  ggplot(aes(x=abundance_cleaners_100m2,fill=score_visitor,y=prob.Vis.Leav))+
+  geom_point()+theme_classic()
+
 
 setorder(fieldData.sum,site.year)
 

### 2021-11-23 manuscript_1.0.Rmd
include a table with the parameter values.

@@
 output:
+  pdf_document: default
   word_document: default
-  pdf_document: default
 fig_caption: yes
@@
 combination of these effects is a better explanation for the pattern
-seen in the data.
+seen in the data. For other parameter values used for the model simulations
+and the MCMC see Table \ref{tab:param}
+
+<!-- # ```{r parTable} -->
+<!-- # source(here("loadData.R")) -->
+<!-- # parFiles<-grep("parametersMCMC",list.files(here("Simulations","MCMCclean_gam_nRew_sca_"), -->
+<!-- #                                               full.names = TRUE),value = TRUE) -->
+<!-- # params<-fromJSON(parFiles[1]) -->
+<!-- # ``` -->
+
+```{r param , echo=FALSE,results='asis'}
+  cat(' Table: \\label{tab:param} Parameter values with which the model was run 
+  in the MCMC.
+$\\sigma$ refers to the amplitud of the perturbation kernel with the subscript indicating the associated parameter.
+New values were taken from a uniform distribution.
+$\\alpha$ refers to the learning rate.
+
+  | Parameter | Value |
+|----------:|:-----:|
+| Learning rounds | 10000 |
+| Reward value | 1 |
+| $\\alpha$ | 0.05 |
+| $\\sigma_{\\gamma}$ | 0.3 |
+| $\\sigma_{\\eta}$ | 4 |
+| $\\sigma_{Sca.Const.}$ | 300 |
+| Number of chains | 5 |
+| Chain lenght | $1^5$ |
+')
+
+```
 
@@
 Line correspond to a perfect match between observed and predicted probabilities.",out.width='90%'}
-fileNogamma<-grep("contour_gamma_",list.files(here("Simulations","ABCclean_nRew_sca_"),
+fileNogamma<-grep("contour_gamma_",list.files(here("Simulations","MCMCclean_nRew_sca_"),
                                               full.names = TRUE),value = TRUE)

### 2021-11-25 ActCritABC.cpp
First version of a model where cleaners are split in two groups (competent and incompetent). Each group has a different gamma and eta values.

@@ -614,17 +614,23 @@ struct cleaner_point {
 	int countVisitor, totalMarket;
 	float market_exp_success;
 	double marketPred;
+	int group=0;
 };
+
+
 
 struct model_param {
 	//model_param(model_param const &obj);
-	double alphaC, alphaA, gamma, negReward, scaleConst;
+	double alphaC, alphaA, scaleConst;
+	double gamma[2], negReward[2];
 	model_param &operator= (model_param const &rhs) {
 		alphaA = rhs.alphaA;
 		alphaC = rhs.alphaC;
-		gamma = rhs.gamma;
-		negReward = rhs.negReward;
 		scaleConst = rhs.scaleConst;
+		gamma[0] = rhs.gamma[0];
+		negReward[0] = rhs.negReward[0];
+		gamma[1] = rhs.gamma[1];
+		negReward[1] = rhs.negReward[1];
 		return *this;
 	}
 };
@@ -682,6 +688,7 @@ vector <cleaner_point> read_CleanData(ifstream& marketData) {
 		marketData >> input.abund_visitors;
 		marketData >> input.abund_resid;
 		marketData >> input.prob_Vis_Leav;
+		marketData >> input.group;
 		if (marketData.eof()) 	break;
 		input.totalMarket = 20;
 		input.market_exp_success = double(input.countVisitor) /
@@ -737,7 +744,10 @@ void initializeChainFile(ofstream &chainOutput,nlohmann::json param){
 	string IndFile = create_filename(namedir, param);
 	chainOutput.open(IndFile.c_str());
 	chainOutput << "iteration	" << "alpha_actor	" << "alpha_critic	" <<
-		"gamma	" << "negReward	" << "scaleConst	" << "fit	" << "ratio" << endl;
+		 "gamma	" << "negReward	";
+	if (bool(param["Group"]))
+		chainOutput	<< "gamma.1	" << "negReward.1	";
+	chainOutput  << "scaleConst	" << "fit	" << "ratio" << endl;
 }
 
 void initializeRoundFile(ofstream& RoundOutput, nlohmann::json param,
@@ -746,9 +756,9 @@ void initializeRoundFile(ofstream& RoundOutput, nlohmann::json param,
 	std::string namedir = param["folder"];
 	namedir.append("round_");
 	namedir.append("gamma_");
[…]

### 2021-11-25 parTOjson.R

@@ -10,7 +10,7 @@ exedir<-'/./ActCrit.exe'
 
 fileName<-"parameters.json"
 
-scenario<-"ABCclean_nRew_sca"
+scenario<-"MCMCclean_gam_nRew_sca_group_"
 
 
 
@@ -168,7 +168,7 @@ for (k in 1:90) print(y)
 
 # MCMC fit - Generate json parameter files for -------------------------------------
 
-scenario<-"MCMCclean_Nrew_sca"
+scenario<-"MCMCclean_gam_nRew_sca_group"
 
 filesList<-list.files(here("Simulations",scenario),full.names = TRUE)
 
@@ -197,9 +197,10 @@ param_mcmc<-list(totRounds=10000,ResReward=1,VisReward=1,
                 seed=1, propfullPrint = 0.7,sdPert=c(0.05,0.05,0.3,4,300),
                 chain_length=100000,
                 init=c(0.05,0.05,0,0,30),# alphaA,AlphaC, Gamma, NegRew
-                pertScen = c(FALSE,FALSE,FALSE,TRUE,TRUE), 
+                pertScen = c(FALSE,FALSE,TRUE,TRUE,TRUE), 
                 MCMC =1, data="clean",nRep=1,
-                  dataFile = here("Data","data_ABC_cleaner_absolute.txt"),
+                Group = TRUE,
+                dataFile = here("Data","data_ABC_cleaner_abs_threa1.5.txt"),
                   # here("Simulations",
                   #                 paste0(scenario,"_"),"data_MCMC_fake.txt"))
                 ##here("Data","data_ABC_site_20.txt")
@@ -211,7 +212,7 @@ check_create.dir(here(simsDir),param = rep(scenario,1),
 
 for(seed in 1:5){
   param_mcmc$folder<-param_mcmc$folderL
-  param_mcmc$init<-c(0.05,0.05,0,runif(1,max = 2,min = 0),#runif(1,max = 0.6),0
+  param_mcmc$init<-c(0.05,0.05,runif(1,max = 0.6),runif(1,max = 2,min = 0),#,0
                      runif(1,max = 300,min = 5))
                      ##runif(1,max = 50,min = 5))
   param_mcmc$seed <- seed

### 2021-11-26 manuscript_1.0.Rmd
Included some of Red1 comments. Updated figure 1.

@@
 [@begon_Ecology_2006]. When partners in these mutualistic interactions
-are able to choose their interaction partner, their investment and the return 
-they obtain can be understood as a 
-biological market [@hammerstein_Biological_2016; @leimar_Cooperation_2010;
+are able to choose their interaction partner, their investment and the
+return they obtain can be understood as a biological market
+[@hammerstein_Biological_2016; @leimar_Cooperation_2010;
 @noe_Biological_1994a,@noe_Biological_1995a]. Thus, we expect supply
-and demand of the good 
-and services being exchanged to determine the action of individuals, at least in
-an evolutionary time-scale.
-However, supply and demand conditions of a biological market, which typically
-depend on the abundance of the species involved, certainly vary in time and space. 
-How do partners adjust their strategy in the face of this variation? 
-In animals an obvious general candidate mechanism for the strategic adjustment is the 
-cognitive machinery. However, it is not clear which specific cognitive mechanisms
-allow individual to adjust their behaviour to the varying market conditions.
+and demand of the good and services being exchanged to determine the
+action of individuals, at least in an evolutionary time-scale. However,
+supply and demand conditions of a biological market, which typically
+depend on the abundance of the species involved, certainly vary in time
+and space. How do partners adjust their strategy in the face of this
+variation? In animals an obvious general candidate mechanism for the
+strategic adjustment is the cognitive machinery. However, it is not
+clear which specific cognitive mechanisms allow individual to adjust
+their behaviour to the varying market conditions.
 
-One example of strategic adjustment in a biological market is the marine cleaning 
-mutualism involving the cleaner
-fish *Labroides dimidiatus* and 'client' fish. Clients offer themselves as
-food patches for cleaners to remove ectoparasites, which gives the
-cleaners calories and the clients improved health
-[@waldie_LongTerm_2011; @ros_Does_2011,@triki_Effects_2016]. As
-cleaners have about 2000 interactions per day
-[@grutter_Relationship_1995], they regularly face the situation that
-two clients seek cleaning simultaneously. In such conditions, it is
-important to distinguish between 'resident' and 'visitor' client species
-[@bshary_Cleaner_2001a]. The residents have small home ranges and hence
-access to only the local cleaner, while the visitors have large home
-ranges and hence access to two or more cleaners. Thus, visitors have the
-option to switch to another cleaner if being made to wait, while
-residents must wait for inspection. Indeed, visitors have been observed
-to use their partner choice options in that way [@bshary_Choosy_2002],
-which may explain why cleaners give visitors service priority in a field
-study in the Red Sea [@bshary_Cleaner_2001a]. Furthermore, early
-laboratory experiments showed that leaving causes service priority
-[@bshary_Asymmetric_2002]. The experiments consisted of a simultaneous
-two-choice task, where cleaners had to decide which of two Plexiglas
-plates to approach first and eat mashed prawn of it. One plate was made
-to behave like a resident, i.e. it would remain until the cleaner had
-eaten off it. The other plate was made to behave like a visitor, i.e. it
-would be removed if not inspected first. Cleaners learned to prefer the
-visitor plate and hence obtained the double amount of food
-[@bshary_Asymmetric_2002]. Adult cleaner fish outperform various
-primates as well as rats and pigeons in this original version of the
-biological market task [@salwiczek_Adult_2012; @zentall_Early_2017];
-only African grey parrots solve this task as well [@pepperberg_Can_2014].
+One example of strategic adjustment in a biological market is the marine
+cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
+'client' fish. Clients offer themselves as food patches for cleaners to
+remove ectoparasites, which gives the cleaners calories and the clients
+improved health [@waldie_LongTerm_2011;
+@ros_Does_2011,@triki_Effects_2016]. As cleaners have about 2000
+interactions per day [@grutter_Relationship_1995], they regularly face
+the situation that two clients seek cleaning simultaneously. In such
+conditions, it is important to distinguish between 'resident' and
+'visitor' client species [@bshary_Cleaner_2001a]. The residents have
+small home ranges and hence access to only the local cleaner, while the
+visitors have large home ranges and hence access to two or more
+cleaners. Thus, visitors have the option to switch to another cleaner if
+being made to wait, while residents must wait for inspection. Indeed,
+visitors have been observed to use their partner choice options in that
+way [@bshary_Choosy_2002], which may explain why cleaners give
+visitors service priority in a field study in the Red Sea
+[@bshary_Cleaner_2001a]. More generally, cleaners apparently prefer to
+interact with visitors [@triki_Zeni_inpress]. Such a preference may be
+linked to a client's value as a food patch; visitors are on average
+larger than residents [ @bshary_Cleaner_2001a], and size is positively
+correlated with parasite load(Grutter xxxx). However, early laboratory
+experiments showed that leaving causes service priority independent of
+client value [@bshary_Asymmetric_2002]. The experiments consisted of a
+simultaneous two-choice task, where cleaners had to decide which of two
+Plexiglas plates, which offered equal amounts of food, to approach first
+and eat mashed prawn of it. One plate was made to behave like a
+resident, i.e. it would remain until the cleaner had eaten off it. The
+other plate was made to behave like a visitor, i.e. it would be removed
+if not inspected first. Cleaners learned to prefer the visitor plate and
+hence obtained the double amount of food [@bshary_Asymmetric_2002].
+Adult cleaner fish outperform various primates as well as rats and
+pigeons in this original version of the biological market task
+[@salwiczek_Adult_2012; @zentall_Early_2017]; only African grey
+parrots solve this task as well [@pepperberg_Can_2014].
 
@@
 presentation may matter; i.e. what relevant information is encoded in
-the design of the two plates (colour versus size; @wismer_Cuebased_2019), or
-whether the information is in the colour of the food
-[@pretot_Comparative_2021; @pretot_Comparing_2016], and whether the
-food is visible or hidden/released only when the subjects made choices
-on a touch screen [@pretot_Factors_2016]. Second, it appears that cleaner and
-client densities in nature affect how cleaners perform in the original
-market task: when densities are low, cleaners are more likely to fail to
-complete the task [@triki_Biological_2019; @triki_Decrease_2018;
-@wismer_Variation_2014]. The apparent explanation for this failure is that visitor
-clients are less likely to exert their choice options (and hence wait
-for inspection if made to wait) when cleaner fish densities are low
-[@triki_Brain_2020]. This change in behaviour may be due to the fact that
-when cleaner densities are low, the cost of switching in terms of
-distance to swim and time invested increases. If visitors are not
-choosy, cleaners should not give them priority anymore. Interestingly,
-while cleaners from high-density areas have larger forebrains harbouring
-more cells than cleaners from low-density areas [@triki_Brain_2020], on a
-local scale those individuals that perform in the market task in a way
-that corresponds to the locally optimal strategy (solving/failing the
-task when living in high/low density) have relatively larger forebrains
-[@triki_Brain_2020]. Taken together, recent empirical results clearly show
-that cognitive processes are involved in the pattern of intra-
-and inter-specific variation in  performance in the market task. One approach to
-identify candidate mechanisms underlying performance in a cognitive task
-is to develop proximate learning models and to evaluated in light of the behavioural data 
-(e.g. @dubois_Model_2021). By varying the cognitive tool
-kit, one can assess which minimal kit is necessary to solve the task at
-hand. This approach has been used to analyse the market task
-[@prat_Role_2021; @quinones_Reinforcement_2019]. Learning models showed that basic reinforcement learning does not suffice to solve the task, this is 
-particularly so when model assume the more complex natural situation 
-in which cleaners
-not only face resident-visitor pairs but also visitor-visitor and
-resident-resident pairs, as well only a resident or only a visitor. To
-be able to give visitors priority over residents, cleaners first need to
-be able to assess a client's value separately for the three possible
-forms of encounter (alone, paired with a fish with the same strategic
-option, paired with a fish with the alternative strategic option)
-[@quinones_Reinforcement_2019]. In addition, cleaners need to account for the future
-consequences of current decisions. In the model, this could be achieved
-in two non-mutually exclusive ways [@quinones_Reinforcement_2019]:
-through low temporal discounting of future effects
-and/or through perceiving a visitor client leaving as psychological
-punishment (i.e. as a negative reinforcer). Given that a negative
-reinforcer is part of the ubiquitous learning mechanism termed operant
-conditioning (Thorndike xxxx), this explanation can be considered the
-simpler one. Furthermore, accounting for the future effects of an action is 
-considered an advanced cognitive ability [@suddendorf_Evolution_2007]. 
-Here, we fitted our model to the empirical data on fish
-densities and cleaner fish performance in the market task with the aim
-to determine the likely mechanism used by cleaners to incorporate future
-consequences of current decisions.
+the design of the two plates (colour versus size;
+@wismer_Cuebased_2019), or whether the information is in the colour of
+the food [@pretot_Comparative_2021; @pretot_Comparing_2016], and
+whether the food is visible or hidden/released only when the subjects
+made choices on a touch screen [@pretot_Factors_2016]. Second, it
+appears that cleaner and client densities in nature affect how cleaners
+perform in the original market task: when densities are low, cleaners
+are more likely to fail to complete the task [@triki_Biological_2019;
+@triki_Decrease_2018; @wismer_Variation_2014]. The apparent explanation
+for this failure is that visitor clients are less likely to exert their
+choice options (and hence wait for inspection if made to wait) when
+cleaner fish densities are low [@triki_Brain_2020]. This change in
+behaviour may be due to the fact that when cleaner densities are low,
+the cost of switching in terms of distance to swim and time invested
+increases. If visitors are not choosy, cleaners should not give them
+priority anymore. Interestingly, while cleaners from high-density areas
+have larger forebrains harbouring more cells than cleaners from
+low-density areas [@triki_Brain_2020], on a local scale those
+individuals that perform in the market task in a way that corresponds to
+the locally optimal strategy (solving/failing the task when living in
+high/low density) have relatively larger forebrains [@triki_Brain_2020].
+Taken together, empirical results clearly show major intra- and
+inter-specific variation in performance in the market task, warranting
+investigation of the underlying cognitive machinery, from perception to
+decision-making [@shettleworth_Cognition_2009].
+
+One approach to identify candidate mechanisms underlying performance in
+a cognitive task is to develop proximate learning models and to evaluate
+their predictions in light of the behavioural data (e.g.
+@dubois_Model_2021). By varying the cognitive tool kit, one can assess
+which minimal kit is necessary to solve the task at hand. This approach
+has been used to analyse the market task [@prat_Role_2021;
+@quinones_Reinforcement_2019]. Learning models showed that basic
+reinforcement learning does not suffice to solve the task, this is
+particularly so when models assume the more complex natural situation in
+which cleaners not only face resident-visitor pairs but also
+visitor-visitor and resident-resident pairs, as well only a resident or
+only a visitor. To be able to give visitors priority over residents,
+cleaners first need to be able to assess a client's value separately for
+the three possible forms of encounter (alone, paired with a fish with
+the same strategic option, paired with a fish with the alternative
+strategic option) [@quinones_Reinforcement_2019]. In addition, cleaners
+need to account for the future consequences of current decisions. In the
+model, this could be achieved in two non-mutually exclusive ways
+[@quinones_Reinforcement_2019]: through low temporal discounting of
+future effects and/or through perceiving a visitor client leaving as
+psychological punishment (i.e. as a negative reinforcer). Given that a
+negative reinforcer is part of the ubiquitous learning mechanism termed
+operant conditioning (Thorndike 1917), this explanation can be
+considered the simpler one. Furthermore, accounting for the future
+effects of an action is considered an advanced cognitive ability
+[@suddendorf_Evolution_2007]. Here, we fitted our model to the empirical
+data on fish densities and cleaner fish performance in the market task
+with the aim to determine the likely mechanism used by cleaners to
+incorporate future consequences of current decisions.
 
@@
 
-We used the data from @triki_Biological_2019 to fit the parameters of the model analysed in
-@quinones_Reinforcement_2019. For the bayesian fitting procedure we used Markov Chain Monte Carlo (MCMC) with a Metropolis sampler to estimate the posterior
-distribution of the model parameters.  The data consist a of the number of times
-that a set of cleaners chose to feed of a visitor client plate in the
-context of the market task. This set of cleaners were collected in
-different locations and in each one of those locations environmental
-variables were measured. These environmental variables recorded from
-each location are used as input for the computational model. Given these
-inputs, the computational model produces a time series of choices in the
-market task. In the model the preference for the two options changes as
-the algorithm learns based on the reward that the two options provide.
-Thus, we run the model for 10000 steps and use the last 3000 steps as
-the preference developed in the model for visitors given the
-environmental variables. We use this preference as the predicted
-probability in a binomial distribution to calculate
-the likelihood that the number of visitor choices obtained from each
-cleaner is generated by the model. We sum over all the cleaners the
-log-likehoods and use the total for the acceptance of new parameters in
-the MCMC.
+We used the data from @triki_Biological_2019 to fit the parameters of
+the model analysed in @quinones_Reinforcement_2019. For the bayesian
+fitting procedure we used Markov Chain Monte Carlo (MCMC) with a
+Metropolis sampler to estimate the posterior distribution of the model
+parameters. The data consist a of the number of times that a set of
+cleaners chose to feed of a visitor client plate in the context of the
+market task. This set of cleaners were collected in different locations
+and in each one of those locations environmental variables were
+measured. These environmental variables recorded from each location are
+used as input for the computational model. Given these inputs, the
+computational model produces a time series of choices in the market
+task. In the model the preference for the two options changes as the
+algorithm learns based on the reward that the two options provide. Thus,
+we run the model for 10000 steps and use the last 3000 steps as the
+preference developed in the model for visitors given the environmental
+variables. We use this preference as the predicted probability in a
+binomial distribution to calculate the likelihood that the number of
+visitor choices obtained from each cleaner is generated by the model. We
+sum over all the cleaners the log-likehoods and use the total for the
+acceptance of new parameters in the MCMC.
 
@@
 the queue (with a certain probability) after they are not given
-priority. Individuals obtain reward from cleaning a client regardless of
-the type. Every time individuals face and make a choice they update the
-probability of making that same choice. The update is based on the
-difference between the expectation of reward and the actual reward the
-individual obtains (prediction error); and it is carried in the
-direction that would lead to more reward being obtained, given the new
-information. Further details of the model implementation can be found in
-the paper [@quinones_Reinforcement_2019]. The model analysis points to
-the relevance of two parameters. First, $\gamma$ measures how much
-individuals include future rewards in their decision updates. If
+priority. Individuals obtain a fixed reward from cleaning a client
+regardless of the type. Every time individuals face and make a choice
+they update the probability of making that same choice. The update is
+based on the difference between the expectation of reward and the actual
+reward the individual obtains (prediction error)
+[@sutton_Reinforcement_2018; \@rescola_Theory_1972]; and it is carried
+in the direction that would lead to more reward being obtained, given
+the new information. Further details of the model implementation can be
+found in the paper [@quinones_Reinforcement_2019]. The model analysis
+points to the relevance of two parameters. First, $\gamma$ measures how
+much individuals include future rewards in their decision updates. If
 $\gamma=0$, individuals only use the immediate reward obtained from a
@@
 reward obtained from the subsequent choices. That amounts to calculating
-and using for decision making the future expected rewards in action. Second,
-$\eta$ measures how much individuals include in their reward the fleeing
-behaviour of visitor clients as a negative component. Both of these
-parameters allow individuals to use in their estimates the future
-effects of their choices. The model analysis shows that without either
-of them, learning individuals cannot develop a preference for the
-visitor client. The analysis contrast the effect of these two parameters
-against variation in the relative abundance of the two client types. The
-analysis shows that under very high client (relative) abundance the presence of
-future rewards ($\gamma>0$) no longer favours the preference for the
-ephemeral clients, this effect does not occur with negative reward
-mediated by client behaviour ($\eta$). Thus, the analysis suggests that
-future reward as an explanation for cleaner ability to prefer visitors
-is supported by @triki_Biological_2019 data. However, a combination of
+and using for decision making the future expected rewards in action.
+Second, $\eta$ measures how much individuals include in their reward the
+fleeing behaviour of visitor clients as a negative component. Both of
+these parameters allow individuals to use in their estimates the future
+effects of their choices. The model analysis by
+@quinones_Reinforcement_2019 shows that without either of them, learning
+individuals cannot develop a preference for the visitor client. The
+analysis contrast the effect of these two parameters against variation
+in the relative abundance of the two client types. The analysis shows
+that under very high client (relative) abundance the presence of future
+rewards ($\gamma>0$) no longer favours the preference for the ephemeral
+clients, this effect does not occur with negative reward mediated by
+client behaviour ($\eta$). Thus, the analysis suggests that future
+reward as an explanation for cleaner ability to prefer visitors is
+supported by @triki_Biological_2019 data. However, a combination of
 future and negative rewards, could still be a better explanation for the
@@
 combination of these effects is a better explanation for the pattern
-seen in the data. For other parameter values used for the model simulations
-and the MCMC see Table \ref{tab:param}
-
-<!-- # ```{r parTable} -->
-<!-- # source(here("loadData.R")) -->
-<!-- # parFiles<-grep("parametersMCMC",list.files(here("Simulations","MCMCclean_gam_nRew_sca_"), -->
-<!-- #                                               full.names = TRUE),value = TRUE) -->
-<!-- # params<-fromJSON(parFiles[1]) -->
-<!-- # ``` -->
-
-```{r param , echo=FALSE,results='asis'}
-  cat(' Table: \\label{tab:param} Parameter values with which the model was run 
-  in the MCMC.
-$\\sigma$ refers to the amplitud of the perturbation kernel with the subscript indicating the associated parameter.
-New values were taken from a uniform distribution.
-$\\alpha$ refers to the learning rate.
-
-  | Parameter | Value |
-|----------:|:-----:|
-| Learning rounds | 10000 |
-| Reward value | 1 |
-| $\\alpha$ | 0.05 |
-| $\\sigma_{\\gamma}$ | 0.3 |
-| $\\sigma_{\\eta}$ | 4 |
-| $\\sigma_{Sca.Const.}$ | 300 |
-| Number of chains | 5 |
-| Chain lenght | $1^5$ |
-')
-
-```
+seen in the data. For other parameter values used for the model
+simulations and the MCMC see Table \ref{tab:param}
 
@@
 but also how often the cleaning station is empty (e.g. there are no
-clients to be cleaned). The more cleaners, the more options available
-for the clients and hence higher probability of empty cleaning stations.
-How often the cleaning stations are empty in turns influences the value
-that cleaners can expect from their choices. The more empty the empty
-cleaning stations are, the more it is to choose a client that will not
-stay. However, the frequency of empty spots in the cleaning station does
-not only depend on cleaner relative abundances but also on the frequency
-with which clients visit the station. The client visitation frequency
-will likely vary among different clients depending on parasite
-abundances and loads. These are variables for which we do not have field
-estimates. Thus, we computed a measure of relative cleaner abundance for
-each site relative to absolute abundances and multiplied by a scaling
-constant that changes the range of the variable. We fit the value of the
-scaling constant as part of the statistical inference. As for the client
-abundances for residents and visitors we computed a relative measure
-with respect to the total client abundance, and weighted that by the
-rescaled relative absence of cleaners (1-relative cleaner abundance).
-Thus, all three measures of relative abundance sum up to 1, and can be
-used in the model as a proxy for the probability of having different
-options in the cleaning station.
+clients to be cleaned). The more cleaners there are, the more options
+are available for the clients and hence the higher is the probability of
+empty cleaning stations. How often the cleaning stations are empty in
+turns influences the value that cleaners can expect from their choices.
+The emptier the cleaning stations are, the more benefitial it is to
+prioritize a client that will not stay. However, the frequency of empty
+spots in the cleaning station does not only depend on cleaner relative
+abundances but also on the frequency with which clients visit the
+station. The client visitation frequency will likely vary among
+different clients depending on parasite abundances and loads. These are
+variables for which we do not have field estimates. Thus, we computed a
+measure of relative cleaner abundance for each site relative to absolute
+abundances and multiplied by a scaling constant that changes the range
+of the variable. We fit the value of the scaling constant as part of the
+statistical inference. As for the client abundances for residents and
+visitors we computed a relative measure with respect to the total client
+abundance, and weighted that by the rescaled relative absence of
+cleaners (1-relative cleaner abundance). Thus, all three measures of
+relative abundance sum up to 1, and can be used in the model as a proxy
+for the probability of having different options in the cleaning station.
 
@@
 On the left (panels a, c and c) posterior distributions for a model with   all three parameters; while on the right (b and d) a model without negative reward.
-Panel f shows a measure of fit for both models, namely the distribution of pseudo-$R^2$ obtained from sampling the posterior distribution of parameter values.",out.width='85%'}
+Panel f shows a measure of fit for both models, namely the distribution of pseudo-$R^2$ obtained from sampling the posterior distribution of parameter values.",out.width='95%'}
 knitr::include_graphics(here("post_both_gamma.png"))
@@
 of the parameter values of the full model, which includes both future
-and negative reward. Posterior distributions show the probability density (y axis)
-of the estimated parameter values (x axis). Notably, the bulk of the posterior distribution for
-the parameter tuning negative reward ($\eta$ panel e) is around zero.
-Thus, the shape of the posterior and its credible intervals suggest the absence of negative reward to be well supported by the data.
-As for $\gamma$, the confidence intervals also includes zero, but the mode of the posterior is around 0.5.
-Thus, we run the analysis setting $\eta$ to zero.
-Figure
-\ref{fig:post} panels b and d, show the posterior distributions of the
-two parameters estimated. Not surprisingly, under a model without
-negative reward, the distribution of $\gamma$ shifts to higher values. Under the new model, the estimates of $\gamma$ do not include zero, the distribution is centered around $0.5$.
-To asses the fit of the two models we calculated the  $pseudo-R^2$ [@mcfadden_Conditional_1974], which is a standard measure of fit for logistic regression. 
-In that context, $pseudo-R^2$ uses the log-likelihood of the data given the model, relative to the log-likelihood of the data given a model without covariates, as a measure of fit. 
-Given that our model is not a logistic regression, we measure the $pseudo-R^2$ 
-in relation to the log-likelihood of a model with parameters $\gamma$ and $\eta$ set to zero. This in practice amounts to a model that has neutral preference between 
-the two options. In panel f of figure \ref{fig:post} we show the distribution of $pseudo-R^2$ calculated using samples from the posterior distributions shown before, as well as from a 
-model with negative reward but without future reward. 
-Note, $pseudo-R^2$ can have negative values, that is when the log-likelihood of 
-the model is lower than that of a model that triggers neutral preferences.
-Even though, the peak of the three $pseudo-R^2$ distributions is not very different, 
-the model with only future reward produces a distribution of $pseudo-R^2$ where more values 
+and negative reward. Posterior distributions show the probability
+density (y axis) of the estimated parameter values (x axis). Notably,
+the bulk of the posterior distribution for the parameter tuning negative
+reward ($\eta$ panel e) is around zero. Thus, the shape of the posterior
+and its credible intervals suggest the absence of negative reward to be
+well supported by the data. As for $\gamma$, the confidence intervals
+also includes zero, but the mode of the posterior is around 0.5. Thus,
+we run the analysis setting $\eta$ to zero. Figure \ref{fig:post} panels
+b and d, show the posterior distributions of the two parameters
+estimated. Not surprisingly, under a model without negative reward, the
+distribution of $\gamma$ shifts to higher values. Under the new model,
+the estimates of $\gamma$ do not include zero, the distribution is
+centered around $0.5$. To asses the fit of the two models we calculated
+the $pseudo-R^2$ [@mcfadden_Conditional_1974], which is a standard
+measure of fit for logistic regression. In that context, $pseudo-R^2$
+uses the log-likelihood of the data given the model, relative to the
+log-likelihood of the data given a model without covariates, as a
+measure of fit. Given that our model is not a logistic regression, we
+measure the $pseudo-R^2$ in relation to the log-likelihood of a model
+with parameters $\gamma$ and $\eta$ set to zero. This in practice
+amounts to a model that has neutral preference between the two options.
+In panel f of figure \ref{fig:post} we show the distribution of
+$pseudo-R^2$ calculated using samples from the posterior distributions
+shown before, as well as from a model with negative reward but without
+future reward. Note, $pseudo-R^2$ can have negative values, that is when
+the log-likelihood of the model is lower than that of a model that
+triggers neutral preferences. Even though, the peak of the three
+$pseudo-R^2$ distributions is not very different, the model with only
+future reward produces a distribution of $pseudo-R^2$ where more values
 are positive (to the right of black line in Fig. \ref{fig:post} f). This
-shows that, accounting for variation in the parameter estimates, the model with only future
-reward gives a better fit to the data, despite having one parameter less.
+shows that, accounting for variation in the parameter estimates, the
+model with only future reward gives a better fit to the data, despite
+having one parameter less.
 
@@
 
-Figure \ref{fig:pred} shows predictions from the two models. 
-On the left, the colour coded contour shows the
-prediction of the model for combinations of cleaner abundance (x axis)
-and visitor leaving probability (y axis). Predictions were generated by
-running the model using the mode of the posterior distributions as the
-parameter values, and varying systematically cleaner abundance and
-visitor leaving probability. Colour of the points on top of the contour
-show the average frequency with which cleaner fish chose the visitor
-plate in their last 20 experimental trials. Each point corresponds to a
-location with a combination of cleaner abundance and visitor leaving
-probability. Panels on the right show how close the observed values for 
-the probability of choosing the visitor (y
-axis) are from the predictions of the models (x axis). The line
-represents the perfect fit between observed and predicted values.
-Predicted values were taken using the mode just as in the left panels.
-On the lower right corner of the panel we show the $pseudo-R^2$ for predictions
-using the mode of the posterior distributions. 
-Despite having one more parameter varying, the full model
+Figure \ref{fig:pred} shows predictions from the two models. On the
+left, the colour coded contour shows the prediction of the model for
+combinations of cleaner abundance (x axis) and visitor leaving
+probability (y axis). Predictions were generated by running the model
+using the mode of the posterior distributions as the parameter values,
+and varying systematically cleaner abundance and visitor leaving
+probability. Colour of the points on top of the contour show the average
+frequency with which cleaner fish chose the visitor plate in their last
+20 experimental trials. Each point corresponds to a location with a
+combination of cleaner abundance and visitor leaving probability. Panels
+on the right show how close the observed values for the probability of
+choosing the visitor (y axis) are from the predictions of the models (x
+axis). The line represents the perfect fit between observed and
+predicted values. Predicted values were taken using the mode just as in
+the left panels. On the lower right corner of the panel we show the
+$pseudo-R^2$ for predictions using the mode of the posterior
+distributions. Despite having one more parameter varying, the full model
 gives a lower fit to the data than the model without negative reward.
-The difference in $pseudo-R^2$ for the point estimates is small, but using the whole posterior distribution
-shows the improvement in fit achieved by eliminating negative reward 
-from the model (Fig. \ref{fig:post}f). In the supplementary
-material, we also show that a model with only negative reward, excluding
-future rewards ($\gamma=0$, Fig. \ref{fig:nogamma}), provides a lower
-fit than either of the ones presented in the main text.
+The difference in $pseudo-R^2$ for the point estimates is small, but
+using the whole posterior distribution shows the improvement in fit
+achieved by eliminating negative reward from the model (Fig.
+\ref{fig:post}f). In the supplementary material, we also show that a
+model with only negative reward, excluding future rewards ($\gamma=0$,
+Fig. \ref{fig:nogamma}), provides a lower fit than either of the ones
+presented in the main text.
 
@@
 
-We have shown that using future rewards in the learning process is the cognitive 
-strategy that best explains the intra-specific variation in cleaner performance 
-in the market task. We compared future reward with the use of visitors behaviour as 
-a primary source of negative reward, and a combination of both. All three models 
-capture well the positive relation between visitor leaving behaviour and cleaner
-performance in the market task [@triki_Biological_2019]. For both mechanisms it is 
-indeed necessary to have fleeing behaviour on the side of the visitors for cleaners
-to develop a preference for them. The key difference between future and negative 
-reward emerges in the predicted relation between cleaner relative abundance and 
-cleaner performance. When cleaner relative abundance is low a model with only 
-future reward predicts low performance in the market task regardless of the visitor
-leaving probability. In contrast, a model with negative reward predicts the highest 
-performance with low relative cleaner abundance, particularly together with high 
-probability of visitor leaving (Fig. \ref{fig:pred}). Low cleaner abundances 
-mean the market has an excess of demand for cleaning services. In the model,
-this translates to a cleaning station that is frequently full. Thus, when visitors 
-leave, it's likely that the cleaner will have access to another client, and there will not be much difference in future reward between choosing a visitor and a resident. Thus,
-estimation of future reward will not develop a preference for the visitor in this 
-conditions, because it matches the state of the market.
-On the other hand, the effect of negative reward is unaffected by this market situation.
-In a full station, cleaners will get more often the heterotypic (resident-visitor) option and will develop a preference for the visitor faster.
-Note that when cleaner abundance is high situations with a resident and a visitor will be so rare that neither mechanism will be very efficient in generating a preference for visitors.
-When such choice arises it is still be best to choose the visitor; however, the learning machinery will not be able to develop this preference efficiently.
-Overall, estimation of future reward is then the cognitive mechanism that allows cleaners to adaptively respond conditions to the ecological conditions of the biological market.
+We have shown that using future rewards in the learning process is the
+cognitive strategy that best explains the intra-specific variation in
+cleaner performance in the market task. We compared future reward with
+the use of visitors behaviour as a primary source of negative reward,
+and a combination of both. All three models capture well the positive
+relation between visitor leaving behaviour and cleaner performance in
+the market task [@triki_Biological_2019]. For both mechanisms it is
+indeed necessary to have fleeing behaviour on the side of the visitors
+for cleaners to develop a preference for them. The key difference
+between future and negative reward emerges in the predicted relation
+between cleaner relative abundance and cleaner performance. When cleaner
+relative abundance is low a model with only future reward predicts low
+performance in the market task regardless of the visitor leaving
+probability. In contrast, a model with negative reward predicts the
+highest performance with low relative cleaner abundance, particularly
+together with high probability of visitor leaving (Fig. \ref{fig:pred}).
+Low cleaner abundances mean the market has an excess of demand for
+cleaning services. In the model, this translates to a cleaning station
+that is frequently full. Thus, when visitors leave, it's likely that the
+cleaner will have access to another client, and there will not be much
+difference in future reward between choosing a visitor and a resident.
+Thus, estimation of future reward will not develop a preference for the
+visitor in this conditions, because it matches the state of the market.
+On the other hand, the effect of negative reward is unaffected by this
+market situation. In a full station, cleaners will get more often the
+heterotypic (resident-visitor) option and will develop a preference for
+the visitor faster. Note that when cleaner abundance is high situations
+with a resident and a visitor will be so rare that neither mechanism
+will be very efficient in generating a preference for visitors. When
+such choice arises it is still be best to choose the visitor; however,
+the learning machinery will not be able to develop this preference
+efficiently. Overall, estimation of future reward is then the cognitive
+mechanism that allows cleaners to adaptively respond conditions to the
+ecological conditions of the biological market.
 
-The cognitive mechanism that our analysis highlights does not require a direct flow of information between the cleaner and its social environment.
-The adaptive responses to local conditions in cleaners is not only evidenced by their performance in  experiments, but also by changes in their brain anatomy.
-Indeed, individual cleaners that performed in the laboratory test according to what is best at the location in which they were caught had larger forebrains harbouring more cells than individuals whose performance mismatched locally optimal decision-rules [@triki_Brain_2020].
-This suggests that despite the overall behavioural adjustment, evidenced here and elsewhere
-[@triki_Biological_2019; @triki_Decrease_2018], there are some cleaners better than others at adjusting their behaviour.
-@triki_Brain_2020 refer to these as socially competent cleaners.
-Social competence refers to the ability to optimise social behaviour depending on the available social information [@taborsky_Social_2012a;@bshary_Cooperation_2015].
-However, we have shown here that cleaners are able to adjust their preference 
-according to the social environment without using any direct information about the social
-environment.
-Instead cleaners seem to adjust their preference using information from the visitation patterns of their clients.
-These patters are of course determined by the feedback with the cleaner's social environment (relative abundance).
-Our models, however, do not include the possibility of cleaners responding to their relative abundance directly.
+The cognitive mechanism that our analysis highlights does not require a
+direct flow of information between the cleaner and its social
+environment. The adaptive responses to local conditions in cleaners is
+not only evidenced by their performance in experiments, but also by
+changes in their brain anatomy. Indeed, individual cleaners that
+performed in the laboratory test according to what is best at the
+location in which they were caught had larger forebrains harbouring more
+cells than individuals whose performance mismatched locally optimal
+decision-rules [@triki_Brain_2020]. This suggests that despite the
+overall behavioural adjustment, evidenced here and elsewhere
+[@triki_Biological_2019; @triki_Decrease_2018], there are some
+cleaners better than others at adjusting their behaviour.
+@triki_Brain_2020 refer to these as socially competent cleaners. Social
+competence refers to the ability to optimise social behaviour depending
+on the available social information [@taborsky_Social_2012a;
+@bshary_Cooperation_2015]. However, we have shown here that cleaners
+are able to adjust their preference according to the social environment
+without using any direct information about the social environment.
+Instead cleaners seem to adjust their preference using information from
+the visitation patterns of their clients. These patters are of course
+determined by the feedback with the cleaner's social environment
+(relative abundance). Our models, however, do not include the
+possibility of cleaners responding to their relative abundance directly.
 The mechanism we present is a general way to learn from the environment
-in both asocial and social situations. It would be interesting to contrast this learning mechanism with one tailored to a social situation, and see whether such mechanism 
-fits the data better. More generally, an important challenge ahead is to understand 
-which cognitive mechanisms evolved exclusively for social life.
+in both asocial and social situations. It would be interesting to
+contrast this learning mechanism with one tailored to a social
+situation, and see whether such mechanism fits the data better. More
+generally, an important challenge ahead is to understand which cognitive
+mechanisms evolved exclusively for social life.
 
-One potential difference between cleaners that adjusted adaptively, like the ones
-reported in @triki_Brain_2020, is configurational learning or chunking. The original analysis of the learning model showed that agents needed two mechanism to solve the 
-market task; to account for the future, and to assign value to the combination of clients
-present in the station [@quinones_Reinforcement_2019]. The latter is referred to as configurational learning or chunking [@sutherland_Configural_1989;@miller_Magical_1956]. 
-Theoretical models suggest that the lack of configurational learning is related to a low 
-frequency with which stimuli are encountered together [@kolodny_Evolution_2014;@prat_Role_2021].
-Thus, cleaner relative abundance could be related to this mechanisms as well.
-Given that the lack of either of these two mechanisms yields
-neutral preferences [@quinones_Reinforcement_2019], we can't, based on the current analysis, distinguish between the two possibilities.
-One potential way to distinguish between the two would be to use an experimental set up designed to test configuration learning on cleaners
-with high and low performance in the market task [@woodbury_Learning_1943;@sutherland_Configural_1989].
+One potential difference between cleaners that adjusted adaptively, like
+the ones reported in @triki_Brain_2020, is configurational learning or
+chunking. The original analysis of the learning model showed that agents
+needed two mechanism to solve the market task; to account for the
+future, and to assign value to the combination of clients present in the
+station [@quinones_Reinforcement_2019]. The latter is referred to as
+configurational learning or chunking [@sutherland_Configural_1989;
+@miller_Magical_1956]. Theoretical models suggest that the lack of
+configurational learning is related to a low frequency with which
+stimuli are encountered together [@kolodny_Evolution_2014;
+@prat_Role_2021]. Thus, cleaner relative abundance could be related to
+this mechanisms as well. Given that the lack of either of these two
+mechanisms yields neutral preferences [@quinones_Reinforcement_2019],
+we can't, based on the current analysis, distinguish between the two
+possibilities. One potential way to distinguish between the two would be
+to use an experimental set up designed to test configuration learning on
+cleaners with high and low performance in the market task
+[@woodbury_Learning_1943; @sutherland_Configural_1989].
 
-Combining mechanistic modelling with empirical data provides a powerful tool to identify the cognitive processes underlying learning in animals.
-The combination overcomes the limitations of the general philosophy in animal cognition to apply the logic of Lloyd Morgen’s canon (Occam’s razor).
-Excluding basic reinforcement learning explanations (operant and/or classical conditioning) as a potential explanation in experiments often warrants animals solving the task on the first possible occasion.
-For example, any theory of mind task needs to be solved in the first trial in order to exclude fast conditioning [@heyes_Theory_1998].
-Similarly, subjects need to solve a social learning task on the first trial to accept imitation as a mechanism over stimulus/local enhancement.
-Such strict conditions are virtually never met.
-For example, potato washing by Japanese macaques, an iconic example of social learning, took several years to spread within the group [@kawamura_Process_1959], meaning that any learner had been repeatedly exposed to demonstrations before acquisition.
-Importantly, @galef_Question_1992 refuted imitation as mechanism not simply because of the repeated exposure but because a (rather qualitative) analysis of the spread of potato washing across individuals did not follow the prediction based on imitation learning (see also @hirata_SweetPotato_2001).
-In our case, the number of trials it took cleaners to learn the solution to the market task would never allow to exclude an important role of negative reinforcement based on the data alone.
-However, fitting model predictions to our data set revealed that a more complex mechanism (estimation of future reward) fits the data better.
+Combining mechanistic modelling with empirical data provides a powerful
+tool to identify the cognitive processes underlying learning in animals.
+The combination overcomes the limitations of the general philosophy in
+animal cognition to apply the logic of Lloyd Morgen's canon (Occam's
+razor). Excluding basic reinforcement learning explanations (operant
+and/or classical conditioning) as a potential explanation in experiments
+often warrants animals solving the task on the first possible occasion.
+For example, any theory of mind task needs to be solved in the first
+trial in order to exclude fast conditioning [@heyes_Theory_1998].
+Similarly, subjects need to solve a social learning task on the first
+trial to accept imitation as a mechanism over stimulus/local
+enhancement. Such strict conditions are virtually never met. For
+example, potato washing by Japanese macaques, an iconic example of
+social learning, took several years to spread within the group
+[@kawamura_Process_1959], meaning that any learner had been repeatedly
+exposed to demonstrations before acquisition. Importantly,
+@galef_Question_1992 refuted imitation as mechanism not simply because
+of the repeated exposure but because a (rather qualitative) analysis of
+the spread of potato washing across individuals did not follow the
+prediction based on imitation learning (see also
+@hirata_SweetPotato_2001). In our case, the number of trials it took
+cleaners to learn the solution to the market task would never allow to
+exclude an important role of negative reinforcement based on the data
+alone. However, fitting model predictions to our data set revealed that
+a more complex mechanism (estimation of future reward) fits the data
+better.
 
@@
 ```
+
+```{r param , echo=FALSE,results='asis'}
+  cat(' Table: \\label{tab:param} Parameter values with which the model was run 
+  in the MCMC.
+$\\sigma$ refers to the amplitud of the perturbation kernel with the subscript indicating the associated parameter.
+New values were taken from a uniform distribution.
+$\\alpha$ refers to the learning rate.
+
+  | Parameter | Value |
+|----------:|:-----:|
+| Learning rounds | 10000 |
+| Reward value | 1 |
+| $\\alpha$ | 0.05 |
+| $\\sigma_{\\gamma}$ | 0.3 |
+| $\\sigma_{\\eta}$ | 4 |
+| $\\sigma_{Sca.Const.}$ | 300 |
+| Number of chains | 5 |
+| Chain lenght | $1^5$ |
+')
+
+```

### 2021-11-26 analysisABC.R

@@ -14,6 +14,7 @@ library(ggdist)
 library(bayesplot) 
 library(lattice)
 source(here("../R_files/posPlots.R"))
+source(here("loadData.R"))
 library("jsonlite")
 
 # simdir<-"D:/Neuch_simulations/ABCfit/Simulations/"
@@ -211,8 +212,10 @@ nullLikehood<-sum(dbinom(x=fieldData[,score_visitor],
 
 loglikehoods.both[,Rsqrd:=1-lglikelihood/nullLikehood]
 
+loglikehoods.both[,model:=factor(model,levels=c("gam","Nrew","both"))]
+
 ggplot(data=loglikehoods.both[is.finite(lglikelihood)],
-       aes(fill=model,x=lglikelihood))+
+       aes(fill=model,x=lglikelihood,y=model))+
   stat_halfeye(alpha=0.4)+theme_classic()+xlim(-750,-400) +
   # scale_fill_manual(values=c("#d73027", "#4575b4"))+
   geom_vline(xintercept = nullLikehood,color="red")
@@ -221,16 +224,16 @@ ggplot(data=loglikehoods.both[is.finite(lglikelihood)],
 df.Rsqrd.mode<-data.table(model=c("both","gam"),
                           Rsqr=c(rsqr.both.McFadden,rsqr.gam.McFadden))
 
-rsqr.both<-ggplot(data=loglikehoods.both[is.finite(lglikelihood)],aes(fill=model,x=Rsqrd))+
-  stat_halfeye(alpha=0.5,point_size=3)+theme_classic()+xlim(-0.5,0.3)+
+rsqr.both<-
+  ggplot(data=loglikehoods.both[is.finite(lglikelihood)],
+                  aes(y=model,x=Rsqrd,fill=model))+
+  stat_halfeye(alpha=0.5,point_size=3)+
+  theme_classic()+xlim(-0.5,0.3)+
   scale_fill_manual(values=c("#1b9e77","#d95f02","#7570b3"),
-                    name = "Model", labels = c("Full", "Future reward","Negative reward"))+
-  
-  # geom_vline(data=df.Rsqrd.mode,aes(xintercept = Rsqr,color=model),size=1)+
-  scale_color_manual(values=c("#d73027", "#4575b4"),
-                     name = "Model", labels = c("Full", "Only future reward"))+
+  name = "Model", labels = c("Future reward","Negative reward","Full"))+
   geom_vline(xintercept = 0,color="black",size=1)+
[…]

### 2021-11-29 ActCritABC.cpp
The competence/incompetence seems to be working well. Will now merge to master

@@ -1039,44 +1039,45 @@ void do_simulation(//del focal_model,
 	abs2rel_abund(emp_data, focal_comb);
 	// Loop through the data points
 	for (int id_data_point = 0; id_data_point < emp_data.size(); ++id_data_point) {
-		/*if (id_data_point > 0 &&
-			emp_data[id_data_point].site_year == emp_data[id_data_point - 1].site_year) {
+		if (id_data_point > 0 &&
+			(emp_data[id_data_point].site_year == emp_data[id_data_point - 1].site_year &&
+			emp_data[id_data_point].group == emp_data[id_data_point - 1].group)) {
 			emp_data[id_data_point].marketPred = emp_data[id_data_point - 1].marketPred;
 		}
 		else
-		{*/
-		init = focal_comb.gamma[emp_data[id_data_point].group]*
-			(1 - pow(1 -
-				emp_data[id_data_point].rel_abund_resid -
-				emp_data[id_data_point].rel_abund_visitors, 2)) / 
-				(1 - focal_comb.gamma[emp_data[id_data_point].group]);
-		cleaners[emp_data[id_data_point].group]->rebirth(init);
-		draw(clientSet, sim_param["totRounds"],
-			emp_data[id_data_point].rel_abund_resid,
-			emp_data[id_data_point].rel_abund_visitors);
-		idClientSet = 0;
-		VisPref = 0, countRVopt = 0;
-		// Loop through the learning rounds
-		for (int trial = 0; trial < sim_param["totRounds"]; ++trial) {
-			cleaners[emp_data[id_data_point].group]->
-				act(clientSet, idClientSet, emp_data[id_data_point].prob_Vis_Leav,
-				sim_param["ResProbLeav"], sim_param["VisReward"],
-				sim_param["ResReward"], sim_param["inbr"], sim_param["outbr"],
-				learnScenario(sim_param["scenario"]));
-		cleaners[emp_data[id_data_point].group]->update();
-			if (trial > int(sim_param["totRounds"]) * float(sim_param["propfullPrint"])) {
-				if (cleaners[emp_data[id_data_point].group]->getstate(0) == 0) {
-					++countRVopt;
-					if (cleaners[emp_data[id_data_point].group]->
-						cleanOptionsT[cleaners[emp_data[id_data_point].group]->getChoice(0)] == visitor) 
-						++VisPref;
+		{
[…]

### 2021-11-29 analysisABC.R

@@ -18,7 +18,7 @@ library("jsonlite")
 
 # simdir<-"D:/Neuch_simulations/ABCfit/Simulations/"
 
-scen1<-"MCMCclean_gam_Nrew_sca_"
+scen1<-"MCMCclean_gam_nRew_sca_group_"
 scen2<-"MCMCclean_gam_sca_"
 scen3<-"MCMCclean_nRew_sca_"
 
@@ -31,7 +31,7 @@ MCMCdata2<-loadMCMCrun(scen2)
 MCMCdata3<-loadMCMCrun(scen3)
 
 # One per onject
-ABCraw<-fread(here("Simulations",scen,ABCruns[3]))
+ABCraw<-fread(here("Simulations",scen1,"MCMCchain_CL100000_seed1.txt"))
 
 # Defaults
 burn.in<-1000
@@ -446,7 +446,7 @@ lines(y=predict.lm(mod1.2,data.frame(negReward=seq(0,5,length=1000))),
 # ABCraw<-fread(ABCruns[3])
 ABCraw<-fread(here("Simulations",scen,ABCruns[2]))
 
-ABC.mcmc<-mcmc(ABCraw[,.(gamma,negReward,scaleConst)])
+ABC.mcmc<-mcmc(ABCraw[,.(gamma,gamma.1,negReward,negReward.1,scaleConst)])
 effectiveSize(ABC.mcmc)
 
 (sumMCMC<-summary(ABC.mcmc))
@@ -460,6 +460,7 @@ HPDinterval(ABC.mcmc)
 raftery.diag(ABC.mcmc)
 autocorr(ABC.mcmc)
 rejectionRate(ABC.mcmc)
+gelman.plot(ABC.mcmc)
 
 hist(ABC.mcmc[,1],breaks = 100)
 hist(ABC.mcmc[,2],breaks = 100)

### 2021-11-29 parTOjson.R

@@ -194,7 +194,7 @@ param_mcmc<-list(totRounds=10000,ResReward=1,VisReward=1,
                 ResProbLeav=0,scenario=0, 
                 #nature, experiment, marketExperiment, ExtendedMarket
                 inbr=0,outbr=0,forRat=0.0,
-                seed=1, propfullPrint = 0.7,sdPert=c(0.05,0.05,0.3,4,300),
+                seed=1, propfullPrint = 0.7,sdPert=c(0.05,0.05,0.25,2,200),
                 chain_length=100000,
                 init=c(0.05,0.05,0,0,30),# alphaA,AlphaC, Gamma, NegRew
                 pertScen = c(FALSE,FALSE,TRUE,TRUE,TRUE), 

### 2021-12-07 manuscript_1.0.Rmd
New file to make figures for grouped models. Updated source code to allow for predictions with grouped data.

@@
 [@bshary_Cleaner_2001a]. More generally, cleaners apparently prefer to
-interact with visitors [@triki_Zeni_inpress]. Such a preference may be
+interact with visitors [@triki_Marine_2021]. Such a preference may be
 linked to a client's value as a food patch; visitors are on average
 larger than residents [ @bshary_Cleaner_2001a], and size is positively
-correlated with parasite load(Grutter xxxx). However, early laboratory
+correlated with parasite load [@grutter_Relationship_1995a]. However, 
+early laboratory
 experiments showed that leaving causes service priority independent of
@@
 negative reinforcer is part of the ubiquitous learning mechanism termed
-operant conditioning (Thorndike 1917), this explanation can be
+operant conditioning [@thorndike_Animal_1898; @skinner_Behavior_1938], 
+this explanation can be
 considered the simpler one. Furthermore, accounting for the future
@@
 performance we need to translate the absolute abundances of the data to
-relative abundances. This is because in the model, relative abundances
+to  a measure of relative abundances that captures client visitation patterns.
+This is because in the model, relative abundances
 of clients define not only the probability of residents and visitors,
@@
 turns influences the value that cleaners can expect from their choices.
-The emptier the cleaning stations are, the more benefitial it is to
+The emptier the cleaning stations are, the more beneficial it is to
 prioritize a client that will not stay. However, the frequency of empty
-spots in the cleaning station does not only depend on cleaner relative
-abundances but also on the frequency with which clients visit the
+spots in the cleaning station does not only depend on cleaner abundances 
+but also on the frequency with which clients visit the
 station. The client visitation frequency will likely vary among
@@
 abundances and multiplied by a scaling constant that changes the range
-of the variable. We fit the value of the scaling constant as part of the
+of the variable. This scaling constant is meant to capture variation in the market 
+conditions driven partly by cleaner abundance. We fit the value of the 
+scaling constant as part of the
 statistical inference. As for the client abundances for residents and

### 2021-12-07 ActCritABC.cpp

@@ -777,7 +777,7 @@ void initializeRoundFile(ofstream& RoundOutput, nlohmann::json param,
 			<< "prob.Vis.Leav	"
 			<< "market_binomial_data	"
 			<< "market_binomial_pred	"
-			<< "competence	" << endl;
+			<< "competence" << endl;
 	}
 	else{
 		RoundOutput << "site_year	"
@@ -787,7 +787,8 @@ void initializeRoundFile(ofstream& RoundOutput, nlohmann::json param,
 			<< "rel.abund.residents	"
 			<< "prob.Vis.Leav	"
 			<< "visitorChoices	"
-			<< "visitorChoices_pred" << endl;
+			<< "visitorChoices_pred	"
+			<< "competence" << endl;
 	}
 	
 }
@@ -814,7 +815,7 @@ void printRoundFile(ofstream& roundOut, const std::vector<cleaner_point>& emp_da
 			emp_data[id_data_point].prob_Vis_Leav << '\t' <<
 			emp_data[id_data_point].countVisitor << '\t' <<
 			emp_data[id_data_point].marketPred << '\t' << 
-			emp_data[id_data_point].group << '\t' << endl;
+			emp_data[id_data_point].group << endl;
 	}
 }
 
@@ -1115,15 +1116,16 @@ int main(int argc, char* argv[]){
 	//// alphaA, alphaC, Gamma, NegRew,scaleConst
 	//sim_param["chain_length"] = 100;
 	//sim_param["init"]       = {0.05, 0.05 , 0.93,0.02, 58};
+	//sim_param["init2"] =	{ 0.05, 0.05 , 0.93,0.02, 58 };
 	// //alphaA, alphaC, gamma, NegRew, scaleConst
 	//sim_param["pertScen"] = {false,false,true,true,true};
 	////enum perturnScen {all,  bothFut, justGam, justNegRew};
-	//sim_param["MCMC"] = 1;
+	//sim_param["MCMC"] = 0;
 	//sim_param["data"] = "clean"; // "loc", "clean"
 	//sim_param["nRep"] = 30 ;
 	//sim_param["folder"]       = "M:/Projects/Clean.ActCrit/Simulations/ABCtest_/";
 	//sim_param["dataFile"] = "M:/Projects/Clean.ActCrit/Data/data_ABC_cleaner_abs_threa1.5.txt";
-	//sim_param["Group"] = false;
+	//sim_param["Group"] = true;
 
[…]

### 2021-12-07 analysisABC.R

@@ -4,7 +4,6 @@
 ## Libraries
 
 require(here)
-library(plot3D) 
 require(data.table)
 library(coda)
 library(mcmcplots)
@@ -20,36 +19,47 @@ library("jsonlite")
 # simdir<-"D:/Neuch_simulations/ABCfit/Simulations/"
 
 scen1<-"MCMCclean_gam_nRew_sca_group_"
-scen2<-"MCMCclean_gam_sca_"
-scen3<-"MCMCclean_nRew_sca_"
+scen2<-"MCMCclean_gam_sca_group_"
+scen3<-"MCMCclean_gam_sca_"
+
+labels.Scen<-c("both.group","gam.group","gam")
+labelsPlot.Scen<-c("Full grouped","Future reward grouped","Future reward")
 
 ## Load files --------------------------------------------------------------
-
-# all chains in a single object
-
-MCMCdata1<-loadMCMCrun(scen1)
-MCMCdata2<-loadMCMCrun(scen2)
-MCMCdata3<-loadMCMCrun(scen3)
-
-# One per onject
-ABCraw<-fread(here("Simulations",scen1,"MCMCchain_CL100000_seed1.txt"))
 
 # Defaults
 burn.in<-1000
 thinning<-100
 
 
+# all chains in a single object
+
+# MCMCdataAll<-lapply(listScen,loadMCMCrun,thinning = thinning,burn.in = burn.in)
+MCMCdata1<-loadMCMCrun(scen1,thinning = thinning,burn.in = burn.in)
+MCMCdata2<-loadMCMCrun(scen2,thinning = thinning,burn.in = burn.in)
+MCMCdata3<-loadMCMCrun(scen3,thinning = thinning,burn.in = burn.in)
+
+# One per onject
+ABCraw<-fread(here("Simulations",scen1,"MCMCchain_CL100000_seed1.txt"))
+
+
 # All in a list
 mcmcList.1<-mcmc.list(do.call(list,lapply(unique(MCMCdata1$seed), function(repli){
   rundata<-MCMCdata1[seed==repli]
-  mcmcRun<-mcmc(rundata[,.(gamma,negReward,scaleConst)],thin = 100)#
+  mcmcRun<-mcmc(rundata[,.(gamma,gamma.1,negReward,negReward.1,scaleConst)],thin = thinning)#
+  return(mcmcRun)
+})))
+
[…]

### 2021-12-07 loadFieldData.R

@@ -80,13 +80,27 @@ fieldData.sum[,competence:=highDens+visPref!=1]
 
 
 
-fieldData.sum %>%
-  ggplot(aes(x=abund.cleaners,y=score_visitor,color=competence))+
+absplot<-fieldData.sum %>%
+  ggplot(aes(x=abund.cleaners,y=score_visitor,
+             color=as.factor(competence)))+
   geom_point()+geom_vline(xintercept = 1.5)+
   scale_color_manual(values=c("#999999", "#E69F00"))+
+  labs(color="Competence")+geom_jitter()+
+  geom_text(x=1.7,y=2,label="Threshold",inherit.aes = FALSE)+
   theme_classic()
 
-fieldData.sum[,sum(competence)]
+ratioPlot<-fieldData.sum %>%
+  ggplot(aes(x=abund.cleaners/(abund.visitors+abund.visitors),y=score_visitor,
+             color=as.factor(competence)))+
+  geom_point()+#geom_vline(xintercept = 1.5)+
+  scale_color_manual(values=c("#999999", "#E69F00"))+
+  labs(color="Competence")+geom_jitter(width = 0.0005)+
+  theme_classic()
+
+png(here("threshold.png"))
+ggarrange(absplot,ratioPlot,labels = c("a","b"),common.legend = TRUE,ncol = 1)
+dev.off()
+
 
 fieldData.site %>%
   ggplot(aes(x=abundance_cleaners_100m2,fill=score_visitor,y=prob.Vis.Leav))+

### 2021-12-07 manFiguresgroup.R

@@ -1,0 +1,346 @@
+#### Figures showing the fit between model and data ############################
+
+library(data.table)
+library(ggplot2)
+library("RColorBrewer")
+library("patchwork") 
+library(ggpubr)
+library(here)
+library(cowplot)
+source(here("loadData.R"))
+source(here("data2interp.R"))
+require('akima')
+source(here("aesth_par.R"))
+source(here("..","R_files",'ternaryAEQP.R'))
+source(here("../R_files/posPlots.R"))
+
+
+scenario<-"MCMCclean_gam_sca_group"
+# scenario2<-"MCMCclean_gam_sca"
+
+predfileMode.both<-grep(".txt",grep("round",list.files(here("Simulations",paste0(scenario,"_"))),value = T),
+                    value = T)
+
+# predfileMode.gam<-grep(".txt",grep("round",list.files(here("Simulations",paste0(scenario2,"_"))),value = T),
+#                    value = T)
+# 
+# predfileSamples.both<-grep(".txt",grep("round",list.files(here("Simulations",
+#                                                                paste0(scenario,"_"),"samplesPost_"),
+#                                                           full.names = TRUE)
+#                                        ,value = T),
+#                          value = T)
+
+# predfileSamples.gam<-grep(".txt",grep("round",list.files(here("Simulations",
+#                                                                paste0(scenario2,"_"),"samplesPost_"),
+#                                                           full.names = TRUE)
+#                                        ,value = T),
+#                            value = T)
+
+
+predictDataMode.both<-fread(here("Simulations",paste0(scenario,"_"),
+                      predfileMode.both))
+
+
+# predictDataMode.gam<-fread(here("Simulations",paste0(scenario2,"_"),
+#                              predfileMode.gam))
+# 
+# predictDataSamps.both<-do.call(rbind,lapply(predfileSamples.both,fread))
+# predictDataSamps.both[,id_samp:=rep(1:100,each=120)]
+# 
+# predictDataSamps.gam<-do.call(rbind,lapply(predfileSamples.gam,fread))
[…]

### 2021-12-07 parTOjson.R

@@ -13,198 +13,47 @@ fileName<-"parameters.json"
 scenario<-"MCMCclean_gam_nRew_sca_group_"
 
 
-
-# For predictions on the observed values ---------------------------------------
-
-scenario<-"MCMCclean_gam_sca"
-
-MCMCdata<-loadMCMCrun(paste0(scenario,"_"))
-densGamma<-density(MCMCdata$gamma)
-modeGamma<-densGamma$x[densGamma$y==max(densGamma$y)]
-densNR<-density(MCMCdata$negReward)
-modeNegrew<-densNR$x[densNR$y==max(densNR$y)]
-densScal<-density(MCMCdata$scaleConst)
-modescal<-densScal$x[densScal$y==max(densScal$y)]
-
-
-param_pred<-list(totRounds=10000,ResReward=1,VisReward=1,
-            ResProbLeav=0,scenario=0, inbr=0,outbr=0,forRat=0.0,
-            seed=1, propfullPrint = 0.7,sdPert=c(0.05,0.05,0.1,0.05,1),
-            chain_length=0,
-            init=c(0.05,0.05,0,0,35),# alphaA,AlphaC, Gamma, NegRew, scalConst
-            pertScen = c(FALSE,FALSE,FALSE,FALSE,TRUE), 	
-            MCMC =0, data="clean",nRep=30,
-            dataFile = here("Data","data_ABC_cleaner_absolute.txt"),
-            # folderL=paste0(here(simsDir),"/",scenario,"_/","samplesPost_/"))
-            folderL=paste0(here(simsDir),"/",scenario,"_/"))
-
-
-check_create.dir(here(simsDir,paste0(scenario,"_")),param = rep("samplesPost",1),
-                 values = c(""))
-
-param_pred$folder<-param_pred$folderL
-# param_pred$init[c(3,4,5)]<- c(modeGamma,0,modeScal)
-  # c(runif(1,max = 1),runif(1,max = 1),
-  #                            runif(1,max = 500,min = 1))
-  #sumMCMC$statistics[,1]
-nsamples<-100
-
-postSamp<-MCMCdata[round(runif(nsamples,min = 1,max=dim(MCMCdata)[1]),
-                              digits = 0),]
-
-for(i in 1:nsamples){
-  param_pred$init[c(3,4,5)]<- c(postSamp[i,.(gamma,negReward,scaleConst)])
-  param_pred$seed<- i
[…]

### 2021-12-21 manuscript_1.0/manuscript_1.0.Rmd
Added a version of the manuscript in the Royal society format

--- 2021-12-07 manuscript_1.0.Rmd
@@
 ---
-title: "Cognitive and ecological aspects determine cleaner performance"
+title: Combination of field and experimental data together with computational models 
+  reveal cognitive mechanisms behind cleaner fish behaviour
+#Cognitive and ecological aspects determine cleaner performance
 author:
-- Andrés E. Quiñones^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland - andresqp@gmail.com]
-- Zegni Triki^[Department of Zoology, Stockholm University, Stockholm, Sweden]
-- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
-date: "`r format(Sys.time(), '%B %d, %Y')`"
+  - name: Andrés E. Quiñones 
+    affiliation: "1"
+  - name: Zegni Triki
+    affiliation: "2"
+  - name: Redouan Bshary
+    affiliation: "1"
+address:
+  - code: "1"
+    address: Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
+  - code: "2"
+    address: Department of Zoology, Stockholm University, Stockholm, Sweden
+
+corresp_author_name:  "AE. Quiñones"
+corresp_author_email: "andreseqp@gmail.com"
+
+subject:
+  - "Behavioural ecology"
+  - "Cognitive ecology"
+  - "Animal behaviour"
+
+keywords:
+  - learning   
+  - cleaners
+  - behaviour
+  - bayesian statisitics
+  - mutualism
+
+abstract: |
+  The same behaviour can be achieved by different mechanistic underpinnings.
+But the environmental conditions under which the behaviour is triggered might differ between these potential mechanisms.
+Thus, by analyzing the varitation of expression of a certain behavior across environmental gradients, it is possible to infer which mechanism is behind the behaviour of interest.
+To do this it is necessary to have clearcut predictions for the different candidate mechanisms.
+In here, we use experimental data on the behaviour of the cleaner fish *Labroides* *dimidiatus*, as well as enviromental variables from the location of fish, together with a computational model of the learning mechanisms.
+Using bayesian statistics, we fit key parameters of the computational model that determine the specific learning mechanism.
+We find that a model where cleaners estimate the long-term consequences of their actions, and use the estimate in their desicion making, explains better under which conditions the fish develop a preference for an ephemeral food source in the context of a biological market.
+We argue that the combination of computational models with data is powerful tool to infer the mechanistic underpinning behind behaviours of interest.
+
+## Remove this if not required
+ethics: |
+  The Animal Ethics Committee of the Queensland government (DAFF)
+  approved the project (CA 2016/05/970 and CA 2017/05/1063).
+
+data_accessibility: |
+  Please provide details on the data availability.
+
+author_contributions: |
+  AQ and RB designed the study. ZT collected the data. AQ developed the model and 
+  fitted the model parameters to the data. AQ, RB and ZT wrote the manuscript.
+
+## Remove this if not required
+conflict_of_interest: |
+  The authors declate no conflict of interests.
+
+## Remove this if not required
+funding: |
+  This work was supported by the Swiss National Science Foundation (grants
+  number: 31003A_153067/1 and 310030B_173334/1 to R.B.).
+
+## Remove this if not required
+acknowledgements: |
+  ZT and RB kindly thank the staff of Lizard Island Research Station, Dominique
+  Roche, Sharon Wismer, Olivia Rey,Sandra Ann Binning, Elena Levorato and William McNeely for their
+  field support. AQ thanks Florian Hartig for statistical advice.
+
+bibliography: Cleanerlearning.bib
+
+## change to true to add optional line numbering
+lineno: true
+
+header-includes: |
+  ```{=latex}
+  \usepackage{float}
+  \usepackage{booktabs}
+  \newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+  ```
+site: bookdown::bookdown_site
 output: 
-  pdf_document: default
-  word_document: default
-fig_caption: yes
-fontsize: 12pt
-header-includes:
-- \usepackage{float}
-- \usepackage{setspace}\doublespacing
-- \linespread{2}
-- \usepackage{lineno}
-- \linenumbers
-- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
-  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
-keep_tex: yes
-keywords: My keywords
-linespread: 2
-bibliography: Cleaner_learning.bib
-style: M:\Zotero\styles\behavioral-ecology-and-sociobiology.csl
-spacing: double
-abstract: Insert a nice abstract here.
-editor_options:
-  markdown:
-    wrap: 72
+  bookdown::pdf_book:
+    base_format: rticles::rsos_article
 ---
+
+## Introduction
 
@@
 ```
-
-## Introduction
 
@@
 [@hammerstein_Biological_2016; @leimar_Cooperation_2010;
-@noe_Biological_1994a,@noe_Biological_1995a]. Thus, we expect supply
+@noe_Biological_1994a;@noe_Biological_1995a]. Thus, we expect supply
 and demand of the good and services being exchanged to determine the
@@
 improved health [@waldie_LongTerm_2011;
-@ros_Does_2011,@triki_Effects_2016]. As cleaners have about 2000
+@ros_Does_2011;@triki_Effects_2016]. As cleaners have about 2000
 interactions per day [@grutter_Relationship_1995], they regularly face
@@
 the design of the two plates (colour versus size;
-@wismer_Cuebased_2019), or whether the information is in the colour of
+[@wismer_Cuebased_2019]), or whether the information is in the colour of
 the food [@pretot_Comparative_2021; @pretot_Comparing_2016], and
@@
 their predictions in light of the behavioural data (e.g.
-@dubois_Model_2021). By varying the cognitive tool kit, one can assess
+[@dubois_Model_2021]). By varying the cognitive tool kit, one can assess
 which minimal kit is necessary to solve the task at hand. This approach
@@
 
-We used the data from @triki_Biological_2019 to fit the parameters of
-the model analysed in @quinones_Reinforcement_2019. For the bayesian
+We used the data from [@triki_Biological_2019] to fit the parameters of
+the model analysed in [@quinones_Reinforcement_2019]. For the bayesian
 fitting procedure we used Markov Chain Monte Carlo (MCMC) with a
@@
 reward the individual obtains (prediction error)
-[@sutton_Reinforcement_2018; \@rescola_Theory_1972]; and it is carried
+[@sutton_Reinforcement_2018; @rescorla_Theory_1972]; and it is carried
 in the direction that would lead to more reward being obtained, given
@@
 these parameters allow individuals to use in their estimates the future
-effects of their choices. The model analysis by
-@quinones_Reinforcement_2019 shows that without either of them, learning
+effects of their choices. The model analysis by Quiñones *et al*
+[@quinones_Reinforcement_2019] shows that without either of them, learning
 individuals cannot develop a preference for the visitor client. The
@@
 reward as an explanation for cleaner ability to prefer visitors is
-supported by @triki_Biological_2019 data. However, a combination of
+supported by Triki *et al* [@triki_Biological_2019] data. However, a combination of
 future and negative rewards, could still be a better explanation for the
 data. In the current analysis, we will use the data from
-@triki_Biological_2019 to fit some of the model parameters, particularly
+[@triki_Biological_2019] to fit some of the model parameters, particularly
 relevant we will fit $\gamma$ and $\eta$ to find out whether each or a
@@
 [@triki_Biological_2019; @triki_Decrease_2018], there are some
-cleaners better than others at adjusting their behaviour.
-@triki_Brain_2020 refer to these as socially competent cleaners. Social
+cleaners better than others at adjusting their behaviour. Triki *et al.*
+[@triki_Brain_2020] refer to these as socially competent cleaners. Social
 competence refers to the ability to optimise social behaviour depending
@@
 One potential difference between cleaners that adjusted adaptively, like
-the ones reported in @triki_Brain_2020, is configurational learning or
+the ones reported in Triki *et al.* [@triki_Brain_2020], is configurational learning or
 chunking. The original analysis of the learning model showed that agents
@@
 [@kawamura_Process_1959], meaning that any learner had been repeatedly
-exposed to demonstrations before acquisition. Importantly,
-@galef_Question_1992 refuted imitation as mechanism not simply because
+exposed to demonstrations before acquisition. Importantly, Galef
+[@galef_Question_1992] refuted imitation as mechanism not simply because
 of the repeated exposure but because a (rather qualitative) analysis of
@@
 prediction based on imitation learning (see also
-@hirata_SweetPotato_2001). In our case, the number of trials it took
+[@hirata_SweetPotato_2001]). In our case, the number of trials it took
 cleaners to learn the solution to the market task would never allow to
@@
 better.
-
-## Cited literature
-
-::: {#refs}
-:::
 
@@
 Right-hand side panel: scatter plot of the observed and predicted probabilities of choosing a visitor for 12 locations.
-Line correspond to a perfect match between observed and predicted probabilities.",out.width='90%'}
+Line correspond to a perfect match between observed and predicted probabilities.",out.width='90%',fig.pos = 'H', out.extra = ""}
 fileNogamma<-grep("contour_gamma_",list.files(here("Simulations","MCMCclean_nRew_sca_"),

### 2021-12-21 manuscript_1.0/manuscript_1.0.tex
Added a version of the manuscript in the Royal society format

--- 2021-12-07 manuscript_1.0.Rmd
@@
----
-title: "Cognitive and ecological aspects determine cleaner performance"
-author:
-- Andrés E. Quiñones^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland - andresqp@gmail.com]
-- Zegni Triki^[Department of Zoology, Stockholm University, Stockholm, Sweden]
-- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
-date: "`r format(Sys.time(), '%B %d, %Y')`"
-output:
-  pdf_document: default
-  word_document: default
-fig_caption: yes
-fontsize: 12pt
-header-includes:
-- \usepackage{float}
-- \usepackage{setspace}\doublespacing
-- \linespread{2}
-- \usepackage{lineno}
-- \linenumbers
-- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
-  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
-keep_tex: yes
-keywords: My keywords
-linespread: 2
-bibliography: Cleaner_learning.bib
-style: M:\Zotero\styles\behavioral-ecology-and-sociobiology.csl
-spacing: double
-abstract: Insert a nice abstract here.
-editor_options:
-  markdown:
-    wrap: 72
----
+%% Author_tex.tex
+%% V1.0
+%% 2012/13/12
+%% developed by Techset
+%%
+%% This file describes the coding for rsproca.cls
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-library(here)
-```
+\documentclass[]{rsos}%%%%where rsos is the template name
 
-## Introduction
+\usepackage{lineno}
+\linenumbers
+
+\usepackage[T1]{fontenc}
+\usepackage[utf8]{inputenc}
+
+% For Pandoc highlighting needs
+
+% Pandoc citation processing
+
+\usepackage{float}
+\usepackage{booktabs}
+\newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+
+%%%% *** Do not adjust lengths that control margins, column widths, etc. ***
+
+%%%%%%%%%%% Defining Enunciations  %%%%%%%%%%%
+\newtheorem{theorem}{\bf Theorem}[section]
+\newtheorem{condition}{\bf Condition}[section]
+\newtheorem{corollary}{\bf Corollary}[section]
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{document}
+
+%%%% Article title to be placed here
+\title{Combination of field and experimental data together with computational models reveal cognitive mechanisms behind cleaner fish behaviour}
+
+\author{
+Andrés E. Quiñones$^{1}$,
+Zegni Triki$^{2}$,
+Redouan Bshary$^{1}$}
+
+\address{
+  $^{1}$Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland\\
+  $^{2}$Department of Zoology, Stockholm University, Stockholm, Sweden}
+%%%% Subject entries to be placed here %%%%
+\subject{
+Behavioural ecology,
+Cognitive ecology,
+Animal behaviour}
+
+%%%% Keyword entries to be placed here %%%%
+\keywords{
+learning,
+cleaners,
+behaviour,
+bayesian statisitics,
+mutualism}
+
+%%%% Insert corresponding author and its email address}
+\corres{
+  AE. Quiñones\\
+  e-mail: \href{mailto:andreseqp@gmail.com}{\nolinkurl{andreseqp@gmail.com}}
+}
+
+%%%% Abstract text to be placed here %%%%%%%%%%%%
+\begin{abstract}
+The same behaviour can be achieved by different mechanistic underpinnings.
+But the environmental conditions under which the behaviour is triggered might differ between these potential mechanisms.
+Thus, by analyzing the varitation of expression of a certain behavior across environmental gradients, it is possible to infer which mechanism is behind the behaviour of interest.
+To do this it is necessary to have clearcut predictions for the different candidate mechanisms.
+In here, we use experimental data on the behaviour of the cleaner fish \emph{Labroides} \emph{dimidiatus}, as well as enviromental variables from the location of fish, together with a computational model of the learning mechanisms.
+Using bayesian statistics, we fit key parameters of the computational model that determine the specific learning mechanism.
+We find that a model where cleaners estimate the long-term consequences of their actions, and use the estimate in their desicion making, explains better under which conditions the fish develop a preference for an ephemeral food source in the context of a biological market.
+We argue that the combination of computational models with data is powerful tool to infer the mechanistic underpinning behind behaviours of interest.
+\end{abstract}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Some pieces required from the pandoc template
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\providecommand{\EndFirstPage}{%
+}
+
+\maketitle
+
+\hypertarget{introduction}{%
+\subsection{Introduction}\label{introduction}}
 
@@
 interactions such as pollination, nitrogen fixation or parasite removal
-[@begon_Ecology_2006]. When partners in these mutualistic interactions
+\citep{begon_Ecology_2006}. When partners in these mutualistic interactions
 are able to choose their interaction partner, their investment and the
 return they obtain can be understood as a biological market
-[@hammerstein_Biological_2016; @leimar_Cooperation_2010;
-@noe_Biological_1994a,@noe_Biological_1995a]. Thus, we expect supply
+\citep{hammerstein_Biological_2016, leimar_Cooperation_2010, noe_Biological_1994a, noe_Biological_1995a}. Thus, we expect supply
 and demand of the good and services being exchanged to determine the
@@
 One example of strategic adjustment in a biological market is the marine
-cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
-'client' fish. Clients offer themselves as food patches for cleaners to
+cleaning mutualism involving the cleaner fish \emph{Labroides dimidiatus} and
+`client' fish. Clients offer themselves as food patches for cleaners to
 remove ectoparasites, which gives the cleaners calories and the clients
-improved health [@waldie_LongTerm_2011;
-@ros_Does_2011,@triki_Effects_2016]. As cleaners have about 2000
-interactions per day [@grutter_Relationship_1995], they regularly face
+improved health \citep{waldie_LongTerm_2011, ros_Does_2011, triki_Effects_2016}. As cleaners have about 2000
+interactions per day \citep{grutter_Relationship_1995}, they regularly face
 the situation that two clients seek cleaning simultaneously. In such
-conditions, it is important to distinguish between 'resident' and
-'visitor' client species [@bshary_Cleaner_2001a]. The residents have
+conditions, it is important to distinguish between `resident' and
+`visitor' client species \citep{bshary_Cleaner_2001a}. The residents have
 small home ranges and hence access to only the local cleaner, while the
@@
 visitors have been observed to use their partner choice options in that
-way [@bshary_Choosy_2002], which may explain why cleaners give
+way \citep{bshary_Choosy_2002}, which may explain why cleaners give
 visitors service priority in a field study in the Red Sea
-[@bshary_Cleaner_2001a]. More generally, cleaners apparently prefer to
-interact with visitors [@triki_Marine_2021]. Such a preference may be
+\citep{bshary_Cleaner_2001a}. More generally, cleaners apparently prefer to
+interact with visitors \citep{triki_Marine_2021}. Such a preference may be
 linked to a client's value as a food patch; visitors are on average
-larger than residents [ @bshary_Cleaner_2001a], and size is positively
-correlated with parasite load [@grutter_Relationship_1995a]. However, 
+larger than residents \citep{bshary_Cleaner_2001a}, and size is positively
+correlated with parasite load \citep{grutter_Relationship_1995a}. However,
 early laboratory
 experiments showed that leaving causes service priority independent of
-client value [@bshary_Asymmetric_2002]. The experiments consisted of a
+client value \citep{bshary_Asymmetric_2002}. The experiments consisted of a
 simultaneous two-choice task, where cleaners had to decide which of two
@@
 and eat mashed prawn of it. One plate was made to behave like a
-resident, i.e. it would remain until the cleaner had eaten off it. The
-other plate was made to behave like a visitor, i.e. it would be removed
+resident, i.e.~it would remain until the cleaner had eaten off it. The
+other plate was made to behave like a visitor, i.e.~it would be removed
 if not inspected first. Cleaners learned to prefer the visitor plate and
-hence obtained the double amount of food [@bshary_Asymmetric_2002].
+hence obtained the double amount of food \citep{bshary_Asymmetric_2002}.
 Adult cleaner fish outperform various primates as well as rats and
 pigeons in this original version of the biological market task
-[@salwiczek_Adult_2012; @zentall_Early_2017]; only African grey
-parrots solve this task as well [@pepperberg_Can_2014].
+\citep{salwiczek_Adult_2012, zentall_Early_2017}; only African grey
+parrots solve this task as well \citep{pepperberg_Can_2014}.
 
 Recent research has refined our view on partner choice and priority of
-access in the *L. dimidiatus* cleaning mutualism, and as a consequence
+access in the \emph{L. dimidiatus} cleaning mutualism, and as a consequence
 performance in the biological market task. First, details of the task
-presentation may matter; i.e. what relevant information is encoded in
+presentation may matter; i.e.~what relevant information is encoded in
 the design of the two plates (colour versus size;
-@wismer_Cuebased_2019), or whether the information is in the colour of
-the food [@pretot_Comparative_2021; @pretot_Comparing_2016], and
+\citep{wismer_Cuebased_2019}), or whether the information is in the colour of
+the food \citep{pretot_Comparative_2021, pretot_Comparing_2016}, and
 whether the food is visible or hidden/released only when the subjects
-made choices on a touch screen [@pretot_Factors_2016]. Second, it
+made choices on a touch screen \citep{pretot_Factors_2016}. Second, it
 appears that cleaner and client densities in nature affect how cleaners
 perform in the original market task: when densities are low, cleaners
-are more likely to fail to complete the task [@triki_Biological_2019;
-@triki_Decrease_2018; @wismer_Variation_2014]. The apparent explanation
+are more likely to fail to complete the task \citep{triki_Biological_2019, triki_Decrease_2018, wismer_Variation_2014}. The apparent explanation
 for this failure is that visitor clients are less likely to exert their
 choice options (and hence wait for inspection if made to wait) when
-cleaner fish densities are low [@triki_Brain_2020]. This change in
+cleaner fish densities are low \citep{triki_Brain_2020}. This change in
 behaviour may be due to the fact that when cleaner densities are low,
@@
 have larger forebrains harbouring more cells than cleaners from
-low-density areas [@triki_Brain_2020], on a local scale those
+low-density areas \citep{triki_Brain_2020}, on a local scale those
 individuals that perform in the market task in a way that corresponds to
 the locally optimal strategy (solving/failing the task when living in
-high/low density) have relatively larger forebrains [@triki_Brain_2020].
+high/low density) have relatively larger forebrains \citep{triki_Brain_2020}.
 Taken together, empirical results clearly show major intra- and
@@
 investigation of the underlying cognitive machinery, from perception to
-decision-making [@shettleworth_Cognition_2009].
+decision-making \citep{shettleworth_Cognition_2009}.
 
@@
 their predictions in light of the behavioural data (e.g.
-@dubois_Model_2021). By varying the cognitive tool kit, one can assess
+\citep{dubois_Model_2021}). By varying the cognitive tool kit, one can assess
 which minimal kit is necessary to solve the task at hand. This approach
-has been used to analyse the market task [@prat_Role_2021;
-@quinones_Reinforcement_2019]. Learning models showed that basic
+has been used to analyse the market task \citep{prat_Role_2021, quinones_Reinforcement_2019}. Learning models showed that basic
 reinforcement learning does not suffice to solve the task, this is
@@
 the same strategic option, paired with a fish with the alternative
-strategic option) [@quinones_Reinforcement_2019]. In addition, cleaners
+strategic option) \citep{quinones_Reinforcement_2019}. In addition, cleaners
 need to account for the future consequences of current decisions. In the
 model, this could be achieved in two non-mutually exclusive ways
-[@quinones_Reinforcement_2019]: through low temporal discounting of
+\citep{quinones_Reinforcement_2019}: through low temporal discounting of
 future effects and/or through perceiving a visitor client leaving as
-psychological punishment (i.e. as a negative reinforcer). Given that a
+psychological punishment (i.e.~as a negative reinforcer). Given that a
 negative reinforcer is part of the ubiquitous learning mechanism termed
-operant conditioning [@thorndike_Animal_1898; @skinner_Behavior_1938], 
+operant conditioning \citep{thorndike_Animal_1898, skinner_Behavior_1938},
 this explanation can be
@@
 effects of an action is considered an advanced cognitive ability
-[@suddendorf_Evolution_2007]. Here, we fitted our model to the empirical
+\citep{suddendorf_Evolution_2007}. Here, we fitted our model to the empirical
 data on fish densities and cleaner fish performance in the market task
@@
 
-## Methods
+\hypertarget{methods}{%
+\subsection{Methods}\label{methods}}
 
-### Overview
+\hypertarget{overview}{%
+\subsubsection{Overview}\label{overview}}
 
-We used the data from @triki_Biological_2019 to fit the parameters of
-the model analysed in @quinones_Reinforcement_2019. For the bayesian
+We used the data from \citep{triki_Biological_2019} to fit the parameters of
+the model analysed in \citep{quinones_Reinforcement_2019}. For the bayesian
 fitting procedure we used Markov Chain Monte Carlo (MCMC) with a
@@
 
-### The model
+\hypertarget{the-model}{%
+\subsubsection{The model}\label{the-model}}
 
@@
 reward the individual obtains (prediction error)
-[@sutton_Reinforcement_2018; \@rescola_Theory_1972]; and it is carried
+\citep{sutton_Reinforcement_2018, rescorla_Theory_1972}; and it is carried
 in the direction that would lead to more reward being obtained, given
 the new information. Further details of the model implementation can be
-found in the paper [@quinones_Reinforcement_2019]. The model analysis
-points to the relevance of two parameters. First, $\gamma$ measures how
+found in the paper \citep{quinones_Reinforcement_2019}. The model analysis
+points to the relevance of two parameters. First, \(\gamma\) measures how
 much individuals include future rewards in their decision updates. If
-$\gamma=0$, individuals only use the immediate reward obtained from a
-cleaning interaction. As $\gamma$ increases individuals include more the
+\(\gamma=0\), individuals only use the immediate reward obtained from a
+cleaning interaction. As \(\gamma\) increases individuals include more the
 reward obtained from the subsequent choices. That amounts to calculating
 and using for decision making the future expected rewards in action.
-Second, $\eta$ measures how much individuals include in their reward the
+Second, \(\eta\) measures how much individuals include in their reward the
 fleeing behaviour of visitor clients as a negative component. Both of
 these parameters allow individuals to use in their estimates the future
-effects of their choices. The model analysis by
-@quinones_Reinforcement_2019 shows that without either of them, learning
+effects of their choices. The model analysis by Quiñones \emph{et al}
+\citep{quinones_Reinforcement_2019} shows that without either of them, learning
 individuals cannot develop a preference for the visitor client. The
@@
 that under very high client (relative) abundance the presence of future
-rewards ($\gamma>0$) no longer favours the preference for the ephemeral
+rewards (\(\gamma>0\)) no longer favours the preference for the ephemeral
 clients, this effect does not occur with negative reward mediated by
-client behaviour ($\eta$). Thus, the analysis suggests that future
+client behaviour (\(\eta\)). Thus, the analysis suggests that future
 reward as an explanation for cleaner ability to prefer visitors is
-supported by @triki_Biological_2019 data. However, a combination of
+supported by Triki \emph{et al} \citep{triki_Biological_2019} data. However, a combination of
 future and negative rewards, could still be a better explanation for the
 data. In the current analysis, we will use the data from
-@triki_Biological_2019 to fit some of the model parameters, particularly
-relevant we will fit $\gamma$ and $\eta$ to find out whether each or a
+\citep{triki_Biological_2019} to fit some of the model parameters, particularly
+relevant we will fit \(\gamma\) and \(\eta\) to find out whether each or a
 combination of these effects is a better explanation for the pattern
@@
 
-### Data explanation
+\hypertarget{data-explanation}{%
+\subsubsection{Data explanation}\label{data-explanation}}
 
-*Insert here brief description of the data and data collection*
+\emph{Insert here brief description of the data and data collection}
 
@@
 of clients define not only the probability of residents and visitors,
-but also how often the cleaning station is empty (e.g. there are no
+but also how often the cleaning station is empty (e.g.~there are no
 clients to be cleaned). The more cleaners there are, the more options
@@
 
-## Results
+\hypertarget{results}{%
+\subsection{Results}\label{results}}
 
-```{r post, fig.cap="Posterior distributions for parameter values $\\gamma$, $\\eta$ and the scaling constant.
-We show the kernel density estimates, below the mode (black point) and the 65\\% (light blue) and 95\\% (grey)  highest posterior density interval for the three paramteres.
+\begin{figure}
+
+{\centering \includegraphics[width=0.95\linewidth]{M:/Projects/Clean.ActCrit/post_both_gamma}
+
+}
+
+\caption{Posterior distributions for parameter values $\gamma$, $\eta$ and the scaling constant.
+We show the kernel density estimates, below the mode (black point) and the 65\% (light blue) and 95\% (grey)  highest posterior density interval for the three paramteres.
 On the left (panels a, c and c) posterior distributions for a model with   all three parameters; while on the right (b and d) a model without negative reward.
-Panel f shows a measure of fit for both models, namely the distribution of pseudo-$R^2$ obtained from sampling the posterior distribution of parameter values.",out.width='95%'}
-knitr::include_graphics(here("post_both_gamma.png"))
-
-```
+Panel f shows a measure of fit for both models, namely the distribution of pseudo-$R^2$ obtained from sampling the posterior distribution of parameter values.}\label{fig:post}
+\end{figure}
 
@@
 the bulk of the posterior distribution for the parameter tuning negative
-reward ($\eta$ panel e) is around zero. Thus, the shape of the posterior
+reward (\(\eta\) panel e) is around zero. Thus, the shape of the posterior
 and its credible intervals suggest the absence of negative reward to be
-well supported by the data. As for $\gamma$, the confidence intervals
+well supported by the data. As for \(\gamma\), the confidence intervals
 also includes zero, but the mode of the posterior is around 0.5. Thus,
-we run the analysis setting $\eta$ to zero. Figure \ref{fig:post} panels
+we run the analysis setting \(\eta\) to zero. Figure \ref{fig:post} panels
 b and d, show the posterior distributions of the two parameters
 estimated. Not surprisingly, under a model without negative reward, the
-distribution of $\gamma$ shifts to higher values. Under the new model,
-the estimates of $\gamma$ do not include zero, the distribution is
-centered around $0.5$. To asses the fit of the two models we calculated
-the $pseudo-R^2$ [@mcfadden_Conditional_1974], which is a standard
-measure of fit for logistic regression. In that context, $pseudo-R^2$
+distribution of \(\gamma\) shifts to higher values. Under the new model,
+the estimates of \(\gamma\) do not include zero, the distribution is
+centered around \(0.5\). To asses the fit of the two models we calculated
+the \(pseudo-R^2\) \citep{mcfadden_Conditional_1974}, which is a standard
+measure of fit for logistic regression. In that context, \(pseudo-R^2\)
 uses the log-likelihood of the data given the model, relative to the
@@
 measure of fit. Given that our model is not a logistic regression, we
-measure the $pseudo-R^2$ in relation to the log-likelihood of a model
-with parameters $\gamma$ and $\eta$ set to zero. This in practice
+measure the \(pseudo-R^2\) in relation to the log-likelihood of a model
+with parameters \(\gamma\) and \(\eta\) set to zero. This in practice
 amounts to a model that has neutral preference between the two options.
 In panel f of figure \ref{fig:post} we show the distribution of
-$pseudo-R^2$ calculated using samples from the posterior distributions
+\(pseudo-R^2\) calculated using samples from the posterior distributions
 shown before, as well as from a model with negative reward but without
-future reward. Note, $pseudo-R^2$ can have negative values, that is when
+future reward. Note, \(pseudo-R^2\) can have negative values, that is when
 the log-likelihood of the model is lower than that of a model that
 triggers neutral preferences. Even though, the peak of the three
-$pseudo-R^2$ distributions is not very different, the model with only
-future reward produces a distribution of $pseudo-R^2$ where more values
+\(pseudo-R^2\) distributions is not very different, the model with only
+future reward produces a distribution of \(pseudo-R^2\) where more values
 are positive (to the right of black line in Fig. \ref{fig:post} f). This
@@
 
-```{r pred, fig.cap="Observed and predicted probability of choosing a visitor.
+\begin{figure}
+
+{\centering \includegraphics[width=0.9\linewidth]{M:/Projects/Clean.ActCrit/contour_both_gamma}
+
+}
+
+\caption{Observed and predicted probability of choosing a visitor.
 Left-hand side panel: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis.
@@
 Pseudo $R^2$ shown in the lower right corner corresponds to the fit of a model parameterized with the mode of the posterior distributions.
-Upper panels show prediction from a model including future and negative reward, lower panels froma  model with only future reward.",out.width='90%'}
-
-knitr::include_graphics(here("contour_both_gamma.png"))
-
-```
+Upper panels show prediction from a model including future and negative reward, lower panels froma  model with only future reward.}\label{fig:pred}
+\end{figure}
 
@@
 the left panels. On the lower right corner of the panel we show the
-$pseudo-R^2$ for predictions using the mode of the posterior
+\(pseudo-R^2\) for predictions using the mode of the posterior
 distributions. Despite having one more parameter varying, the full model
 gives a lower fit to the data than the model without negative reward.
-The difference in $pseudo-R^2$ for the point estimates is small, but
+The difference in \(pseudo-R^2\) for the point estimates is small, but
 using the whole posterior distribution shows the improvement in fit
@@
 \ref{fig:post}f). In the supplementary material, we also show that a
-model with only negative reward, excluding future rewards ($\gamma=0$,
+model with only negative reward, excluding future rewards (\(\gamma=0\),
 Fig. \ref{fig:nogamma}), provides a lower fit than either of the ones
@@
 
-## Discussion
+\hypertarget{discussion}{%
+\subsection{Discussion}\label{discussion}}
 
@@
 relation between visitor leaving behaviour and cleaner performance in
-the market task [@triki_Biological_2019]. For both mechanisms it is
+the market task \citep{triki_Biological_2019}. For both mechanisms it is
 indeed necessary to have fleeing behaviour on the side of the visitors
@@
 cells than individuals whose performance mismatched locally optimal
-decision-rules [@triki_Brain_2020]. This suggests that despite the
+decision-rules \citep{triki_Brain_2020}. This suggests that despite the
 overall behavioural adjustment, evidenced here and elsewhere
-[@triki_Biological_2019; @triki_Decrease_2018], there are some
-cleaners better than others at adjusting their behaviour.
-@triki_Brain_2020 refer to these as socially competent cleaners. Social
+\citep{triki_Biological_2019, triki_Decrease_2018}, there are some
+cleaners better than others at adjusting their behaviour. Triki \emph{et al.}
+\citep{triki_Brain_2020} refer to these as socially competent cleaners. Social
 competence refers to the ability to optimise social behaviour depending
-on the available social information [@taborsky_Social_2012a;
-@bshary_Cooperation_2015]. However, we have shown here that cleaners
+on the available social information \citep{taborsky_Social_2012a, bshary_Cooperation_2015}. However, we have shown here that cleaners
 are able to adjust their preference according to the social environment
@@
 One potential difference between cleaners that adjusted adaptively, like
-the ones reported in @triki_Brain_2020, is configurational learning or
+the ones reported in Triki \emph{et al.} \citep{triki_Brain_2020}, is configurational learning or
 chunking. The original analysis of the learning model showed that agents
@@
 future, and to assign value to the combination of clients present in the
-station [@quinones_Reinforcement_2019]. The latter is referred to as
-configurational learning or chunking [@sutherland_Configural_1989;
-@miller_Magical_1956]. Theoretical models suggest that the lack of
+station \citep{quinones_Reinforcement_2019}. The latter is referred to as
+configurational learning or chunking \citep{sutherland_Configural_1989, miller_Magical_1956}. Theoretical models suggest that the lack of
 configurational learning is related to a low frequency with which
-stimuli are encountered together [@kolodny_Evolution_2014;
-@prat_Role_2021]. Thus, cleaner relative abundance could be related to
+stimuli are encountered together \citep{kolodny_Evolution_2014, prat_Role_2021}. Thus, cleaner relative abundance could be related to
 this mechanisms as well. Given that the lack of either of these two
-mechanisms yields neutral preferences [@quinones_Reinforcement_2019],
+mechanisms yields neutral preferences \citep{quinones_Reinforcement_2019},
 we can't, based on the current analysis, distinguish between the two
@@
 cleaners with high and low performance in the market task
-[@woodbury_Learning_1943; @sutherland_Configural_1989].
+\citep{woodbury_Learning_1943, sutherland_Configural_1989}.
 
@@
 For example, any theory of mind task needs to be solved in the first
-trial in order to exclude fast conditioning [@heyes_Theory_1998].
+trial in order to exclude fast conditioning \citep{heyes_Theory_1998}.
 Similarly, subjects need to solve a social learning task on the first
@@
 social learning, took several years to spread within the group
-[@kawamura_Process_1959], meaning that any learner had been repeatedly
-exposed to demonstrations before acquisition. Importantly,
-@galef_Question_1992 refuted imitation as mechanism not simply because
+\citep{kawamura_Process_1959}, meaning that any learner had been repeatedly
+exposed to demonstrations before acquisition. Importantly, Galef
+\citep{galef_Question_1992} refuted imitation as mechanism not simply because
 of the repeated exposure but because a (rather qualitative) analysis of
@@
 prediction based on imitation learning (see also
-@hirata_SweetPotato_2001). In our case, the number of trials it took
+\citep{hirata_SweetPotato_2001}). In our case, the number of trials it took
 cleaners to learn the solution to the market task would never allow to
@@
 
-## Cited literature
-
-::: {#refs}
-:::
-
 \newpage
 
-# Supplementary material
+\hypertarget{supplementary-material}{%
+\section{Supplementary material}\label{supplementary-material}}
 
-```{=tex}
 \beginsupplement
-```
-```{r nogamma, fig.cap="Observed and predicted probability of choosing a visitor for a model without future rewards.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.9\linewidth,]{M:/Projects/Clean.ActCrit/Simulations/MCMCclean_nRew_sca_/round_gamma_0regRew_0.13scaC_61.66alphA_0.05alphC_0.05_contour_gamma_ggplot}
+
+}
+
+\caption{Observed and predicted probability of choosing a visitor for a model without future rewards.
 Left-hand side panel: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis.
@@
 Right-hand side panel: scatter plot of the observed and predicted probabilities of choosing a visitor for 12 locations.
-Line correspond to a perfect match between observed and predicted probabilities.",out.width='90%'}
-fileNogamma<-grep("contour_gamma_",list.files(here("Simulations","MCMCclean_nRew_sca_"),
-                                              full.names = TRUE),value = TRUE)
-knitr::include_graphics(fileNogamma)
+Line correspond to a perfect match between observed and predicted probabilities.}\label{fig:nogamma}
+\end{figure}
 
-```
+\begin{longtable}[]{@{}rc@{}}
+\caption{\label{tab:param} Parameter values with which the model was run
+in the MCMC.
+\(\sigma\) refers to the amplitud of the perturbation kernel with the subscript indicating the associated parameter.
+New values were taken from a uniform distribution.
+\(\alpha\) refers to the learning rate.}\tabularnewline
+\toprule
+Parameter & Value \\
+\midrule
+\endfirsthead
+\toprule
+Parameter & Value \\
+\midrule
+\endhead
+Learning rounds & 10000 \\
+Reward value & 1 \\
+\(\alpha\) & 0.05 \\
+\(\sigma_{\gamma}\) & 0.3 \\
+\(\sigma_{\eta}\) & 4 \\
+\(\sigma_{Sca.Const.}\) & 300 \\
+Number of chains & 5 \\
+Chain lenght & \(1^5\) \\
+\bottomrule
+\end{longtable}
 
-```{r param , echo=FALSE,results='asis'}
-  cat(' Table: \\label{tab:param} Parameter values with which the model was run 
-  in the MCMC.
-$\\sigma$ refers to the amplitud of the perturbation kernel with the subscript indicating the associated parameter.
-New values were taken from a uniform distribution.
-$\\alpha$ refers to the learning rate.
+\ethics{The Animal Ethics Committee of the Queensland government (DAFF)
+approved the project (CA 2016/05/970 and CA 2017/05/1063).}
 
-  | Parameter | Value |
-|----------:|:-----:|
-| Learning rounds | 10000 |
-| Reward value | 1 |
-| $\\alpha$ | 0.05 |
-| $\\sigma_{\\gamma}$ | 0.3 |
-| $\\sigma_{\\eta}$ | 4 |
-| $\\sigma_{Sca.Const.}$ | 300 |
-| Number of chains | 5 |
-| Chain lenght | $1^5$ |
-')
+\dataccess{Please provide details on the data availability.}
 
-```
+\aucontribute{AQ and RB designed the study. ZT collected the data. AQ developed the model and
+fitted the model parameters to the data. AQ, RB and ZT wrote the manuscript.}
+
+\competing{The authors declate no conflict of interests.}
+
+\funding{This work was supported by the Swiss National Science Foundation (grants
+number: 31003A\_153067/1 and 310030B\_173334/1 to R.B.).}
+
+\ack{ZT and RB kindly thank the staff of Lizard Island Research Station, Dominique
+Roche, Sharon Wismer, Olivia Rey,Sandra Ann Binning, Elena Levorato and William McNeely for their
+field support. AQ thanks Florian Hartig for statistical advice.}
+
+\bibliographystyle{RS}
+\bibliography{Cleanerlearning.bib}
+
+\end{document}

### 2021-12-21 manFigures.R

@@ -82,7 +82,6 @@ fieldatabyLoc.gam
 myPalette <- colorRampPalette(rev(brewer.pal(11, "Spectral")))
 sc <- scale_colour_gradientn(colours = myPalette(100), limits=c(0,0.8))
 
-typeof(sc)
 
 obs.both<-ggplot(data = fieldatabyLoc.both, 
             aes(x=re.abund.clean, y=prob.Vis.leave,color=probVisi.data))+

### 2022-05-26 prog_report/report_attention.Rmd
Preliminary results on the implementation of Mackintosh and pearce and hall to the market experiment. PRogress report.

---
title: "The role of attention in learning to solve the ephemeral reward task"
author: "Andres Quiñones"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:  
  bookdown::pdf_document2:
     number_sections : false 
     toc: false 
bibliography: "../manuscript_1.0/Cleanerlearning.bib"
keep_tex: yes
header-includes:
  - \usepackage{amsmath}
---

```{r setup,include=FALSE,}
knitr::opts_chunk$set(echo = FALSE,message = FALSE)
```

## The model

Just like like in previous versions of the model, we assume that clients that 
reach the cleaning station of a cleaner fish can be classified into resident 
or visitors. Up to two of these clients can demand cleaning service simultaneously.
In such case the cleaner fish must choose which client to clean first. In order to
make that choice, the cleaner fish estimates a value for each client combination
that it faces. Hence, each client combination is a **state** in Reinforcement 
Learning (RL) jargon. In order to update the estimates and use the estimations 
based on the client types, cleaners must discriminate between 
the two types and their combinations in the cleaning stations. We, thus, assume
that each client type is associated with a specific stimulus. The cleaner fish
stores an estimate of value for each of these stimuli in a look-up table 
\ref{tab:lookup}. These estimates of individual client values are used independently
or combined depending on the state of the cleaning station. For example, 
when the cleaning station only has a resident the estimated value is 
$S_r$ of the state; while when it has a resident and a visitor the 
estimated value is $S_r+S_v$. This combination of stimuli to estimate 
value is in line with classical accounts of associative 
learning [@rescorla_Theory_1972].

The estimation and decision making are structured as in the **actor-critic** 
implementation in RL. Both the value estimation update (the critic) and the 
decision making update (the actor), change proportionally to the 
prediction error ($\delta$)

$$
\delta^t = R^t + \gamma \sum_{i}^{n^{t+1}} S_i^{t+1} - \sum_{i}^{n^t} S_i^t ,
$$
where $t$ and $t+1$ refer to the current and future cleaning 
interactions respectively; $R$ is the primary reward obtained from the 
choice made; $\gamma$ is measure of how much future reward 
influences the estimation process; and the sum terms correspond 
to the estimate of value made by summing over the client 
fish present in the corresponding cleaning
interaction. The total update of the value estimation is given by the product
of $\delta$ and $\alpha_i$, which can be interpreted as how much 
attention the cleaner fish pays to each stimuli. The fact that each stimuli 
can have a different $\alpha_i$ implies that the model implement differential 
attention. Furthermore, the actual value of $\alpha_i$ may be updated 
and depend on specific rules, which will be explained later. 
So, the update rule for the critic is given by

$$ \Delta S_i^t = \alpha^t_i \delta^t.  $$

The decision between resident and visitor is implemented as a probability, 
which is given by

$$ \pi_i = \frac{1}{1+e^{-(\theta_i-\theta_j)}}$$
where $\theta_i$ and $\theta_j$ are the preferences for each of the two 
available options. These preferences start on zero and are updated on every 
cleaning interaction where the cleaner fish chooses among of the two
options according to

$$ \Delta (\theta_i-\theta_j) = \beta_i^t \delta_t2(1-\pi_i)$$
where $\beta$ is the speed of the actor update and can be also interpreted as 
the attention that is paid to the respective stimuli in the decision-making
process.

```{r lookup}
lookupTab<-data.frame(client=c("Resident","Visitor","absence"),
                      Value=c("$S_r$","$S_v$",0))
knitr::kable(lookupTab,caption = "Look up table where a cleaner fish stores the 
             estimated values for each client type.",escape=FALSE)
```

## Attention updates

In the following sections we explain three different possible rules for 
the change in the selective attention variables.

### Static attention

As a benchmark we implement first a model where attention is not selective, 
all stimuli get the same attention and this level of attention does not vary
as learning proceeds. Figure \ref{fig:bench} shows the dynamics of the 
resident *vs* visitor choices that agents make along the learning process. 
The dynamics are discriminated according to how important future reward 
is in the estimation ($\gamma$). The dynamics show that regardless of the value 
of $\gamma$ used in the simulations, this set up does not allow cleaner fish 
to develop a preference for the visitor over the resident. Interestingly, 
under $\gamma=0$ the learning process leads to a small proference fir the 
resident option; while with $\gamma>0$ agents stay with a neutral preference.

```{r bench, fig.cap="Dynamics of the resident-visitor choices under a constant and even attention to all the stimuli"}
library("jsonlite")
library("here")
library("ggplot2")
source(here("loadData.R"))
source(here("aesth_par.R"))

scenario<-"equalAttAC1";simsDir<-"Simulations"
listPar<-c("AttMech","alph");listVal<-c(0,0.01)

# param<-list(totRounds=10000,ResReward=1,VisReward=1,
#             ResProb=c(0.3),
#             VisProb=c(0.3),
#             ResProbLeav=0,VisProbLeav=1,negativeRew=-0,
#             scenario=0,
#             inbr=0,outbr=0,trainingRep=10,forRat=0.0,
#             alphaT=0.05,printGen=1,seed=1, gammaRange=I(c(0,0.5,0.8)),
#             netaRange=I(c(0)),alphaThRange=I(c(0.05)),numlearn=1,
#             propfullPrint = 0.7,
#             # alphaThNch=0.01,
#             attenMech=0, equalAttAC=TRUE,
#             folderL=paste0(here(simsDir),scenario,"_/"))
# 
# check_create.dir(here(simsDir),param = rep(scenario,1),
#                  values = c(""))
# 
# parRange<-c(1)
# 
# for(parVal in parRange){
#   param$folderL<-paste0(here("Simulations",paste0(scenario,"_")),"/")
#   param$folder<-param$folderL #paste0(folderSims,"/",listfolders[i],"/") 
#   # param$attenMech<-parVal
#   outParam<-toJSON(param,auto_unbox = TRUE,pretty = TRUE)
#   fileName<-paste("parameters_",parVal,".json",sep="") 
#   if(file.exists(paste(param$folderL,fileName,sep = ''))){
#     currFile<-fromJSON(paste(param$folderL,fileName,sep = ''))
#     if(sum(unlist(currFile)!=unlist(param))>0){
#       write(outParam,paste(param$folderL,fileName,sep = "/"))
#     }
#   }
#   else{
#     write(outParam,paste(param$folderL,fileName,sep = ""))
#   }
# }

# runSims(paste0(scenario,"_"))
# 
# system("cd Simulations \n cd staticAtt_ \n ../.././CleanAC. exe parameters:1.json")

GamVar<-do.call(rbind,lapply(
                getFilelist(here("Simulations",paste0(scenario,"_")),
                            listparam = listPar,
                            values = listVal,
                            fullNam = TRUE)$PAA,
                file2timeInter,interV=501,parOFint="AttMech"))

ggplot(data = GamVar,aes(x=Interv,y=Prob.RV.V,group=Gamma,
                         color=as.factor(Gamma)))+
        stat_summary(fun.data = function(x){
                x.r<-fivenum(x)
                return(list(ymin=x.r[2],ymax=x.r[4],y=x.r[3]))
        },geom ="pointrange",position = position_dodge(0.9))+
        geom_hline(yintercept = 0.5,color="red")+
        geom_line(aes(group=interaction(Gamma,Training),
                  color=as.factor(Gamma)),alpha=0.2,size=0.2,
                  position = position_dodge(0.9))+
        scale_color_manual(values =colorValues, name = expression(gamma))+
        xlab("Time")+ylab("Probability of V over R")+
        theme_classic()
```

### The Mackintosh update

@mackintosh_Theory_1975 proposed a model of attentional dynamics, where the 
central idea is that stimuli which are the best predictors of reward get an
increase in the attention they enjoy during the learning process. Formally,
@mackintosh_Theory_1975 defined the attentional update as

\begin{align*}
\Delta \alpha_i &> 0 \text{ if } |\lambda - V_i | <  |\lambda - V_P| \\
\Delta \alpha_i &< 0 \text{ if } |\lambda - V_i | < |\lambda- V_P|
\end{align*}

Where $\lambda$ is the "real" associative value in a given trial, which translates
to RL jargon as the reward; $V_i$ is the associative strength of the focal stimulus;
and $V_p$ is sum of the associative strengths of the stimuli other than the focal.

Following a notation more in line with the current model, we implemented the 
@mackintosh_Theory_1975 idea as follows

$$
\alpha_i^{t+1} = \hat{\alpha}(|R^t + \gamma \sum_{j}^{n^{t+1}} S_j^{t+1}-\sum_{j\neq i}^{n^{t+1}}S_j^{t}|-
|R^t + \gamma \sum_{j}^{n^{t+1}} S_j^{t+1}-S_i^t|)
$$

```{r Mackintosh,fig.cap="Dynamics of the resident-visitor choices under a update inspired by @mackintosh_Theory_1975"}
library(cowplot)
scenario<-"equalAttAC1"
listPar<-c("AttMech","alph");listVal<-c(1,0.01)
list1<-getFilelist(here("Simulations",paste0(scenario,"_")),
            listparam = listPar,
            values = listVal,
            fullNam = TRUE)$PAA
TimeInterv<-do.call(rbind,lapply(
                list1,
                file2timeInter,interV=501,parOFint="AttMech"))

ggplot(data = TimeInterv,aes(x=Interv,y=Prob.RV.V,group=Gamma,
                         color=as.factor(Gamma)))+
        stat_summary(fun.data = function(x){
                x.r<-fivenum(x)
                return(list(ymin=x.r[2],ymax=x.r[4],y=x.r[3]))
        },geom ="pointrange",position = position_dodge(0.9))+
        geom_hline(yintercept = 0.5,color="red")+
        geom_line(aes(group=interaction(Gamma,Training),
                  color=as.factor(Gamma)),alpha=0.2,size=0.2,
                  position = position_dodge(0.9))+
        scale_color_manual(values =colorValues, name = expression(gamma))+
        xlab("Time")+ylab("Probability of V over R")+
        theme_classic()

```

```{r attenDynMac,fig.cap="Dynamics of attention to the different stimuli under the model inspired by  @mackintosh_Theory_1975"}
# Load raw data
rawdynamics<-do.call(rbind,lapply(list1,loadFilewithPar,parOFint="AttMech"))

# Transform associability (\alpha) data to long format
long_alpha<-melt(rawdynamics[Age%%200==0],
                    id.vars = c("Age","Training","Gamma","AttMech"),
                    variable.name = "Client",
                    measure.vars = c("alphaRes","alphaVis"))

long_alpha[,Client:=as.factor(Client)]

ggplot(data=long_alpha,aes(x=Age,y=value,colour=Client,
                              group=as.factor(Client)))+
        stat_summary(fun.data = function(x){
                x.r<-fivenum(x)
                return(list(ymin=x.r[2],ymax=x.r[4],y=x.r[3]))
        },geom = "pointrange")+
        geom_line(aes(group=interaction(Client,Training)),
                  alpha=0.2,size=0.2,)+
                facet_grid(Gamma~Client,scales = "free_y",
                           labeller = labeller(Client=NULL))+
        scale_color_manual(values =colorValues,name="Client",
                           labels=c("Resident","Visitor"))+
  scale_y_continuous(sec.axis = sec_axis(~ . , name = expression(gamma),
                                         breaks = NULL, labels = NULL))+
      theme_classic()+
        theme(legend.position = "top",strip.background = element_blank(),
   strip.text.x = element_blank())

```

### The Pearce and Hall

@pearce_Model_1980 proposed a model of attentional dynamics, where the 
central idea is that the level of attention is proportional to the difference
between the real associative strength and that predicted by the stimulus. 
Formally, @pearce_Model_1980 defined the attentional update as

$$
\alpha^t = \rho|\lambda^{t-1}-V^{t-1}_{\sum}| + (1-\rho)\alpha^{t-1},
$$

where $V_{\sum}^{t-1}$ represents the total associative strength triggered
by the stimulus; and $\rho$ is a constant that measure how fast are the 
jumps in attention.

Following a notation more in line with the current model, we implemented the 
@pearce_Model_1980 idea as follows

$$
\alpha_i^{t+1} = \hat{\alpha}|R^t + \gamma \sum_{j}^{n^{t+1}}S_i^{t+1}-S_i^t|
$$

```{r PandH,fig.cap="Dynamics of the resident-visitor choices under a update inspired by @pearce_Model_1980"}
scenario<-"equalAttAC1"
listPar<-c("AttMech","alph");listVal<-c(2,0.01)
list1<-getFilelist(here("Simulations",paste0(scenario,"_")),
            listparam = listPar,
            values = listVal,
            fullNam = TRUE)$PAA
GamVar<-do.call(rbind,lapply(list1,
                file2timeInter,interV=501,parOFint="AttMech"))

ggplot(data = GamVar,aes(x=Interv,y=Prob.RV.V,group=Gamma,
                         color=as.factor(Gamma)))+
        stat_summary(fun.data = function(x){
                x.r<-fivenum(x)
                return(list(ymin=x.r[2],ymax=x.r[4],y=x.r[3]))
        },geom ="pointrange",position = position_dodge(0.9))+
        geom_hline(yintercept = 0.5,color="red")+
        geom_line(aes(group=interaction(Gamma,Training),
                  color=as.factor(Gamma)),alpha=0.2,size=0.2,
                  position = position_dodge(0.9))+
        scale_color_manual(values =colorValues, name = expression(gamma))+
        xlab("Time")+ylab("Probability of V over R")+
        theme_classic()

```

```{r attenDynPandH,fig.cap="Dynamics of attention to the different stimuli under the model inspired by  @pearce_Model_1980"}

# Load raw data
rawdynamics<-do.call(rbind,lapply(list1,loadFilewithPar,parOFint="AttMech"))

# Transform associability (\alpha) data to long format
long_alpha<-melt(rawdynamics[Age%%200==0],
                    id.vars = c("Age","Training","Gamma","AttMech"),
                    variable.name = "Client",
                    measure.vars = c("alphaRes","alphaVis"))

long_alpha[,Client:=as.factor(Client)]

ggplot(data=long_alpha,aes(x=Age,y=value,colour=Client,
                              group=as.factor(Client)))+
        stat_summary(fun.data = function(x){
                x.r<-fivenum(x)
                return(list(ymin=x.r[2],ymax=x.r[4],y=x.r[3]))
        },geom = "pointrange")+
        geom_line(aes(group=interaction(Client,Training)),
                  alpha=0.2,size=0.2,)+
                facet_grid(Gamma~Client,scales = "free_y",
                           labeller = labeller(Client=NULL))+
        scale_color_manual(values =colorValues,name="Client",
                           labels=c("Resident","Visitor"))+
  scale_y_continuous(sec.axis = sec_axis(~ . , name = expression(gamma),
                                         breaks = NULL, labels = NULL))+
      theme_classic()+
        theme(legend.position = "top",strip.background = element_blank(),
   strip.text.x = element_blank())

```

## An alternative Mackintosh implementation

During the development of the model an alternative implementation to the @mackintosh_Theory_1975. This alternative implementation captures the original conceptual idea. However, it implements it in a slightly different way:

$$
\alpha_i^{t+1} = \hat{\alpha}(|R^t + \gamma \sum_{j}^{n^{t+1}} S_j^{t+1}-\sum_{j}^{n^{t+1}}S_j^{t}|-
|R^t + \gamma \sum_{j}^{n^{t+1}} S_j^{t+1}-S_i^t|)
$$

```{r MackAltRV,fig.cap="Dynamics of the resident-visitor choices under a update inspired by a model with an alternative implementation similar to @mackintosh_Theory_1975"}
scenario<-"equalAttAC1"
listPar<-c("AttMech","alph");listVal<-c(3,0.01)
list1<-getFilelist(here("Simulations",paste0(scenario,"_")),
            listparam = listPar,
            values = listVal,
            fullNam = TRUE)$PAA
GamVar<-do.call(rbind,lapply(list1,
                file2timeInter,interV=501,parOFint="AttMech"))

ggplot(data = GamVar,aes(x=Interv,y=Prob.RV.V,group=Gamma,
                         color=as.factor(Gamma)))+
        stat_summary(fun.data = function(x){
                x.r<-fivenum(x)
                return(list(ymin=x.r[2],ymax=x.r[4],y=x.r[3]))
        },geom ="pointrange",position = position_dodge(0.9))+
        geom_hline(yintercept = 0.5,color="red")+
        geom_line(aes(group=interaction(Gamma,Training),
                  color=as.factor(Gamma)),alpha=0.2,size=0.2,
                  position = position_dodge(0.9))+
        scale_color_manual(values =colorValues, name = expression(gamma))+
        xlab("Time")+ylab("Probability of V over R")+
        theme_classic()

```

```{r MackAltAtt,fig.cap="Dynamics of attention to the different stimuli under a model inspired by  @mackintosh_Theory_1975"}

# Load raw data
rawdynamics<-do.call(rbind,lapply(list1,loadFilewithPar,parOFint="AttMech"))

# Transform associability (\alpha) data to long format
long_alpha<-melt(rawdynamics[Age%%200==0],
                    id.vars = c("Age","Training","Gamma","AttMech"),
                    variable.name = "Client",
                    measure.vars = c("alphaRes","alphaVis"))

long_alpha[,Client:=as.factor(Client)]

ggplot(data=long_alpha,aes(x=Age,y=value,colour=Client,
                              group=as.factor(Client)))+
        stat_summary(fun.data = function(x){
                x.r<-fivenum(x)
                return(list(ymin=x.r[2],ymax=x.r[4],y=x.r[3]))
        },geom = "pointrange")+
        geom_line(aes(group=interaction(Client,Training)),
                  alpha=0.2,size=0.2,)+
                facet_grid(Gamma~Client,scales = "free_y",
                           labeller = labeller(Client=NULL))+
        scale_color_manual(values =colorValues,name="Client",
                           labels=c("Resident","Visitor"))+
  scale_y_continuous(sec.axis = sec_axis(~ . , name = expression(gamma),
                                         breaks = NULL, labels = NULL))+
      theme_classic()+
        theme(legend.position = "top",strip.background = element_blank(),
   strip.text.x = element_blank())

```

# References

### 2022-05-26 ActCrit.cpp

@@ -22,7 +22,7 @@ types of agent. Fully Aware Agents (FAA) estimate value for
 
 Written by:
 
-Andr?s E. Qui?ones
+Andres E. Quinones
 Posdoctoral researcher
 Behavioural Ecology Group
 Institute of Biology
@@ -66,7 +66,7 @@ class agent													// Learning agent
 {
 public:
 	agent(double alphaI, double gammaI, bool netaI, 
-		double alphathI, double initVal);
+		double alphathI, double initVal, bool equalAttI);
 	// constructor providing values for the learning parameters
 	~agent();																
 	// destructor not really necessary
@@ -144,13 +144,14 @@ protected:
 	double cumulReward;	// Cumulative reward
 	int age;
 	double negReward;
+	bool equalAttAC; // TRUE if both Actor and Critic use the same \alpha
 };
 
 // Members of agent class
 
 agent::agent(double alphaI = 0.01, double gammaI = 0.5, 
 	bool netaI = 0, double alphathI = 0.01,
-	double initVal = 0){
+	double initVal = 0, bool equalAttI = true){
 // parameterized constructor with defaults
 	theta[0] = 0, theta[1] = 0;
 	numEst = 6;
@@ -171,6 +172,7 @@ agent::agent(double alphaI = 0.01, double gammaI = 0.5,
 	currentReward = 0, cumulReward = 0;
 	age = 0;
 	countExp = 2;
+	equalAttAC = equalAttI;
 }
 
 void agent::rebirth(double initVal = 0)
@@ -457,8 +459,47 @@ void agent::updateAlpha(int attenMech, int currState, double totValcurr,
 			break;
 		case 1:
 			if (cleanOptionsT[0] == 0 || cleanOptionsT[1] == 0) {
-				alphas[0] = alpha*(abs(lambda - totValcurr)-
+				alphas[0] = alpha*(abs(lambda - totValcurr + values[0])-
+				//alphas[0] = alpha*(abs(lambda - values[1]) -
 					abs(lambda-values[0]));
+				clip_low(alphas[0], 0);
+				if (isnan(alphas[0])) {
+					wait_for_return();
+				}
+			}
+			if (cleanOptionsT[0] == 1 || cleanOptionsT[1] == 1) {
+				alphas[1] = alpha * (abs(lambda - totValcurr + values[0]) -
+					abs(lambda - values[1]));
+				clip_low(alphas[1], 0);
[…]

### 2022-05-26 loadData.R

@@ -112,8 +112,11 @@ getParam<-function(folder,agent,listparam=NULL,values=NULL){
 # divides raw data in interval according to interV and calculates 
 # the RV preference is such interval
 
-file2timeInter<-function(filename,interV,maxAge=-2){
-  extPar<-strsplit(filename,split ="_/")[[1]][1]
+file2timeInter<-function(filename,interV,maxAge=-2,parOFint){
+  extPar<-strsplit(filename,split ="/")[[1]]
+  extPar<-grep(parOFint,x = extPar,value = TRUE)
+  extPar<-strsplit(extPar,split ="_")[[1]]
+  extPar<-grep(parOFint,x = extPar,value = TRUE)
   parVal<-as.numeric(gsub("[[:alpha:]]",extPar,replacement = ''))
   extPar<-gsub("[[:digit:]]",extPar,replacement = '')
   tmp<-fread(filename,nrows = maxAge+1)
@@ -127,6 +130,20 @@ file2timeInter<-function(filename,interV,maxAge=-2){
     tmptimeInter[,eval(extPar):=parVal]
   }
   return(tmptimeInter)
+}
+
+loadFilewithPar<-function(filename,interV,maxAge=-2,parOFint){
+  extPar<-strsplit(filename,split ="/")[[1]]
+  extPar<-grep(parOFint,x = extPar,value = TRUE)
+  extPar<-strsplit(extPar,split ="_")[[1]]
+  extPar<-grep(parOFint,x = extPar,value = TRUE)
+  parVal<-as.numeric(gsub("[[:alpha:]]",extPar,replacement = ''))
+  extPar<-gsub("[[:digit:]]",extPar,replacement = '')
+  tmp<-fread(filename,nrows = maxAge+1)
+  if(length(extPar)>0){
+    tmp[,eval(extPar):=parVal]
+  }
+  return(tmp)
 }
 
 # Computes RV preference for a certain final proportion of the learning trial --
@@ -224,6 +241,31 @@ loadMCMCrun<-function(scen,burn.in=1000,thinning=100){
   return(filtered)
 }
   
+# Function to recursively filter a list of files
+
+greploop<-function(x,mylist,and=TRUE){
+  newlist<-mylist
+  for(i in x){
+    newlist<-grep(pattern = i,x=newlist,value = TRUE)
+  }
+  return(newlist)
+}
   
+# Function to run simulations from parameter files in a folder -----------------
+
[…]

### 2022-05-26 parTOjson.R

@@ -6,11 +6,11 @@ source(here("loadData.R"))
 
 simsDir<-"Simulations"
 
-exedir<-'/./ActCrit.exe'
+# exedir<-'/./ActCrit.exe'
 
 fileName<-"parameters.json"
 
-scenario<-"MCMCclean_gam_nRew_sca_group_"
+
 
 
 # MCMC fit - Generate json parameter files for -------------------------------------
@@ -265,8 +265,47 @@ for (i in 1:length(rangLeav)) {
     #   ,sep = " "))
   }
 }
-# 
-for (k in 1:90) print(y)
+#
+
+## Create Job files for individuaL runs and single parameter variation ---------
+
+scenario<-"equalAttAC1"
+
+param<-list(totRounds=10000,ResReward=1,VisReward=1,
+            ResProb=c(0.3),
+            VisProb=c(0.3),
+            ResProbLeav=0,VisProbLeav=1,negativeRew=-0,
+            scenario=0,
+            inbr=0,outbr=0,trainingRep=10,forRat=0.0,
+            alphaT=0.01,printGen=1,seed=1, gammaRange=I(c(0,0.5,0.8)),
+            netaRange=I(c(0)),alphaThRange=I(c(0.05)),numlearn=1,
+            propfullPrint = 0.7,
+            # alphaThNch=0.01,
+            attenMech=1, equalAttAC=TRUE,
+            folderL=paste0(here(simsDir),scenario,"_/"))
+
+check_create.dir(here(simsDir),param = rep(scenario,1),
+                 values = c(""))
+
+parRange<-c(0,1,2,3)
+
+for(parVal in parRange){
+  param$folderL<-paste0(here("Simulations",paste0(scenario,"_")),"/")
+  param$folder<-param$folderL #paste0(folderSims,"/",listfolders[i],"/") 
+  param$attenMech<-parVal
+  outParam<-toJSON(param,auto_unbox = TRUE,pretty = TRUE)
+  fileName<-paste("parameters_",parVal,".json",sep="") 
+  if(file.exists(paste(param$folderL,fileName,sep = ''))){
+    currFile<-fromJSON(paste(param$folderL,fileName,sep = ''))
+    if(sum(unlist(currFile)!=unlist(param))>0){
+      write(outParam,paste(param$folderL,fileName,sep = "/"))
+    }
+  }
+  else{
+    write(outParam,paste(param$folderL,fileName,sep = ""))
+  }
+}
+
 
 
 

### 2022-05-26 runSim2.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+
+cd Simulations
+cd $1
+
+
+nohup ../.././CleanAC.exe $2 &

### 2022-05-26 runsSims.sh

@@ -1,0 +1,8 @@
+#!/bin/bash
+
+cd Simulations
+cd $1
+
+
+echo find . -name "*.json"
+find . -name "*.json" | xargs -n 1 -P $2 nohup ../.././CleanAC.exe &

### 2022-05-26 test-discrim.R

@@ -8,20 +8,32 @@ source(here("loadData.R"))
 source(here("aesth_par.R"))
 
 # Files to be loaded
-list1<-list.files(here("Simulations","test_"),full.names = TRUE)
+scenario<-"equalAttAC1_"
+
+parFocal<-"AttMech"
+
+listPar<-c("AttMech","alph")
+listVal<-c(0,0.01)
+(list1<-getFilelist(here("Simulations",scenario),
+            listparam = listPar,
+            values = listVal,
+            fullNam = TRUE)$PAA)
+
+
+# list1<-greploop(params,mylist = list1)
 
 # Load raw data
-Mckintosh<-do.call(rbind,lapply(list1,fread))
+Mckintosh<-do.call(rbind,lapply(list1,loadFilewithPar,parOFint="AttMech"))
 
 # Transform value data to long format
 longMck_Value<-melt(Mckintosh[Age%%100==0],
-                    id.vars = c("Age","Training","Gamma"),
+                    id.vars = c("Age","Training","Gamma","AttMech"),
                     variable.name = "Client",
                     measure.vars = c("Resident","Visitor"))
 
 # Transform associability (\alpha) data to long format
 longMck_alpha<-melt(Mckintosh[Age%%500==0],
-                    id.vars = c("Age","Training","Gamma"),
+                    id.vars = c("Age","Training","Gamma","AttMech"),
                     variable.name = "Client",
                     measure.vars = c("alphaRes","alphaVis"))
 
@@ -35,7 +47,7 @@ ggplot(data=longMck_Value,aes(x=Age,y=value,colour=as.factor(Client),
         geom_line(aes(group=interaction(Client,Training)),
                   alpha=0.2,size=0.2)+
         scale_color_manual(values =colorValues)+
-        facet_wrap(~Gamma)+
+        facet_wrap(~Gamma*AttMech,scales = "free_y")+
         theme_classic()
 
 ggplot(data=longMck_alpha,aes(x=Age,y=value,colour=as.factor(Client),
@@ -46,22 +58,17 @@ ggplot(data=longMck_alpha,aes(x=Age,y=value,colour=as.factor(Client),
         },geom = "pointrange")+
         geom_line(aes(group=interaction(Client,Training)),
                   alpha=0.2,size=0.2,)+
-                facet_wrap(~Gamma*Client,nrow = 3)+
+                facet_wrap(~Gamma*Client*AttMech,nrow = 3,scales = "free_y")+
[…]
